### 2019-10-09 docs/manuscript/draft_manuscript.tex
first commit

\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={The title},
            pdfauthor={First Author~\& Ernst-August Doelle},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{The title}
    \author{First Author\textsuperscript{1}~\& Ernst-August Doelle\textsuperscript{1,2}}
    \date{}

\shorttitle{Title}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Wilhelm-Wundt-University\\\textsuperscript{2} Konstanz Business School}
\keywords{keywords\newline\indent Word count: X}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers

\authornote{Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

Enter author note here.

Correspondence concerning this article should be addressed to First Author, Postal address. E-mail: \href{mailto:my@email.com}{\nolinkurl{my@email.com}}}

\abstract{
One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.

Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.

One sentence clearly stating the \textbf{general problem} being addressed by this particular study.

One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).

Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.

One or two sentences to put the results into a more \textbf{general context}.

Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.

}

\begin{document}
\maketitle

\hypertarget{methods}{%
\section{Methods}\label{methods}}

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

\hypertarget{material}{%
\subsection{Material}\label{material}}

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

We used R (Version 3.6.0; R Core Team, 2019) for all our analyses.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-R-base}{}%
R Core Team. (2019). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}

\endgroup

\end{document}

### 2019-10-09 docs/slides/vertical_intro.Rmd
first commit

---
title: "Reproducible psychological research and project management in R"
subtitle: "A solution stack"
author: "Matthew Crump"
date: "2018/07/20 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: ["slides_css/defaultb.css", "slides_css/metropolisb.css", "slides_css/metropolis-fontsb.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, echo=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning=FALSE, cache = TRUE)
```

class: pink, center, middle, clear

# Everything in its right place - Radiohead

---

# Hodge Podge Management

Using lots of different programs to accomplish different jobs

1. Experiment builders (E-prime, Presentation, PsychoPy, in-house software, etc.)
2. Data Analysis (Excel, SPSS, R, Python, Matlab, etc.)
3. Write up (Word, Latex, R markdown)
4. Presentation (powerpoint, keynote)

Outcome: lots of different files everywhere, lots of work by hand, **can be very difficult to reconstruct or redo specific steps**

---

# Solution Stack Management

From Wikipedia: In computing, a **solution stack or software stack** is a set of software subsystems or components needed to create a complete platform such that no additional software is needed to support applications. Applications are said to "run on" or "run on top of" the resulting platform.

---

# R as a solution stack

1. Program Experiments (exprmntr and JsPsych)
2. Data analysis (e.g., tidyverse)
3. Graphing (ggplot2)
4. Reproducible reports (R markdown)
5. Slide decks (xaringan)
6. APA papers (papaja)
7. Sharing custom functions (R packages)
8. Sharing everything all in one place (R markdown websites, Github, and pkgdown)

---

# R as a control center

Having a full solution stack in R doesn't mean you can't use other things.

R can often be used as a control center to run other code (e.g., python, C++, etc.)

---

### 2019-10-09 slides/vertical_intro.Rmd
first commit

--- 2019-10-09 docs/slides/vertical_intro.Rmd
@@
 ---
+
+# R markdown (.Rmd files)
+
+A text file for **notes and R scripts** that compiles to different formats (html, pdf, etc.)
+
+```{r}
+knitr::include_graphics("imgs/rmarkdown.png")
+```
+
+---
+
+# Data frames
+
+```{r, echo=TRUE}
+all_data <- data.frame(trial = 1:6,
+                       congruency = rep(c("con","inc"), each = 3),
+                       rt = rnorm(6, mean = 0, sd = 100),
+                       accuracy = c(1,1,1,1,0,1))
+all_data
+```
+
+---
+
+# dplyr - tidyverse Data Wrangling
+
+```{r, echo=T}
+library(tidyverse)
+df <- all_data %>%
+  filter(accuracy == 1) %>%
+  group_by(congruency) %>%
+  summarise(mean_rt = mean(rt))
+```
+
+---
+
+# ggplot2 - tidyverse plotting
+
+```{r, echo=TRUE, fig.height=3}
+ggplot(df, aes(x=congruency, y=mean_rt))+
+  geom_bar(stat="identity", position="dodge") +
+  theme_classic()
+```

### 2019-10-09 Package Helper.R

@@ -1,0 +1,19 @@
+# Package helper
+library(usethis) # helpful automation functions for common tasks in package development
+library(pkgdown) # compile your R package to a website
+
+pkgdown::build_site()
+
+# generate description file with your info (need to set your info in advanve)
+usethis::use_description()
+
+# generate readme file, the contents control front page of pkgdown website
+usethis::use_readme_md()
+
+# create a vignette (.Rmd) document that gets compiled and displayed on pkdown website
+usethis::use_vignette("get_started", title="Getting Started")
+
+pkgdown::template_navbar() # generate yml for controlling navbar
+
+render_folder_to_docs("slides")
+

### 2019-10-09 R/render_folder_to_docs.R

@@ -1,0 +1,17 @@
+render_folder_to_docs <- function(folder_name,
+                                  dest_folder = "docs",
+                                  assets_folder){
+
+  file_names <- list.files(folder_name, pattern = "\\.Rmd$")
+
+  for(i in file_names){
+    rmarkdown::render(paste(folder_name,file_names[1],sep="/"),
+                      output_dir = paste(".",dest_folder,folder_name,sep="/"))
+  }
+
+  if (missing(assets_folder) == FALSE){
+
+  }
+}
+
+

### 2019-10-09 experiments/Experiment 1/experiment/xprmntr/xprmntr.js

@@ -1,0 +1,15 @@
+/**
+ * xprmntr.js
+ * Danielle Navarro
+ *
+ **/
+
+var xprmntr = {};
+xprmntr.save_locally = function() {
+  var data = jsPsych.data.get().csv();
+  var file = "xprmntr_local_name";
+  var xhr = new XMLHttpRequest();
+  xhr.open('POST', 'submit');
+  xhr.setRequestHeader('Content-Type', 'application/json');
+  xhr.send(JSON.stringify({filename: file, filedata: data}));
+};

### 2019-10-09 manuscript/example2/ReproducibleReport.Rmd
update papaja example

---
title: "Reproducible Report"
author: "Matt"
date: "1/31/2019"
output: 
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
```

This is an example of a reproducible report for part I of the PSYC 7709 midterm assignment.

## Introduction

This report re-produces the analysis of Experiment 3 reported in Rosenbaum, Mama, Algom (2017). The citation for the article is:

Rosenbaum, D., Mama, Y., & Algom, D. (2017). Stand by your Stroop: Standing up enhances selective attention and cognitive control. Psychological science, 28(12), 1864-1867.

The data were downloaded from [https://osf.io/b7x8q/](https://osf.io/b7x8q/)

Rosenbaum et al. (2017) had participants perform a Stroop task in one of two posture conditions. Participants either sat and performed the Stroop task, or stood and performed the Stroop task. The question was whether the size of the Stroop effect would change as a function of posture. The Stroop effect is measued as a difference between reaction times on congruent vs. incongruent trials.

The experiment involved a 2 (Posture: sitting vs standing) x 2 (congruency: congruent vs. incongruent) repeated measures design.

## Re-analysis

```{r}
# load libraries and data
library(data.table)
library(dplyr)
library(ggplot2)
all_data <- fread("data/stroop_stand.csv")

# pre-processing and conversion to long format
RTs <- c(as.numeric(unlist(all_data[,1])),
         as.numeric(unlist(all_data[,2])),
         as.numeric(unlist(all_data[,3])),
         as.numeric(unlist(all_data[,4]))
         )

Congruency <- rep(rep(c("Congruent","Incongruent"),each=50),2)
Posture <- rep(c("Stand","Sit"),each=100)
Subject <- rep(1:50,4)

stroop_df <- data.frame(Subject,Congruency,Posture,RTs)

```

### table of means for each condition

```{r}
overall_means <- stroop_df %>%
                  group_by(Posture,Congruency) %>%
                  summarise(meanRT = mean(RTs),
                            SEMRT = (sd(RTs)/sqrt(length(RTs))))

# make a table of overall means
knitr::kable(overall_means)
```

### Repoducing the figure

```{r}
ggplot(overall_means, aes(x=Posture,
                          y=meanRT, 
                          group=Congruency,
                          fill=Congruency))+
  geom_bar(stat="identity",position="dodge")+
  theme_classic(base_size=12)+
  ylab("Mean Reaction Time (ms)")+
  geom_errorbar(aes(ymin=meanRT-SEMRT,
                    ymax=meanRT+SEMRT),
                position=position_dodge(width=0.9),
                width=.2,
                color="black")+
  coord_cartesian(ylim=c(750,1000))
```

### Repeated measure ANOVA

```{r}
# Make sure Subject is  a factor
stroop_df$Subject <-  as.factor(stroop_df$Subject)

aov_out <- aov(RTs~Posture*Congruency + Error(Subject/(Posture*Congruency)), stroop_df)

# prints a nicer ANOVA table
summary_out <- summary(aov_out)
library(xtable)
knitr::kable(xtable(summary_out))

# prints means for each effect
print(model.tables(aov_out,"means"), format="markdown")

```

### Results

Mean reaction times for each subject in each condition to a 2 (Congruency: congruecnt vs. incongruent) x 2 (Posture: Standing vs. Sitting) were submitted to a repeated measures ANOVA.

There was a main effect of Congruency, F (1, 49) = 342.45, MSE = 1684.39, p < 0.001. Mean reaction times were slower for incongruent (922 ms) than congruent groups (815 ms).

There main effect of Posture was significant, F (1, 49) = 7.33, MSE = 4407.09, p =.009. Mean reaction times were slower for sitting (881 ms) than standing groups (855 ms).

The two-way interaction between Congruency and Posture was significant, F (1, 49) = 8.96, MSE = 731.82, p < 0.004. The Stroop effect was 23 ms smaller in the standing than sitting conditions.

## Discussion

The re-analysis succesfully reproduced the analysis reported by Rosenbaum, Mama, Algom (2017)

### 2019-10-09 docs/slides/vertical_intro.Rmd
update slides

@@
 
-class: pink, center, middle, clear
+# link to this website
 
-# Everything in its right place - Radiohead
+Contains these slides and example of using R to share many components of a research project
+
+<https://crumplab.github.io/vertical/>
 
@@
 
-# Hodge Podge Management
+# Online Resources for R
+
+Lots of great and free online resources. I've been making a list:
+
+<https://crumplab.github.io/blogposts/LearningR/LearningR.html>
+
+---
+
+# Everything everywhere workflow
 
@@
 
-# Solution Stack Management
-
-From Wikipedia: In computing, a **solution stack or software stack** is a set of software subsystems or components needed to create a complete platform such that no additional software is needed to support applications.
-Applications are said to "run on" or "run on top of" the resulting platform.
-
----
-
-# R as a solution stack
+# vertical: R for everything
 
@@
 
-# R as a control center
+# R markdown (.Rmd files)
 
-Having a full solution stack in R doesn't mean you can't use other things.
+A text file for **notes and R scripts** that compiles to different formats (html, pdf, etc.)
 
-R can often be used as a control center to run other code (e.g., python, C++, etc.)
+```{r}
+knitr::include_graphics("imgs/rmarkdown.png")
+```
 
 ---
+
+# R Markdown guide
+
+Definitive guide to R Markdown
+
+<https://bookdown.org/yihui/rmarkdown/>
+
+---
+
+# Experiments in R (jsPsych)
+
+- jsPsych documentation <https://www.jspsych.org>
+- xprmntr R package (in dev) <https://github.com/CrumpLab/xprmntr>
+- examples <https://github.com/CrumpLab/jsPsychRexamples>
+
+---
+
+# Data analysis
+
+- tidyverse <https://www.tidyverse.org> lots of great packages for data analysis
+- dplyr <https://dplyr.tidyverse.org> data-wrangling
+- ggplot2 <https://ggplot2.tidyverse.org> plotting
+
+---
+
+# Data frames
+
+```{r, echo=TRUE}
+all_data <- data.frame(trial = 1:6,
+                       congruency = rep(c("con","inc"), each = 3),
+                       rt = rnorm(6, mean = 0, sd = 100),
+                       accuracy = c(1,1,1,1,0,1))
+all_data
+```
+
+---
+
+# dplyr - tidyverse Data Wrangling
+
+```{r, echo=T}
+library(tidyverse)
+df <- all_data %>%
+  filter(accuracy == 1) %>%
+  group_by(congruency) %>%
+  summarise(mean_rt = mean(rt))
+df
+```
+
+---
+
+# ggplot2 - tidyverse plotting
+
+```{r, echo=TRUE, fig.height=3}
+ggplot(df, aes(x=congruency, y=mean_rt))+
+  geom_bar(stat="identity", position="dodge") +
+  theme_classic()
+```
+
+---
+
+# slide decks
+
+This slide deck was written in R markdown the`xaringan` package
+
+<https://github.com/yihui/xaringan>
+
+---
+
+# APA papers
+
+`papaja` is an R package for writing APA papers in R markdown (then compile to .pdf)
+
+<https://crsh.github.io/papaja_man/>
+
+---
+
+# R packages and pkgdown
+
+`pkgdown` is an R package that builds a website for R packages.
+
+<https://pkgdown.r-lib.org>
+
+---
+
+# bookdown
+
+A package for writing books (compile to web-book, .pdf, and epub)
+
+<https://bookdown.org>
+
+---
+
+# R markdown websites
+
+Write in R markdown, and make all kinds of websites
+
+<https://rmarkdown.rstudio.com/lesson-13.html>

### 2019-11-08 inst/vertical/docs/manuscript/example2/APAreport.Rmd
add project template

---
title             : "Reproducing the analysis of Rosenbaum, Mama, & Algom (2017)"
shorttitle        : "Stroop Analysis"

author: 
  - name          : "Matthew Crump"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "2900 Bedford Ave"
    email         : "mcrump@brooklyn.cuny.edu"

affiliation:
  - id            : "1"
    institution   : "Brooklyn College of the City University of New York"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

abstract: |
  A reproduction of the analysis for Experiment 3 from Rosenbaum, Mama, and Algom (2017).

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "Stroop, Reproducibilty"
wordcount         : "X"

bibliography      : ["r-references.bib","myreferences.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

This is an example of creating an APA manuscript using papaja, and also an example for part 2 of the midterm assignment. Another sentence.

This report re-produces the analysis of Experiment 3 reported in @rosenbaum2017stand. The data were downloaded from [https://osf.io/b7x8q/](https://osf.io/b7x8q/)

@rosenbaum2017stand had participants perform a Stroop task in one of two posture conditions. Participants either sat and performed the Stroop task, or stood and performed the Stroop task. The question was whether the size of the Stroop effect would change as a function of posture. The Stroop effect is measured as a difference between reaction times on congruent vs. incongruent trials.The experiment involved a 2 (Posture: sitting vs standing) x 2 (congruency: congruent vs. incongruent) repeated measures design.

# Methods

## Participants

There were 50 participants

## Material

The details of the Stroop experiment are report in @rosenbaum2017stand

## Procedure

In each posture condition, participants completed 72 Stroop trials, half congruent and half incongruent.

# Results

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# load libraries and data
library(data.table)
library(dplyr)
library(ggplot2)
all_data <- fread("data/stroop_stand.csv")

# pre-processing and conversion to long format
RTs <- c(as.numeric(unlist(all_data[,1])),
         as.numeric(unlist(all_data[,2])),
         as.numeric(unlist(all_data[,3])),
         as.numeric(unlist(all_data[,4]))
         )

Congruency <- rep(rep(c("Congruent","Incongruent"),each=50),2)
Posture <- rep(c("Stand","Sit"),each=100)
Subject <- rep(1:50,4)

stroop_df <- data.frame(Subject,Congruency,Posture,RTs)

##table
overall_means <- stroop_df %>%
                  group_by(Posture,Congruency) %>%
                  summarise(meanRT = round(mean(RTs), digits=0),
                            SEMRT = round((sd(RTs)/sqrt(length(RTs))), digits=0))

```

Mean reaction times for each subject in each condition to a 2 (Congruency: congruent vs. incongruent) x 2 (Posture: Standing vs. Sitting) were submitted to a repeated measures ANOVA. Mean RTs in each condition are displayed in Table 1, and in Figure 1. The full ANOVA table is reported in Table 2.

```{r, warning=FALSE, message=FALSE}
# Make and print the table

library(kableExtra)
library(tidyr)

# format from long to wide
wide_means <- overall_means %>% 
  gather(variable, value, -(Posture:Congruency)) %>%
  unite(temp, Congruency, variable) %>%
  spread(temp, value)

# rename columns
names(wide_means)[2:5] <- c("RT","SEM","RT","SEM")

# print table
kable(wide_means, "latex", caption = "Mean Reaction Times and Standard Errors of the Mean for Experiment 3", booktabs = T) %>%
  kable_styling() %>%
  add_header_above(c(" " = 1, "Congruent" = 2, "Incongruent" = 2))

```

```{r stroopfig, fig.cap="Mean reaction times wth standard error bars as a function of Posture and Congruency for Experiment 3"}
# make and print the figure

ggplot(overall_means, aes(x=Posture,
                          y=meanRT, 
                          group=Congruency,
                          fill=Congruency))+
  geom_bar(stat="identity",position="dodge")+
  theme_classic(base_size=12)+
  ylab("Mean Reaction Time (ms)")+
  geom_errorbar(aes(ymin=meanRT-SEMRT,
                    ymax=meanRT+SEMRT),
                position=position_dodge(width=0.9),
                width=.2,
                color="black")+
  coord_cartesian(ylim=c(750,1000))
```

```{r aovtable, results="asis"}
## ANOVA

# Make sure Subject is  a factor
stroop_df$Subject <-  as.factor(stroop_df$Subject)

aov_out <- aov(RTs~Posture*Congruency + Error(Subject/(Posture*Congruency)), stroop_df)

apa_aov <- apa_print(summary(aov_out))

apa_table(
  apa_aov$table
  , caption = "ANOVA table for Experiment 3"
)
```

Below are examples of writing the results using two methods. The first method is to report all of the values by hand. The second method is to embed the results of R variables into the reporting using papaja. Both results sections appear similar in the .pdf, so look at the .rmd file for this example to see how to use papaja.

## By hand reporting

There was a main effect of Congruency, F (1, 49) = 342.45, MSE = 1684.39, p < 0.001. Mean reaction times were slower for incongruent (922 ms) than congruent groups (815 ms).

There main effect of Posture was significant, F (1, 49) = 7.33, MSE = 4407.09, p =.009. Mean reaction times were slower for sitting (881 ms) than standing groups (855 ms).

The two-way interaction between Congruency and Posture was significant, F (1, 49) = 8.96, MSE = 731.82, p < 0.004. The Stroop effect was 23 ms smaller in the standing than sitting conditions.

## papaja reporting

There was a main effect of Congruency, `r apa_aov$full_result$Congruency`. Mean reaction times were slower for incongruent (922 ms) than congruent groups (815 ms).

There main effect of Posture was significant, `r apa_aov$full_result$Posture`. Mean reaction times were slower for sitting (881 ms) than standing groups (855 ms).

The two-way interaction between Congruency and Posture was significant, `r apa_aov$full_result$Posture_Congruency`. The Stroop effect was 23 ms smaller in the standing than sitting conditions.

# Discussion

The re-analysis successfully reproduced the analysis reported by @rosenbaum2017stand. In the following section, I show an example of completing a simulation based power analysis for this design.

## Simulation-based power analysis

The design was a 2x2 repeated measures design with 50 subjects. This design leads to three different effects, the main effect of congruency, the main effect of posture, and the posture by congruency interaction. A power analysis could be applied to any of these effects. I will show examples for the main effect of congruency first, and then for the interaction.

Several features of the design can go into the power analysis, the major ones being the number of subjects, the effect-size, and assumptions about the distributions underlying the sample data. Power is the probability that the design will reject the null-hypothesis, when the null-hypothesis is false and there is a true difference. Power depends on the size of the true difference. The same design will have greater power to detect a large versus small difference. Commonly, the assumed true difference is defined in terms of Cohen's D, a mean difference in terms of standard deviation units.

```{r}
null_distribution <- stroop_df %>%
                      group_by(Subject) %>%
                      summarise(meanRT= mean(RTs)) %>%
                      summarise(overall_meanRT = mean(meanRT),
                                overall_sd = sd(meanRT))
```

We will first estimate the overall mean reaction time, and the standard deviation of the mean reaction from the data. The overall mean was `r null_distribution$overall_meanRT`, and the overall standard deviation was `r null_distribution$overall_sd`.

To conduct the simulation we generate data for each subject using the rnorm function. Each subject contributed had two mean RTs in the congruent condition (sitting and standing), and two mean RTs in the incongruent condition (sitting and standing). There were 50 subjects, for the congruent condition we sample 2 scores for each subject from the above normal distribution (100 total scores), and 2 scores for each subject from the above normal distribution for the incongruent condition (100 total scores). To model the Stroop effect, we systemically increase the mean in the incongruent condition by a proportion of the standard deviation. We use effect-sizes of .05, .1, .2, .4, .5, and .8; which range from small to large. For each effect-size, we run 100 simulated experiments, and save p-value for the main effect of congruency for each simulated experiment. Then, for each effect-size, we find the proportion of experiments that resulted in p<.05. The proportion of experiments that reject the null is the power of the design to detect an effect of each size. The simulation below finds that this design had power of .8, to detect an effect of d=.4. It had power of .99 to detect effects of d=.8 or larger. The full power-curve for this design is displayed in Figure 2.

```{r}

s_mean <- null_distribution$overall_meanRT
s_sd <- null_distribution$overall_sd
effect_sizes <-c(.05,.1,.2,.4,.5,.8)

save_es <- c()
save_p <- c()
for(es in effect_sizes){
  for(i in 1:100){
  sim_congruent <- rnorm(50*2,s_mean,s_sd)
  sim_incongruent <- rnorm(50*2,(s_mean+(es*s_sd)),s_sd)
  subjects <- as.factor(rep(rep(1:50,each=2),2))
  congruency <- as.factor(rep(c("C","I"),each=100))

  sim_df <- data.frame(subjects,
                       RTs = c(sim_congruent,sim_incongruent),
                       congruency)

  sim_aov <- summary(aov(RTs~congruency+Error(subjects/congruency),sim_df))
  get_pvalue <- sim_aov$`Error: subjects:congruency`[[1]]$`Pr(>F)`[1]
  save_es <- c(save_es,es)
  save_p <- c(save_p,get_pvalue)
   }
}

power_sim <- data.frame(save_es, save_p)

plot_df <- power_sim %>%
  group_by(save_es)%>%
  summarise(power = length(save_p[save_p<.05])/length(save_p))
```

```{r powerfig, fig.cap="Simulation-based power curve for this design"}
ggplot(plot_df, aes(x=save_es, y=power))+
  geom_line()+
  geom_point()+
  xlab("Effect-size")
```

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup

### 2019-11-08 inst/vertical/docs/manuscript/example2/APAreport.tex
add project template

\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Reproducing the analysis of Rosenbaum, Mama, \& Algom (2017)},
            pdfauthor={Matthew Crump},
            pdfkeywords={Stroop, Reproducibilty},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Reproducing the analysis of Rosenbaum, Mama, \& Algom (2017)}
    \author{Matthew Crump\textsuperscript{1}}
    \date{}

\shorttitle{Stroop Analysis}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Brooklyn College of the City University of New York}
\keywords{Stroop, Reproducibilty\newline\indent Word count: X}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother

\authornote{Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

Correspondence concerning this article should be addressed to Matthew Crump, 2900 Bedford Ave. E-mail: \href{mailto:mcrump@brooklyn.cuny.edu}{\nolinkurl{mcrump@brooklyn.cuny.edu}}}

\abstract{
A reproduction of the analysis for Experiment 3 from Rosenbaum, Mama, and Algom (2017).

}

\begin{document}
\maketitle

This is an example of creating an APA manuscript using papaja, and also an example for part 2 of the midterm assignment. Another sentence.

This report re-produces the analysis of Experiment 3 reported in Rosenbaum, Mama, and Algom (2017). The data were downloaded from \url{https://osf.io/b7x8q/}

Rosenbaum et al. (2017) had participants perform a Stroop task in one of two posture conditions. Participants either sat and performed the Stroop task, or stood and performed the Stroop task. The question was whether the size of the Stroop effect would change as a function of posture. The Stroop effect is measured as a difference between reaction times on congruent vs.~incongruent trials.The experiment involved a 2 (Posture: sitting vs standing) x 2 (congruency: congruent vs.~incongruent) repeated measures design.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

There were 50 participants

\hypertarget{material}{%
\subsection{Material}\label{material}}

The details of the Stroop experiment are report in Rosenbaum et al. (2017)

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

In each posture condition, participants completed 72 Stroop trials, half congruent and half incongruent.

\hypertarget{results}{%
\section{Results}\label{results}}

Mean reaction times for each subject in each condition to a 2 (Congruency: congruent vs.~incongruent) x 2 (Posture: Standing vs.~Sitting) were submitted to a repeated measures ANOVA. Mean RTs in each condition are displayed in Table 1, and in Figure 1. The full ANOVA table is reported in Table 2.

\begin{table}

\caption{\label{tab:unnamed-chunk-2}Mean Reaction Times and Standard Errors of the Mean for Experiment 3}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Congruent} & \multicolumn{2}{c}{Incongruent} \\
\cmidrule(l{2pt}r{2pt}){2-3} \cmidrule(l{2pt}r{2pt}){4-5}
Posture & RT & SEM & RT & SEM\\
\midrule
Sit & 822 & 17 & 941 & 18\\
Stand & 808 & 15 & 904 & 15\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{APAreport_files/figure-latex/stroopfig-1.pdf}
\caption{\label{fig:stroopfig}Mean reaction times wth standard error bars as a function of Posture and Congruency for Experiment 3}
\end{figure}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:aovtable}ANOVA table for Experiment 3}
\begin{tabular}{lllllll}
\toprule
Effect & \multicolumn{1}{c}{$F$} & \multicolumn{1}{c}{$\mathit{df}_1$} & \multicolumn{1}{c}{$\mathit{df}_2$} & \multicolumn{1}{c}{$\mathit{MSE}$} & \multicolumn{1}{c}{$p$} & \multicolumn{1}{c}{$\hat{\eta}^2_G$}\\
\midrule
Posture & 7.33 & 1 & 49 & 4,407.09 & .009 & .012\\
Congruency & 342.45 & 1 & 49 & 1,684.39 & < .001 & .182\\
Posture $\times$ Congruency & 8.96 & 1 & 49 & 731.82 & .004 & .003\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\end{table}

Below are examples of writing the results using two methods. The first method is to report all of the values by hand. The second method is to embed the results of R variables into the reporting using papaja. Both results sections appear similar in the .pdf, so look at the .rmd file for this example to see how to use papaja.

\hypertarget{by-hand-reporting}{%
\subsection{By hand reporting}\label{by-hand-reporting}}

There was a main effect of Congruency, F (1, 49) = 342.45, MSE = 1684.39, p \textless{} 0.001. Mean reaction times were slower for incongruent (922 ms) than congruent groups (815 ms).

There main effect of Posture was significant, F (1, 49) = 7.33, MSE = 4407.09, p =.009. Mean reaction times were slower for sitting (881 ms) than standing groups (855 ms).

The two-way interaction between Congruency and Posture was significant, F (1, 49) = 8.96, MSE = 731.82, p \textless{} 0.004. The Stroop effect was 23 ms smaller in the standing than sitting conditions.

\hypertarget{papaja-reporting}{%
\subsection{papaja reporting}\label{papaja-reporting}}

There was a main effect of Congruency, \(F(1, 49) = 342.45\), \(\mathit{MSE} = 1,684.39\), \(p < .001\), \(\hat{\eta}^2_G = .182\). Mean reaction times were slower for incongruent (922 ms) than congruent groups (815 ms).

There main effect of Posture was significant, \(F(1, 49) = 7.33\), \(\mathit{MSE} = 4,407.09\), \(p = .009\), \(\hat{\eta}^2_G = .012\). Mean reaction times were slower for sitting (881 ms) than standing groups (855 ms).

The two-way interaction between Congruency and Posture was significant, \(F(1, 49) = 8.96\), \(\mathit{MSE} = 731.82\), \(p = .004\), \(\hat{\eta}^2_G = .003\). The Stroop effect was 23 ms smaller in the standing than sitting conditions.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The re-analysis successfully reproduced the analysis reported by Rosenbaum et al. (2017). In the following section, I show an example of completing a simulation based power analysis for this design.

\hypertarget{simulation-based-power-analysis}{%
\subsection{Simulation-based power analysis}\label{simulation-based-power-analysis}}

The design was a 2x2 repeated measures design with 50 subjects. This design leads to three different effects, the main effect of congruency, the main effect of posture, and the posture by congruency interaction. A power analysis could be applied to any of these effects. I will show examples for the main effect of congruency first, and then for the interaction.

Several features of the design can go into the power analysis, the major ones being the number of subjects, the effect-size, and assumptions about the distributions underlying the sample data. Power is the probability that the design will reject the null-hypothesis, when the null-hypothesis is false and there is a true difference. Power depends on the size of the true difference. The same design will have greater power to detect a large versus small difference. Commonly, the assumed true difference is defined in terms of Cohen's D, a mean difference in terms of standard deviation units.

We will first estimate the overall mean reaction time, and the standard deviation of the mean reaction from the data. The overall mean was 868.65, and the overall standard deviation was 107.16.

To conduct the simulation we generate data for each subject using the rnorm function. Each subject contributed had two mean RTs in the congruent condition (sitting and standing), and two mean RTs in the incongruent condition (sitting and standing). There were 50 subjects, for the congruent condition we sample 2 scores for each subject from the above normal distribution (100 total scores), and 2 scores for each subject from the above normal distribution for the incongruent condition (100 total scores). To model the Stroop effect, we systemically increase the mean in the incongruent condition by a proportion of the standard deviation. We use effect-sizes of .05, .1, .2, .4, .5, and .8; which range from small to large. For each effect-size, we run 100 simulated experiments, and save p-value for the main effect of congruency for each simulated experiment. Then, for each effect-size, we find the proportion of experiments that resulted in p\textless{}.05. The proportion of experiments that reject the null is the power of the design to detect an effect of each size. The simulation below finds that this design had power of .8, to detect an effect of d=.4. It had power of .99 to detect effects of d=.8 or larger. The full power-curve for this design is displayed in Figure 2.

\begin{figure}
\centering
\includegraphics{APAreport_files/figure-latex/powerfig-1.pdf}
\caption{\label{fig:powerfig}Simulation-based power curve for this design}
\end{figure}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-rosenbaum2017stand}{}%
Rosenbaum, D., Mama, Y., \& Algom, D. (2017). Stand by your stroop: Standing up enhances selective attention and cognitive control. \emph{Psychological Science}, \emph{28}(12), 1864--1867.

\endgroup

\end{document}

### 2019-11-08 inst/vertical/slides/r_markdown.Rmd
add project template

---
title: "Untitled"
author: "Matt Crump"
date: "10/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2019-11-08 R/vertical_project.R

@@ -1,0 +1,6 @@
+vertical_project <- function(path, ...) {
+  dir.create(path, recursive = TRUE, showWarnings = FALSE)
+  file.copy(list.files(system.file("vertical", package="vertical"),
+                       full.names = TRUE),
+            path, recursive=TRUE, overwrite=TRUE)
+}

### 2019-11-08 inst/vertical/R/vertical_project.R

@@ -1,0 +1,6 @@
+vertical_project <- function(path, ...) {
+  dir.create(path, recursive = TRUE, showWarnings = FALSE)
+  file.copy(list.files(system.file("vertical", package="ptexamples"),
+                       full.names = TRUE),
+            path, recursive=TRUE, overwrite=TRUE)
+}

### 2019-11-08 inst/vertical/experiments/Experiment 1/run.R

@@ -1,0 +1,17 @@
+# Run all of this to run the experiment in a browser
+# critical randomization is done by R right now
+# so this should be run to initiate the experiment for each new subject
+# After the experiment is complete, the data should be stored in the data folder
+# when finished, "stop" the server that will be be started by this code
+
+library(xprmntr)
+
+rmarkdown::render("experiments/Experiment 1/experiment/index.Rmd", "html_document")
+
+run_locally(path="experiments/Experiment 1/",
+            show_in = "browser",
+            xprmntr_host = "127.0.0.1",
+            xprmntr_port = 8000)
+
+
+

### 2019-11-08 Package Helper.R
add logo

@@ -17,3 +17,4 @@ pkgdown::template_navbar() # generate yml for controlling navbar
 
 render_folder_to_docs("slides")
 
+use_logo("xtra/logo.png")

### 2019-11-08 inst/vertical/slides/slidy_example.Rmd
new template

---
title: "Example slide presentation using slidy"
subtitle: "A place for a subtitle"
author: "Matthew Crump"
date: "10/31/2019"
output:
  slidy_presentation:
    self_contained: true
    theme: yeti
    incremental: true
    css: custom_css/style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Example slidy presentation {.middle .center}

Check out the documentation on slidy for more info <https://bookdown.org/yihui/rmarkdown/slidy-presentation.html>

- Note that this slide deck has some custom CSS controlling the styling and look and feel. This can be customized further by editing the style.css file.

## A slide with a header {.bigger}

## A slide with a header {.bigger}

And some text in it.

## A slide with a header and bullet points {.bigger}

- bullet 1
- bullet 2
- bullet 3

## Something in the middle{.middle .center}

## Slide with a picture

```{r, echo = FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("imgs/logo.png")
```

## The end {.middle .center}

- don't forget there are other R markdown options for slide decks
- eg., xaringan, and ioslides, and probably a few others

### 2019-11-08 R/render_folder_to_docs.R

@@ -1,16 +1,19 @@
+#' Render rmd ina folder to docs
+#'
+#' @param folder_name string, name of folder , eg "slides"
+#' @param dest_folder string, name of destination folder, default = "docs
+#'
+#' @return rendered rmd files in destination folder
+#' @export
+#'
 render_folder_to_docs <- function(folder_name,
-                                  dest_folder = "docs",
-                                  assets_folder){
+                                  dest_folder = "docs"){
 
   file_names <- list.files(folder_name, pattern = "\\.Rmd$")
 
   for(i in file_names){
     rmarkdown::render(paste(folder_name,file_names[1],sep="/"),
                       output_dir = paste(".",dest_folder,folder_name,sep="/"))
-  }
-
-  if (missing(assets_folder) == FALSE){
-
   }
 }
 

### 2019-11-08 inst/vertical/Package Helper.R

@@ -1,19 +1,0 @@
-# Package helper
-library(usethis) # helpful automation functions for common tasks in package development
-library(pkgdown) # compile your R package to a website
-
-pkgdown::build_site()
-
-# generate description file with your info (need to set your info in advanve)
-usethis::use_description()
-
-# generate readme file, the contents control front page of pkgdown website
-usethis::use_readme_md()
-
-# create a vignette (.Rmd) document that gets compiled and displayed on pkdown website
-usethis::use_vignette("get_started", title="Getting Started")
-
-pkgdown::template_navbar() # generate yml for controlling navbar
-
-render_folder_to_docs("slides")
-

### 2019-11-08 inst/vertical/R/render_folder_to_docs.R

@@ -1,17 +1,0 @@
-render_folder_to_docs <- function(folder_name,
-                                  dest_folder = "docs",
-                                  assets_folder){
-
-  file_names <- list.files(folder_name, pattern = "\\.Rmd$")
-
-  for(i in file_names){
-    rmarkdown::render(paste(folder_name,file_names[1],sep="/"),
-                      output_dir = paste(".",dest_folder,folder_name,sep="/"))
-  }
-
-  if (missing(assets_folder) == FALSE){
-
-  }
-}
-
-

### 2019-11-08 inst/vertical/R/vertical_project.R

@@ -1,6 +1,0 @@
-vertical_project <- function(path, ...) {
-  dir.create(path, recursive = TRUE, showWarnings = FALSE)
-  file.copy(list.files(system.file("vertical", package="ptexamples"),
-                       full.names = TRUE),
-            path, recursive=TRUE, overwrite=TRUE)
-}

### 2019-11-08 inst/vertical/Vertical Helper.R

@@ -1,0 +1,25 @@
+# for more help with vertical go to https://crumplab.github.io/vertical/
+
+# Vertical helper
+library(usethis) # helpful automation functions for common tasks in package development
+library(pkgdown) # helpful automation functions for common tasks in package development
+library(vertical) # vertical helper functions
+
+# compile the website using pkgdown
+pkgdown::build_site()
+
+# clean the site (deletes everythng in docs folder)
+pkgdown::clean_site()
+
+# render slides to docs (for display on website)
+vertical::render_folder_to_docs("slides")
+
+# generate description file with your info (need to set your info in advance)
+# see https://usethis.r-lib.org/reference/use_description.html
+usethis::use_description()
+
+# create a vignette (.Rmd) document that gets compiled and displayed on website under supplementary
+usethis::use_vignette("Supplemental_2", title="Supplemental 2")
+
+
+

### 2019-11-08 inst/vertical/docs/experiments/Experiment 1/experiment/index.Rmd
delete template structure

---
title: "Stroop Demo"
author: "Matthew Crump"
output:
  html_document:
    mathjax: null
    highlight: null
    theme: null
    css: jspsych-6/css/jspsych.css
    includes:
      in_header: plugins.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

```{r}
# load jspsych and plugin scripts
library(htmltools)
tagList(
tags$script(src='jspsych-6/jspsych.js'),
tags$script(src='jspsych-6/plugins/jspsych-html-keyboard-response.js'),
tags$script(src='xprmntr/xprmntr.js')
)
```

```{r, results="asis"}
# Define stimuli in R, output as script 
library(xprmntr)
library(dplyr)

# create dataframe to define stimuli
stroop_stim <- data.frame( stimulus = length(16),
                           word = rep(c("red","green","blue","yellow"), each=4),
                           color = rep(c("red","green","blue","yellow"), 4),
                           response = rep(c("r","g","b","y"), 4),
                           stim_type = length(16),
                           id = "stroop_stim",
                           fontsize = "60pt") %>%
  mutate(stim_type = as.numeric(word==color)) %>%
  mutate(stim_type = recode(stim_type, `1` = "congruent", `0` = "incongruent"))

# write html definitions to the stimulus column
# note this could be added as a pipe to the above, setting df=.
stroop_stim$stimulus <- html_stimulus(df = stroop_stim, 
                                html_content = "word",
                                html_element = "p",
                                column_names = c("color","fontsize"),
                                css = c("color", "font-size"),
                                id = "id")

# create json object from dataframe
stimulus_json <- stimulus_df_to_json(df = stroop_stim,
                                     stimulus = "stimulus",
                                     data = c("word","color","response","stim_type"))

# write json object to script
write_to_script(stimulus_json,"test_stimuli")

```

```{js}
/* experiment parameters */
var reps_per_trial_type = 1;

/*set up welcome block*/
var welcome = {
  type: "html-keyboard-response",
  stimulus: "Welcome to the experiment. Press any key to begin."
};

/*set up instructions block*/
var instructions = {
  type: "html-keyboard-response",
  stimulus: "<p>In this task, you will see word in a color</p>"+
    "<p style='color:red;font-size:60pt;'>BLUE</p>"+
    "<p>Press r, g, b, or y to identify the color, not the word</p>"+
    "<p>E.g., press r for red in this example</p>"+
    "<p>Press any key to begin.</p>",
  post_trial_gap: 1000
};

/* defining test timeline */
var test = {
  timeline: [{
    type: 'html-keyboard-response',
    choices: ["r","g","b","y"],
    trial_duration: 2000,
    response_ends_trial: true,
    stimulus: jsPsych.timelineVariable('stimulus'),
    data: jsPsych.timelineVariable('data'),
    on_finish: function(data){
      var correct = false;
      if(data.response == 'r' &&  data.key_press == 82 && data.rt > -1){
        correct = true;
      } else if(data.response == 'g' &&  data.key_press == 71 && data.rt > -1){
        correct = true;
      } else if(data.response == 'b' &&  data.key_press == 66 && data.rt > -1){
        correct = true;
      } else if(data.response == 'y' &&  data.key_press == 89 && data.rt > -1){
        correct = true;
      }
      data.correct = correct;
    },
    post_trial_gap: function() {
        return Math.floor(Math.random() * 1500) + 500;
    }
  }],
  timeline_variables: test_stimuli,
  sample: {type: 'fixed-repetitions', size: reps_per_trial_type}
};

/*defining debriefing block*/
var debrief = {
  type: "html-keyboard-response",
  stimulus: function() {
    var total_trials = jsPsych.data.get().filter({trial_type: 'html-keyboard-response'}).count();
    var accuracy = Math.round(jsPsych.data.get().filter({correct: true}).count() / (total_trials-2) * 100);
    var congruent_rt = Math.round(jsPsych.data.get().filter({correct: true, stim_type: 'congruent'}).select('rt').mean());
    var incongruent_rt = Math.round(jsPsych.data.get().filter({correct: true, stim_type: 'incongruent'}).select('rt').mean());
    return "<p>You responded correctly on <strong>"+accuracy+"%</strong> of the trials.</p> " +
    "<p>Your average response time for congruent trials was <strong>" + congruent_rt + "ms</strong>.</p>"+
    "<p>Your average response time for incongruent trials was <strong>" + incongruent_rt + "ms</strong>.</p>"+
    "<p>Press any key to complete the experiment. Thank you!</p>";
  }
};

/*set up experiment structure*/
var timeline = [];
timeline.push(welcome);
timeline.push(instructions);
timeline.push(test);
timeline.push(debrief);

/*start experiment*/
jsPsych.init({
    timeline: timeline,
    on_finish: function() {
        xprmntr.save_locally();
    }
});
```

### 2019-11-08 R/render_folder_to_docs.R

@@ -1,4 +1,4 @@
-#' Render rmd ina folder to docs
+#' Render rmd in a folder to docs
 #'
 #' @param folder_name string, name of folder , eg "slides"
 #' @param dest_folder string, name of destination folder, default = "docs

### 2019-11-08 inst/vertical/Vertical Helper.R

@@ -21,5 +21,8 @@ usethis::use_description()
 # create a vignette (.Rmd) document that gets compiled and displayed on website under supplementary
 usethis::use_vignette("Supplemental_2", title="Supplemental 2")
 
+# initialize your repository as a git repository
+usethis::use_git("initial")
+
 
 

### 2019-11-08 Package Helper.R
Update Package Helper.R

@@ -18,3 +18,5 @@ pkgdown::template_navbar() # generate yml for controlling navbar
 render_folder_to_docs("slides")
 
 use_logo("xtra/logo.png")
+
+usethis::use_news_md()

### 2019-11-11 R/vertical_project.R
modularize papaja

@@ -1,6 +1,11 @@
 vertical_project <- function(path, ...) {
   dir.create(path, recursive = TRUE, showWarnings = FALSE)
+  #copy all vertical template files
   file.copy(list.files(system.file("vertical", package="vertical"),
                        full.names = TRUE),
             path, recursive=TRUE, overwrite=TRUE)
+
+  # copy papaja template to manuscript folder
+  try(rmarkdown::draft(paste(path,"/manuscript/manuscript.Rmd",sep=""),
+                   template="apa6", package="papaja", edit=FALSE, create_dir=FALSE))
 }

### 2019-11-11 R/vertical_project.R
get most recent jspsych

@@ -8,4 +8,14 @@ vertical_project <- function(path, ...) {
   # copy papaja template to manuscript folder
   try(rmarkdown::draft(paste(path,"/manuscript/manuscript.Rmd",sep=""),
                    template="apa6", package="papaja", edit=FALSE, create_dir=FALSE))
+
+  # copy latest jspsych release as .zip to experiments folder
+  end_point <- "repos/jspsych/jsPsych/releases/latest"
+  response <- devtools:::github_GET(end_point)
+  latest_tag_name <- response$assets[[1]]$browser_download_url
+  download.file(url = latest_tag_name,file.path("experiments", basename(latest_tag_name)))
+
 }
+
+
+#response$assets[[1]]$browser_download_url

### 2019-11-11 inst/vertical/experiments/Experiment 1/experiment/jspsych-6/plugins/jspsych-survey-multi-choice.js

@@ -1,198 +1,0 @@
-/**
- * jspsych-survey-multi-choice
- * a jspsych plugin for multiple choice survey questions
- *
- * Shane Martin
- *
- * documentation: docs.jspsych.org
- *
- */
-
-
-jsPsych.plugins['survey-multi-choice'] = (function() {
-  var plugin = {};
-
-  plugin.info = {
-    name: 'survey-multi-choice',
-    description: '',
-    parameters: {
-      questions: {
-        type: jsPsych.plugins.parameterType.COMPLEX,
-        array: true,
-        pretty_name: 'Questions',
-        nested: {
-          prompt: {
-            type: jsPsych.plugins.parameterType.STRING,
-            pretty_name: 'Prompt',
-            default: undefined,
-            description: 'The strings that will be associated with a group of options.'
-          },
-          options: {
-            type: jsPsych.plugins.parameterType.STRING,
-            pretty_name: 'Options',
-            array: true,
-            default: undefined,
-            description: 'Displays options for an individual question.'
-          },
-          required: {
-            type: jsPsych.plugins.parameterType.BOOL,
-            pretty_name: 'Required',
-            default: false,
-            description: 'Subject will be required to pick an option for each question.'
-          },
-          horizontal: {
-            type: jsPsych.plugins.parameterType.BOOL,
-            pretty_name: 'Horizontal',
-            default: false,
-            description: 'If true, then questions are centered and options are displayed horizontally.'
-          },
-          name: {
-            type: jsPsych.plugins.parameterType.STRING,
-            pretty_name: 'Question Name',
-            default: '',
-            description: 'Controls the name of data values associated with this question'
-          }
-        }
-      },
-      randomize_question_order: {
-        type: jsPsych.plugins.parameterType.BOOL,
-        pretty_name: 'Randomize Question Order',
-        default: false,
-        description: 'If true, the order of the questions will be randomized'
-      },
-      preamble: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Preamble',
-        default: null,
-        description: 'HTML formatted string to display at the top of the page above all the questions.'
-      },
-      button_label: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Button label',
-        default:  'Continue',
-        description: 'Label of the button.'
-      }
-    }
-  }
[…]

### 2019-11-11 inst/vertical/experiments/Experiment 1/experiment/jspsych-6/plugins/jspsych-video-button-response.js

@@ -1,279 +1,0 @@
-/**
- * jspsych-video-button-response
- * Josh de Leeuw
- *
- * plugin for playing a video file and getting a button response
- *
- * documentation: docs.jspsych.org
- *
- **/
-
-jsPsych.plugins["video-button-response"] = (function() {
-
-  var plugin = {};
-
-  jsPsych.pluginAPI.registerPreload('video-button-response', 'stimulus', 'video');
-
-  plugin.info = {
-    name: 'video-button-response',
-    description: '',
-    parameters: {
-      sources: {
-        type: jsPsych.plugins.parameterType.VIDEO,
-        pretty_name: 'Video',
-        default: undefined,
-        description: 'The video file to play.'
-      },
-      choices: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Choices',
-        default: undefined,
-        array: true,
-        description: 'The labels for the buttons.'
-      },
-      button_html: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Button HTML',
-        default: '<button class="jspsych-btn">%choice%</button>',
-        array: true,
-        description: 'The html of the button. Can create own style.'
-      },
-      prompt: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Prompt',
-        default: null,
-        description: 'Any content here will be displayed below the buttons.'
-      },
-      width: {
-        type: jsPsych.plugins.parameterType.INT,
-        pretty_name: 'Width',
-        default: '',
-        description: 'The width of the video in pixels.'
-      },
-      height: {
-        type: jsPsych.plugins.parameterType.INT,
-        pretty_name: 'Height',
-        default: '',
-        description: 'The height of the video display in pixels.'
-      },
-      autoplay: {
-        type: jsPsych.plugins.parameterType.BOOL,
-        pretty_name: 'Autoplay',
-        default: true,
-        description: 'If true, the video will begin playing as soon as it has loaded.'
-      },
-      controls: {
-        type: jsPsych.plugins.parameterType.BOOL,
-        pretty_name: 'Controls',
-        default: false,
-        description: 'If true, the subject will be able to pause the video or move the playback to any point in the video.'
-      },
-      start: {
-        type: jsPsych.plugins.parameterType.FLOAT,
-        pretty_name: 'Start',
-        default: null,
-        description: 'Time to start the clip.'
-      },
-      stop: {
[…]

### 2019-11-11 inst/vertical/experiments/Experiment 1/experiment/jspsych-6/plugins/jspsych-video-keyboard-response.js

@@ -1,236 +1,0 @@
-/**
- * jspsych-video-keyboard-response
- * Josh de Leeuw
- *
- * plugin for playing a video file and getting a keyboard response
- *
- * documentation: docs.jspsych.org
- *
- **/
-
-jsPsych.plugins["video-keyboard-response"] = (function() {
-
-  var plugin = {};
-
-  jsPsych.pluginAPI.registerPreload('video-keyboard-response', 'stimulus', 'video');
-
-  plugin.info = {
-    name: 'video-keyboard-response',
-    description: '',
-    parameters: {
-      sources: {
-        type: jsPsych.plugins.parameterType.VIDEO,
-        pretty_name: 'Video',
-        default: undefined,
-        description: 'The video file to play.'
-      },
-      choices: {
-        type: jsPsych.plugins.parameterType.KEYCODE,
-        pretty_name: 'Choices',
-        array: true,
-        default: jsPsych.ALL_KEYS,
-        description: 'The keys the subject is allowed to press to respond to the stimulus.'
-      },
-      prompt: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Prompt',
-        default: null,
-        description: 'Any content here will be displayed below the stimulus.'
-      },
-      width: {
-        type: jsPsych.plugins.parameterType.INT,
-        pretty_name: 'Width',
-        default: '',
-        description: 'The width of the video in pixels.'
-      },
-      height: {
-        type: jsPsych.plugins.parameterType.INT,
-        pretty_name: 'Height',
-        default: '',
-        description: 'The height of the video display in pixels.'
-      },
-      autoplay: {
-        type: jsPsych.plugins.parameterType.BOOL,
-        pretty_name: 'Autoplay',
-        default: true,
-        description: 'If true, the video will begin playing as soon as it has loaded.'
-      },
-      controls: {
-        type: jsPsych.plugins.parameterType.BOOL,
-        pretty_name: 'Controls',
-        default: false,
-        description: 'If true, the subject will be able to pause the video or move the playback to any point in the video.'
-      },
-      start: {
-        type: jsPsych.plugins.parameterType.FLOAT,
-        pretty_name: 'Start',
-        default: null,
-        description: 'Time to start the clip.'
-      },
-      stop: {
-        type: jsPsych.plugins.parameterType.FLOAT,
-        pretty_name: 'Stop',
-        default: null,
-        description: 'Time to stop the clip.'
-      },
-      rate: {
[…]

### 2019-11-11 inst/vertical/experiments/Experiment 1/experiment/jspsych-6/plugins/jspsych-video-slider-response.js

@@ -1,291 +1,0 @@
-/**
- * jspsych-video-slider-response
- * Josh de Leeuw
- *
- * plugin for playing a video file and getting a slider response
- *
- * documentation: docs.jspsych.org
- *
- **/
-
-jsPsych.plugins["video-slider-response"] = (function() {
-
-  var plugin = {};
-
-  jsPsych.pluginAPI.registerPreload('video-slider-response', 'stimulus', 'video');
-
-  plugin.info = {
-    name: 'video-slider-response',
-    description: '',
-    parameters: {
-      sources: {
-        type: jsPsych.plugins.parameterType.VIDEO,
-        pretty_name: 'Video',
-        default: undefined,
-        description: 'The video file to play.'
-      },
-      prompt: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Prompt',
-        default: null,
-        description: 'Any content here will be displayed below the stimulus.'
-      },
-      width: {
-        type: jsPsych.plugins.parameterType.INT,
-        pretty_name: 'Width',
-        default: '',
-        description: 'The width of the video in pixels.'
-      },
-      height: {
-        type: jsPsych.plugins.parameterType.INT,
-        pretty_name: 'Height',
-        default: '',
-        description: 'The height of the video display in pixels.'
-      },
-      autoplay: {
-        type: jsPsych.plugins.parameterType.BOOL,
-        pretty_name: 'Autoplay',
-        default: true,
-        description: 'If true, the video will begin playing as soon as it has loaded.'
-      },
-      controls: {
-        type: jsPsych.plugins.parameterType.BOOL,
-        pretty_name: 'Controls',
-        default: false,
-        description: 'If true, the subject will be able to pause the video or move the playback to any point in the video.'
-      },
-      start: {
-        type: jsPsych.plugins.parameterType.FLOAT,
-        pretty_name: 'Start',
-        default: null,
-        description: 'Time to start the clip.'
-      },
-      stop: {
-        type: jsPsych.plugins.parameterType.FLOAT,
-        pretty_name: 'Stop',
-        default: null,
-        description: 'Time to stop the clip.'
-      },
-      rate: {
-        type: jsPsych.plugins.parameterType.FLOAT,
-        pretty_name: 'Rate',
-        default: 1,
-        description: 'The playback rate of the video. 1 is normal, <1 is slower, >1 is faster.'
-      },
-      min: {
-        type: jsPsych.plugins.parameterType.INT,
[…]

### 2019-11-11 R/render_folder_to_docs.R
Include dependencies dynamically instead of hard-coding

@@ -1,20 +1,0 @@
-#' Render rmd in a folder to docs
-#'
-#' @param folder_name string, name of folder , eg "slides"
-#' @param dest_folder string, name of destination folder, default = "docs
-#'
-#' @return rendered rmd files in destination folder
-#' @export
-#'
-render_folder_to_docs <- function(folder_name,
-                                  dest_folder = "docs"){
-
-  file_names <- list.files(folder_name, pattern = "\\.Rmd$")
-
-  for(i in file_names){
-    rmarkdown::render(paste(folder_name,file_names[1],sep="/"),
-                      output_dir = paste(".",dest_folder,folder_name,sep="/"))
-  }
-}
-
-

### 2019-11-11 R/vertical_project.R

@@ -1,6 +1,116 @@
+#' Initialize a **vertical** project
+#'
+#' This function is called when a **vertical** R project is created.
+#'
+#' @param path Where to create the project. This must be a valid R package name.
+#' @param ...  Not used.
+#'
+#' @return
+#' @export
 vertical_project <- function(path, ...) {
+
+  dots <- list(...)
   dir.create(path, recursive = TRUE, showWarnings = FALSE)
-  file.copy(list.files(system.file("vertical", package="vertical"),
-                       full.names = TRUE),
-            path, recursive=TRUE, overwrite=TRUE)
+
+  # This is a package
+  usethis::create_package(path, open = FALSE)
+  setwd(paste0(getwd(), "/", path)) # [TODO] improve this hack
+
+  # pkgdown template
+  vertical_pkgdown <- system.file("vertical/_pkgdown.yml", package = "vertical")
+  base::file.copy(vertical_pkgdown, "_pkgdown.yml")
+
+  # papaja manuscript
+  if (dots$init_ms) {
+    usethis::use_directory("manuscript", ignore = TRUE)
+    rmarkdown::draft(
+      file = "manuscript/manuscript.Rmd",
+      template = "apa6",
+      package = "papaja",
+      edit = FALSE,
+      create_dir = FALSE
+    )
+  }
+
+  # supplementary analysis file
+  if (dots$init_som) {
+    usethis::use_article("Supplemental_1", title = "Supplementary analyses")
+  }
+
+  # Slides
+  if (dots$init_slides) {
+    usethis::use_directory("slides", ignore = TRUE)
+    rmarkdown::draft(
+      file = "slides/slides.Rmd",
+      template = "slidy",
+      package = "vertical",
+      edit = FALSE,
+      create_dir = FALSE
+    )
+  }
+
+  # Poster
+  if (dots$init_poster) {
+    usethis::use_directory("posters", ignore = TRUE)
+    rmarkdown::draft(
+      file = "posters/poster.Rmd",
+      template = "posterdown_html",
+      package = "posterdown",
+      edit = FALSE,
+      create_dir = FALSE
+    )
+  }
+
+  # Data
+  usethis::use_data_raw()
+
+  # Experiments
+  if (dots$init_exp) {
+    usethis::use_directory("experiments", ignore = TRUE)
+    usethis::use_directory("experiments/experiment-1")
[…]

### 2019-11-11 inst/vertical/experiments/Experiment 1/experiment/jspsych-6/jspsych.js

@@ -1,2721 +1,0 @@
-/**
- * jspsych.js
- * Josh de Leeuw
- *
- * documentation: docs.jspsych.org
- *
- **/
-window.jsPsych = (function() {
-
-  var core = {};
-
-  //
-  // private variables
-  //
-
-  // options
-  var opts = {};
-  // experiment timeline
-  var timeline;
-  // flow control
-  var global_trial_index = 0;
-  var current_trial = {};
-  var current_trial_finished = false;
-  // target DOM element
-  var DOM_container;
-  var DOM_target;
-  // time that the experiment began
-  var exp_start_time;
-  // is the experiment paused?
-  var paused = false;
-  var waiting = false;
-  // done loading?
-  var loaded = false;
-  var loadfail = false;
-
-  // storing a single webaudio context to prevent problems with multiple inits
-  // of jsPsych
-  core.webaudio_context = null;
-  // temporary patch for Safari
-  if (typeof window !== 'undefined' && window.hasOwnProperty('webkitAudioContext') && !window.hasOwnProperty('AudioContext')) {
-    window.AudioContext = webkitAudioContext;
-  }
-  // end patch
-  core.webaudio_context = (typeof window !== 'undefined' && typeof window.AudioContext !== 'undefined') ? new AudioContext() : null;
-
-  // enumerated variables for special parameter types
-  core.ALL_KEYS = 'allkeys';
-  core.NO_KEYS = 'none';
-
-  //
-  // public methods
-  //
-
-  core.init = function(options) {
-
-    if(typeof options.timeline === 'undefined'){
-      console.error('No timeline declared in jsPsych.init. Cannot start experiment.')
-    }
-
-    // reset variables
-    timeline = null;
-    global_trial_index = 0;
-    current_trial = {};
-    current_trial_finished = false;
-    paused = false;
-    waiting = false;
-    loaded = false;
-    loadfail = false;
-    jsPsych.data.reset();
-
-    var defaults = {
-      'display_element': undefined,
-      'on_finish': function(data) {
-        return undefined;
-      },
-      'on_trial_start': function(trial) {
-        return undefined;
-      },
-      'on_trial_finish': function() {
-        return undefined;
-      },
-      'on_data_update': function(data) {
-        return undefined;
-      },
-      'on_interaction_data_update': function(data){
-        return undefined;
-      },
-      'on_close': function(){
-        return undefined;
-      },
-      'preload_images': [],
-      'preload_audio': [],
[…]

### 2019-11-11 inst/vertical/experiments/Experiment 1/experiment/jspsych-6/plugins/jspsych-image-button-response.js

@@ -1,224 +1,0 @@
-/**
- * jspsych-image-button-response
- * Josh de Leeuw
- *
- * plugin for displaying a stimulus and getting a keyboard response
- *
- * documentation: docs.jspsych.org
- *
- **/
-
-jsPsych.plugins["image-button-response"] = (function() {
-
-  var plugin = {};
-
-  jsPsych.pluginAPI.registerPreload('image-button-response', 'stimulus', 'image');
-
-  plugin.info = {
-    name: 'image-button-response',
-    description: '',
-    parameters: {
-      stimulus: {
-        type: jsPsych.plugins.parameterType.IMAGE,
-        pretty_name: 'Stimulus',
-        default: undefined,
-        description: 'The image to be displayed'
-      },
-      stimulus_height: {
-        type: jsPsych.plugins.parameterType.INT,
-        pretty_name: 'Image height',
-        default: null,
-        description: 'Set the image height in pixels'
-      },
-      stimulus_width: {
-        type: jsPsych.plugins.parameterType.INT,
-        pretty_name: 'Image width',
-        default: null,
-        description: 'Set the image width in pixels'
-      },
-      maintain_aspect_ratio: {
-        type: jsPsych.plugins.parameterType.BOOL,
-        pretty_name: 'Maintain aspect ratio',
-        default: true,
-        description: 'Maintain the aspect ratio after setting width or height'
-      },
-      choices: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Choices',
-        default: undefined,
-        array: true,
-        description: 'The labels for the buttons.'
-      },
-      button_html: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Button HTML',
-        default: '<button class="jspsych-btn">%choice%</button>',
-        array: true,
-        description: 'The html of the button. Can create own style.'
-      },
-      prompt: {
-        type: jsPsych.plugins.parameterType.STRING,
-        pretty_name: 'Prompt',
-        default: null,
-        description: 'Any content here will be displayed under the button.'
-      },
-      stimulus_duration: {
-        type: jsPsych.plugins.parameterType.INT,
-        pretty_name: 'Stimulus duration',
-        default: null,
-        description: 'How long to hide the stimulus.'
-      },
-      trial_duration: {
-        type: jsPsych.plugins.parameterType.INT,
-        pretty_name: 'Trial duration',
-        default: null,
-        description: 'How long to show the trial.'
-      },
[…]

### 2019-11-13 inst/rmarkdown/templates/slidy/skeleton/skeleton.Rmd

--- 2019-11-08 inst/vertical/slides/slidy_example.Rmd
@@
     incremental: true
-    css: custom_css/style.css
 ---
@@
 ```{r, echo = FALSE, out.width="50%", fig.align="center"}
-knitr::include_graphics("imgs/logo.png")
+plot(pressure$temperature, pressure$pressure)
 ```

### 2019-12-09 R/vertical_project.R
manuscript rendering more robust

@@ -108,16 +108,11 @@ build_vertical <- function() {
     "slides/slides.Rmd",
     output_dir = "docs/slides", output_file = "slides.html"
   )
-  # [TODO] file paths for references throw errors here
-  try(
-    {
-      rmarkdown::render(
-        "manuscript/manuscript.Rmd",
-        output_dir = "docs/manuscript",
-        output_file = "manuscript.pdf"
-      )
-    },
-    silent = TRUE
+  # Works, but throws error if file not created first
+  file.create("manuscript/r-references.bib")
+  rmarkdown::render(
+    "manuscript/manuscript.Rmd",
+    output_dir = "docs/manuscript"
   )
   if (dir.exists("experiments")) {
     dir.create("docs/experiments")

### 2019-12-09 R/vertical_project.R
Define functions to init components

@@ -7,6 +7,7 @@
 #'
 #' @export
 vertical_project <- function(path, ...) {
+
   dots <- list(...)
   dir.create(path, recursive = TRUE, showWarnings = FALSE)
 
@@ -16,80 +17,105 @@ vertical_project <- function(path, ...) {
 
   # Git?
   if (dots$init_git) {
-    usethis::use_git(message = "Initial commit")
+    git2r::init(usethis::proj_get())
+    usethis::use_git_ignore(c(".Rhistory", ".RData", ".Rproj.user"))
   }
 
   # pkgdown template
   vertical_pkgdown <- system.file("vertical/_pkgdown.yml", package = "vertical")
   file.copy(vertical_pkgdown, "_pkgdown.yml")
 
-  # Data
   usethis::use_data_raw(open = FALSE)
+  if (dots$init_ms) init_papaja()
+  if (dots$init_som) init_supplemental()
+  if (dots$init_slides) init_slides()
+  if (dots$init_poster) init_poster()
+  if (dots$init_exp) init_jspsych()
+}
 
-  # papaja manuscript
-  if (dots$init_ms) {
-    usethis::use_directory("manuscript", ignore = TRUE)
-    rmarkdown::draft(
-      file = "manuscript/manuscript.Rmd",
-      template = "apa6",
-      package = "papaja",
-      edit = FALSE,
-      create_dir = FALSE
-    )
-  }
+#' Initialize manuscript
+#'
+#' Initialize Rmarkdown APA manuscript in the appropriate location of a vertical project
+#'
+#' @export
+init_papaja <- function() {
+  usethis::use_directory("manuscript", ignore = TRUE)
+  rmarkdown::draft(
+    file = "manuscript/manuscript.Rmd",
+    template = "apa6",
+    package = "papaja",
+    edit = FALSE,
+    create_dir = FALSE
+  )
+}
 
-  # supplementary analysis file
-  if (dots$init_som) {
-    usethis::use_article(
-      "Supplemental_1",
-      title = "Supplementary analyses"
-    )
-  }
+#' Initialize supplemental materials
+#'
+#' Initialize Rmarkdown SOM in the appropriate location of a vertical project
+#'
+#' @export
+init_supplemental <- function() {
+  usethis::use_article(
+    "Supplemental_1",
+    title = "Supplementary analyses"
+  )
+}
 
-  # Slides
-  if (dots$init_slides) {
[…]

### 2019-12-11 Package Helper.R
update getting started

@@ -1,0 +1,22 @@
+# Package helper
+library(usethis) # helpful automation functions for common tasks in package development
+library(pkgdown) # compile your R package to a website
+
+pkgdown::build_site()
+
+# generate description file with your info (need to set your info in advanve)
+usethis::use_description()
+
+# generate readme file, the contents control front page of pkgdown website
+usethis::use_readme_md()
+
+# create a vignette (.Rmd) document that gets compiled and displayed on pkdown website
+usethis::use_vignette("get_started", title="Getting Started")
+
+pkgdown::template_navbar() # generate yml for controlling navbar
+
+render_folder_to_docs("slides")
+
+use_logo("xtra/logo.png")
+
+usethis::use_news_md()

### 2019-12-13 R/vertical_project.R
update version
see news: added yml udpating to build_vertical()

@@ -118,28 +118,119 @@ init_jspsych <- function() {
   file.copy(vertical_jspsych, "experiments/experiment-1/index.html")
 }
 
+#' Update _pkgdown.yml for Rmds in Folders
+#'
+#' @param folder character, name of a vertical folder, e.g., "vignettes"
+#' @param yml_component character, name of the corresponding yml component, e.g., "articles"
+#' @param tab_title character, the title of the component tab written to the website, e.g., "Supplementary"
+#'
+#' @return file, an updated _pkgdown.yml file
+#'
+#' @description This function scrapes a folder and all sub-folders for R markdown documents, and then updates the a designated yml component in _pkgdown.yml with a new list of items. The list of items in shown in the component's tab on the website, and the title of the tab is set to `tab_title`. In the tab, the titles of each .rmd in the main folder are added; then, for each sub-folder the sub-folder name is added (as a text seperator in the tab), followed by the titles for each .rmd in the sub-folder. This function is run automatically by `build_vertical()`.
+#'
+#' @export
+#'
+
+update_yml <- function(folder,yml_component,tab_title){
+
+  if (dir.exists(folder)){
+    # get rmds in folder
+    top_rmds <- list.files(folder, pattern = "\\.Rmd$", full.names = TRUE)
+    top_rmds_yml <- lapply(top_rmds,rmarkdown::yaml_front_matter)
+    top_rmds_title <- unlist(sapply(top_rmds_yml,"[","title"))
+    top_rmds_html <- list.files(folder, pattern = "\\.Rmd$")
+    top_rmds_html <- gsub(".Rmd",".html",top_rmds_html)
+
+    # create temporary articles list
+    new_articles <- list()
+    new_articles <- list(text = tab_title,
+                         menu  = list())
+
+    # add main folder vignettes to list
+    if(length(top_rmds) > 0) {
+      for(i in 1:length(top_rmds_title)){
+        new_articles$menu[[i]] <- list(text = top_rmds_title[i],
+                                       href = paste(folder,top_rmds_html[i],sep="/"))
+      }
+    }
+
+    # add subfolder vignettes
+    sub_folders <- list.dirs(folder, full.names=FALSE, recursive=FALSE)
+    for(i in sub_folders){
+      # get rmds in sub folder
+      sub_rmds <- list.files(paste(folder,i,sep="/"),
[…]

### 2019-12-13 R/vertical_project.R
bux fix to update_yml

@@ -120,7 +120,8 @@ init_jspsych <- function() {
 
 #' Update _pkgdown.yml for Rmds in Folders
 #'
-#' @param folder character, name of a vertical folder, e.g., "vignettes"
+#' @param vertical_folder character, name of a vertical folder, e.g., "vignettes"
+#' @param docs_folder character, name of folder in docs, e.g., "articles"
 #' @param yml_component character, name of the corresponding yml component, e.g., "articles"
 #' @param tab_title character, the title of the component tab written to the website, e.g., "Supplementary"
 #'
@@ -131,15 +132,23 @@ init_jspsych <- function() {
 #' @export
 #'
 
-update_yml <- function(folder,yml_component,tab_title){
+update_yml <- function(vertical_folder,
+                       docs_folder,
+                       yml_component,
+                       tab_title){
 
-  if (dir.exists(folder)){
+  if (dir.exists(vertical_folder)){
     # get rmds in folder
-    top_rmds <- list.files(folder, pattern = "\\.Rmd$", full.names = TRUE)
+    top_rmds <- list.files(vertical_folder, pattern = "\\.Rmd$", full.names = TRUE)
     top_rmds_yml <- lapply(top_rmds,rmarkdown::yaml_front_matter)
     top_rmds_title <- unlist(sapply(top_rmds_yml,"[","title"))
-    top_rmds_html <- list.files(folder, pattern = "\\.Rmd$")
+    top_rmds_html <- list.files(vertical_folder, pattern = "\\.Rmd$")
     top_rmds_html <- gsub(".Rmd",".html",top_rmds_html)
+
+    if (vertical_folder == "manuscript"){
+      top_rmds_html <- list.files(vertical_folder, pattern = "\\.Rmd$")
+      top_rmds_html <- gsub(".Rmd",".pdf",top_rmds_html)
+    }
 
     # create temporary articles list
     new_articles <- list()
@@ -150,15 +159,15 @@ update_yml <- function(folder,yml_component,tab_title){
     if(length(top_rmds) > 0) {
       for(i in 1:length(top_rmds_title)){
         new_articles$menu[[i]] <- list(text = top_rmds_title[i],
-                                       href = paste(folder,top_rmds_html[i],sep="/"))
+                                       href = paste(docs_folder,top_rmds_html[i],sep="/"))
       }
     }
 
     # add subfolder vignettes
-    sub_folders <- list.dirs(folder, full.names=FALSE, recursive=FALSE)
[…]

### 2019-12-13 R/vertical_project.R
add init_vertical_project()
for creating a vertical project in an existing project

@@ -1,4 +1,4 @@
-#' Initialize a **vertical** project
+#' Initialize a **vertical** project as RStudio project template
 #'
 #' This function is called when a **vertical** R project is created.
 #'
@@ -31,6 +31,48 @@ vertical_project <- function(path, ...) {
   if (dots$init_slides) init_slides()
   if (dots$init_poster) init_poster()
   if (dots$init_exp) init_jspsych()
+}
+
+#' Initialize a vertical project inside an existing R project
+#'
+#' @param init_git logical, enables git, init_git=TRUE by default
+#' @param init_ms logical, enables manuscript with papaja, init_ms=TRUE by default
+#' @param init_som logical, enables supplementary with vignettes, init_som=TRUE by default
+#' @param init_slides logical, enables slides with slidy, init_slides=TRUE by default
+#' @param init_poster logical, enables posters with posterdown, init_poster=TRUE by default
+#' @param init_exp logical, enables experiments with jspsych, init_exp=TRUE by default
+#'
+#' @return files, a vertical project template inside an existing R project
+#' @description If you are working inside an existing R project, typically an empty one, or perhaps a project in RStudio cloud, use this function to initialize a vertical project inside the existing project. The name of the existing project (e.g., the parent folder name) must be a valid R package name (numbers, letters, and periods, but no periods at the end, and no spaces, dashes, or underscores).
+#' @export
+#'
+init_vertical_project <- function(init_git = TRUE,
+                                  init_ms = TRUE,
+                                  init_som = TRUE,
+                                  init_slides = TRUE,
+                                  init_poster = TRUE,
+                                  init_exp = TRUE) {
+  path <- getwd()
+  # This is a package
+  usethis::create_package(path, open = FALSE)
+  # setwd(paste0(getwd(), "/", path)) # [TODO] improve this hack
+
+  # Git?
+  if (init_git) {
+    git2r::init(usethis::proj_get())
+    usethis::use_git_ignore(c(".Rhistory", ".RData", ".Rproj.user"))
+  }
+
+  # pkgdown template
+  vertical_pkgdown <- system.file("vertical/_pkgdown.yml", package = "vertical")
+  file.copy(vertical_pkgdown, "_pkgdown.yml")
+
+  usethis::use_data_raw(open = FALSE)
+  if (init_ms) init_papaja()
+  if (init_som) init_supplemental()
+  if (init_slides) init_slides()
+  if (init_poster) init_poster()
+  if (init_exp) init_jspsych()
 }
 
 #' Initialize manuscript

### 2019-12-17 R/vertical_project.R
update with document_data()

@@ -287,3 +287,27 @@ build_vertical <- function(clean=TRUE,update_yml=TRUE,...) {
     file.copy("experiments", "docs", recursive = TRUE)
   }
 }
+
+
+
+#' Add R dataframe and document
+#'
+#' document_data() is a helper function for quickly adding an existing dataframe to an R package. For example, if mydf was an existing data frame, document_data(mydf) would create the data folder (if it doesn't exist), add mydf.rda to the data folder, add mydf.R to the R folder with a roxygen skeleton for documenting the data, and open mydf.R for editing.
+#'
+#' @param ... the name of a single dataframe in the global environment
+#'
+#' @return files, adds dataframe as .rda to data folder, adds .R to R for documentation
+#'
+#' @description A wrapper to usethis::use_data(), usethis::use_r(), and sinew::makeOxygen()
+#'
+#' @export
+#'
+#' @examples
+document_data <- function(...){
+  usethis::use_data(...)
+  data_name <- deparse(substitute(...))
+  usethis::use_r(data_name)
+  switch(menu(c("Yes, add new Roxygen skeleton", "No, show me the file"), title="Overwrite existing .R file?"),
+         cat(sinew::makeOxygen(...),file="R/mydf.R",sep="\n"),2
+  )
+}

### 2019-12-18 R/vertical_project.R
fix some building issues

@@ -9,11 +9,15 @@
 vertical_project <- function(path, ...) {
 
   dots <- list(...)
-  dir.create(path, recursive = TRUE, showWarnings = FALSE)
+  #dir.create(path, recursive = TRUE, showWarnings = FALSE)
+  prjct_name <- path
+  path <- file.path(getwd(), path)
 
   # This is a package
   usethis::create_package(path, open = FALSE)
-  setwd(paste0(getwd(), "/", path)) # [TODO] improve this hack
+  usethis::proj_set(path)
+  #setwd(paste0(getwd(), "/", path)) # [TODO] improve this hack
+  setwd(path)
 
   # Git?
   if (dots$init_git) {
@@ -22,8 +26,10 @@ vertical_project <- function(path, ...) {
   }
 
   # pkgdown template
-  vertical_pkgdown <- system.file("vertical/_pkgdown.yml", package = "vertical")
-  file.copy(vertical_pkgdown, "_pkgdown.yml")
+  #vertical_pkgdown <- system.file("vertical/_pkgdown.yml", package = "vertical")
+  #file.copy(vertical_pkgdown, "_pkgdown.yml")
+  usethis::use_template(template = "_pkgdown.yml",
+                        package = "vertical")
 
   usethis::use_data_raw(open = FALSE)
   if (dots$init_ms) init_papaja()
@@ -96,11 +102,20 @@ init_papaja <- function() {
 #' Initialize Rmarkdown SOM in the appropriate location of a vertical project
 #'
 #' @export
-init_supplemental <- function() {
-  usethis::use_article(
-    "Supplemental_1",
-    title = "Supplementary analyses"
-  )
+init_supplemental <- function(prjct_name=NULL) {
+
+  if(is.null(prjct_name) == FALSE){
+    usethis::use_template(template = "vignette.Rmd",
+                          save_as = "vignettes/Supplementary_1.Rmd",
+                          data = list(vignette_title="Supplementary analyses",
+                                      Package = prjct_name),
+                          package = "usethis")
+  } else{
+    usethis::use_article(
+      "Supplemental_1",
+      title = "Supplementary analyses"
+    )
+  }
 }
 
 #' Initialize slides

### 2019-12-18 R/vertical_project.R
add use_template for _pkgdown.yml

@@ -70,8 +70,10 @@ init_vertical_project <- function(init_git = TRUE,
   }
 
   # pkgdown template
-  vertical_pkgdown <- system.file("vertical/_pkgdown.yml", package = "vertical")
-  file.copy(vertical_pkgdown, "_pkgdown.yml")
+  #vertical_pkgdown <- system.file("vertical/_pkgdown.yml", package = "vertical")
+  #file.copy(vertical_pkgdown, "_pkgdown.yml")
+  usethis::use_template(template = "_pkgdown.yml",
+                        package = "vertical")
 
   usethis::use_data_raw(open = FALSE)
   if (init_ms) init_papaja()

### 2019-12-18 R/vertical_project.R
delete
accidentally added a fix to the wrong function, fix was already present in vertical_project

@@ -70,10 +70,8 @@ init_vertical_project <- function(init_git = TRUE,
   }
 
   # pkgdown template
-  #vertical_pkgdown <- system.file("vertical/_pkgdown.yml", package = "vertical")
-  #file.copy(vertical_pkgdown, "_pkgdown.yml")
-  usethis::use_template(template = "_pkgdown.yml",
-                        package = "vertical")
+  vertical_pkgdown <- system.file("vertical/_pkgdown.yml", package = "vertical")
+  file.copy(vertical_pkgdown, "_pkgdown.yml")
 
   usethis::use_data_raw(open = FALSE)
   if (init_ms) init_papaja()

### 2019-12-18 R/vertical_project.R
fixed init_supplemental()

@@ -10,7 +10,7 @@ vertical_project <- function(path, ...) {
 
   dots <- list(...)
   #dir.create(path, recursive = TRUE, showWarnings = FALSE)
-  prjct_name <- path
+  p_name <- path
   path <- file.path(getwd(), path)
 
   # This is a package
@@ -33,7 +33,7 @@ vertical_project <- function(path, ...) {
 
   usethis::use_data_raw(open = FALSE)
   if (dots$init_ms) init_papaja()
-  if (dots$init_som) init_supplemental()
+  if (dots$init_som) init_supplemental(prjct_name = p_name)
   if (dots$init_slides) init_slides()
   if (dots$init_poster) init_poster()
   if (dots$init_exp) init_jspsych()
@@ -105,7 +105,8 @@ init_papaja <- function() {
 init_supplemental <- function(prjct_name=NULL) {
 
   if(is.null(prjct_name) == FALSE){
-    usethis::use_template(template = "vignette.Rmd",
+    usethis::use_directory("vignettes")
+    usethis::use_template(template = "article.Rmd",
                           save_as = "vignettes/Supplementary_1.Rmd",
                           data = list(vignette_title="Supplementary analyses",
                                       Package = prjct_name),

### 2019-12-18 R/vertical_project.R
improve init_vertical_project
all creation to new project folder from command line

@@ -39,8 +39,10 @@ vertical_project <- function(path, ...) {
   if (dots$init_exp) init_jspsych()
 }
 
-#' Initialize a vertical project inside an existing R project
+#' Initialize a vertical project from command line
 #'
+#' @param project_name character, name of new vertical project, default is project_name = NULL, for initializing inside an existing empty R studio project
+#' @param project_path character, path where new project should be created, default is project_path = NULL for initializing inside an existing empty R studio project
 #' @param init_git logical, enables git, init_git=TRUE by default
 #' @param init_ms logical, enables manuscript with papaja, init_ms=TRUE by default
 #' @param init_som logical, enables supplementary with vignettes, init_som=TRUE by default
@@ -48,19 +50,36 @@ vertical_project <- function(path, ...) {
 #' @param init_poster logical, enables posters with posterdown, init_poster=TRUE by default
 #' @param init_exp logical, enables experiments with jspsych, init_exp=TRUE by default
 #'
-#' @return files, a vertical project template inside an existing R project
-#' @description If you are working inside an existing R project, typically an empty one, or perhaps a project in RStudio cloud, use this function to initialize a vertical project inside the existing project. The name of the existing project (e.g., the parent folder name) must be a valid R package name (numbers, letters, and periods, but no periods at the end, and no spaces, dashes, or underscores).
-#' @export
+#' @return files, a vertical project template inside an existing or new R Studio project
+#' @description Initiliazing a vertical project inside an existing project: The name of the existing project (e.g., the project folder name) must be a valid R package name (numbers, letters, and periods, but no periods at the end, and no spaces, dashes, or underscores). It is not necessary to supply project_name or project_path, they default to the current existing project.
 #'
-init_vertical_project <- function(init_git = TRUE,
+#' Initializing a vertical project as a new project: provide a name and path, and a new vertical project will be set up in an R studio project.
+#'
+#' @export
+#' @examples
+#' \dontrun{
+#' init_vertical_project()
+#' }
+#'
+#'
+init_vertical_project <- function(project_name = NULL,
+                                  project_path = NULL,
+                                  init_git = TRUE,
                                   init_ms = TRUE,
[…]

### 2019-12-18 R/vertical_project.R
udpate documentation with markdown

@@ -1,6 +1,6 @@
 #' Initialize a **vertical** project as RStudio project template
 #'
-#' This function is called when a **vertical** R project is created.
+#' This function is called when a **vertical** R project is created from the RStudio new project template. This function is not intended for use in the console.
 #'
 #' @param path Where to create the project. This must be a valid R package name.
 #' @param ...  Not used.
@@ -41,6 +41,18 @@ vertical_project <- function(path, ...) {
 
 #' Initialize a vertical project from command line
 #'
+#' @description **Initializing a vertical project inside an existing project**: The name of the existing project (e.g., the project folder name) must be a valid R package name (numbers, letters, and periods, but no periods at the end, and no spaces, dashes, or underscores). It is not necessary to supply project_name or project_path, they default to the current existing project.
+#' ```
+#' init_vertical_project()
+#' ```
+#'
+#' **Initializing a vertical project as a new project**: provide a name and path, and a new vertical project will be set up in an R studio project. For example, the following will create the folder `~Desktop/yourname/`, initialize a vertical project template in that folder, and open a new RStudio session with the new project loaded.
+#' ```
+#' init_vertical_project(project_name = "yourname",
+#'                       project_path = "~/Desktop/"
+#'              )
+#' ```
+#'
 #' @param project_name character, name of new vertical project, default is project_name = NULL, for initializing inside an existing empty R studio project
 #' @param project_path character, path where new project should be created, default is project_path = NULL for initializing inside an existing empty R studio project
 #' @param init_git logical, enables git, init_git=TRUE by default
@@ -51,17 +63,8 @@ vertical_project <- function(path, ...) {
 #' @param init_exp logical, enables experiments with jspsych, init_exp=TRUE by default
 #'
 #' @return files, a vertical project template inside an existing or new R Studio project
-#' @description Initiliazing a vertical project inside an existing project: The name of the existing project (e.g., the project folder name) must be a valid R package name (numbers, letters, and periods, but no periods at the end, and no spaces, dashes, or underscores). It is not necessary to supply project_name or project_path, they default to the current existing project.
-#'
[…]

### 2019-12-19 R/vertical_project.R
vertical 0.0.0.9300

@@ -1,6 +1,9 @@
-#' Initialize a **vertical** project as RStudio project template
+#' Initialize **vertical** project from RStudio project template
 #'
-#' This function is called when a **vertical** R project is created from the RStudio new project template. This function is not intended for use in the console.
+#' **NOT for use in the console**
+#'
+#' `vertical_project()` is triggered by loading a new vertical project template from RStudio, and is a wrapper function to create a vertical project structure in the new project folder. The inputs  to this function are selections from the new project template window. Checkbox option for the template are defined in `inst/rstudio/templates/project/vertical_project.dcf`.
+#'
 #'
 #' @param path Where to create the project. This must be a valid R package name.
 #' @param ...  Not used.
@@ -16,8 +19,8 @@ vertical_project <- function(path, ...) {
   # This is a package
   usethis::create_package(path, open = FALSE)
   usethis::proj_set(path)
-  #setwd(paste0(getwd(), "/", path)) # [TODO] improve this hack
-  setwd(path)
+  setwd(path) # [TODO] improve this hack
+
 
   # Git?
   if (dots$init_git) {
@@ -26,11 +29,10 @@ vertical_project <- function(path, ...) {
   }
 
   # pkgdown template
-  #vertical_pkgdown <- system.file("vertical/_pkgdown.yml", package = "vertical")
-  #file.copy(vertical_pkgdown, "_pkgdown.yml")
   usethis::use_template(template = "_pkgdown.yml",
                         package = "vertical")
 
+  # additional content modules
   usethis::use_data_raw(open = FALSE)
   if (dots$init_ms) init_papaja()
   if (dots$init_som) init_supplemental(prjct_name = p_name)
@@ -39,19 +41,7 @@ vertical_project <- function(path, ...) {
   if (dots$init_exp) init_jspsych()
 }
 
-#' Initialize a vertical project from command line
-#'
-#' @description **Initializing a vertical project inside an existing project**: The name of the existing project (e.g., the project folder name) must be a valid R package name (numbers, letters, and periods, but no periods at the end, and no spaces, dashes, or underscores). It is not necessary to supply project_name or project_path, they default to the current existing project.
-#' ```
-#' init_vertical_project()
-#' ```
-#'
[…]

### 2019-12-19 R/vertical_project.R
enhanced init_vertical_project

@@ -43,6 +43,8 @@ vertical_project <- function(path, ...) {
 
 #' Initialize vertical project from command line
 #'
+#' Use `vertical::init_vertical_project()` to initialize a vertical project from the command line.
+#'
 #' @param project_name character, name of new vertical project, default is project_name = NULL, for initializing inside an existing empty R studio project
 #' @param project_path character, path where new project should be created, default is project_path = NULL for initializing inside an existing empty R studio project
 #' @param init_git logical, enables git, init_git=TRUE by default
@@ -55,61 +57,141 @@ vertical_project <- function(path, ...) {
 #' @return **files**, a vertical project template inside an existing or new R Studio project
 #'
 #' @details
-#' **Initializing a vertical project inside an existing project**:
+#'
+#' **Interactive: Get asked what you want to do**
 #' ```
-#' init_vertical_project()
+#' vertical::init_vertical_project()
+#' ```
+#' This opens a dialogue in the console. You choose whether to make a brand new RStudio project, or to initialize vertical in an existing project that you are already working. Follow the prompts and choose which vertical modules to load.
+#'
+#' **Scripted: Do what you want to do**
+#'
+#' If any input parameters are specified, then initialization is done directly without a dialogue.
+#'
+#' **Initializing a vertical project inside an existing project**:
+#' It is not necessary to name the project, because a named project should already be open. At least one of the modules should be specified. Default NULL values are set to FALSE, and those modules are not loaded.
+#' ```
+#' init_vertical_project(init_git=TRUE,...)
 #' ```
 #' The name of the existing project (e.g., the project folder name) must be a valid R package name (numbers, letters, and periods, but no periods at the end, and no spaces, dashes, or underscores). It is not necessary to supply project_name or project_path, they default to the current existing project.
 #'
 #'
 #' **Initializing a vertical project as a new project**:
+#' Provide a name and path to set up a new vertical project as an R studio project, and specify which modules should be loaded by setting their parameters to TRUE.
 #' ```
 #' init_vertical_project(project_name = "yourname",
-#'                       project_path = "~/Desktop/")
[…]

### 2019-12-19 R/vertical_project.R
fix document_data

@@ -482,6 +482,6 @@ document_data <- function(...){
   data_name <- deparse(substitute(...))
   usethis::use_r(data_name)
   switch(menu(c("Yes, add new Roxygen skeleton", "No, show me the file"), title="Overwrite existing .R file?"),
-         cat(sinew::makeOxygen(...),file="R/mydf.R",sep="\n"),2
+         cat(sinew::makeOxygen(...),file=paste0(file.path("R",data_name),".R"),sep="\n"),2
   )
 }

### 2019-12-19 R/vertical_project.R
make jspsych independent from jspsychr

@@ -286,15 +286,14 @@ init_poster <- function() {
 #' This function does the following:
 #' 1. creates the `experiments` folder
 #' 2. Downloads the most recent `jspsych`` library from <https://github.com/jspsych/jsPsych/releases>
-#' 3. Adds a `jspsychr` template (Experiment_1), which is an example of using R Studio and R Markdown to author a `jspsych` experiment
+#' 3. Adds a minimal `jspsych`` example experiment to `experiments/experiment-1/`
 #'
 #' See the [jspsych documentation](https://www.jspsych.org) for more information about using jspsych to build behavioral experiments for the web.
-#'
-#' See the [jspsychr documentation](https://crumplab.github.io/jspsychr/) for more information about using R Markdown to write `jspsych` experiments.
 #'
 #' @export
 init_jspsych <- function() {
   usethis::use_directory("experiments", ignore = TRUE)
+  usethis::use_directory("experiments/experiment-1")
   # Get latest jsPsych version download link
   ver <- basename(httr::GET("https://github.com/jspsych/jsPsych/releases/latest")$url)
   loc_from <- paste0(
@@ -308,7 +307,26 @@ init_jspsych <- function() {
   unlink(loc_to)
   # Suggest deleting unnecessary large folder
   message(paste0("Consider removing ", sub(".zip", "", loc_to), "/examples"))
+  usethis::use_template(template = "experiment.html",
+                        save_as = "experiments/experiment-1/experiment.html",
+                        package = "vertical")
+}
 
+#' Initialize jspsychr experiment
+#'
+#' Initialize a jspsychr template in the `experiments` folder of a vertical project
+#'
+#' This function does the following:
+#' 1. creates the `experiments` folder
+#' 2. Adds a `jspsychr` template (Experiment_1), which is an example of using R Studio and R Markdown to author a `jspsych` experiment
+#'
+#' See the [jspsych documentation](https://www.jspsych.org) for more information about using jspsych to build behavioral experiments for the web.
+#'
+#' See the [jspsychr documentation](https://crumplab.github.io/jspsychr/) for more information about using R Markdown to write `jspsych` experiments.
+#'
+#' @export
+init_jspsychr <- function(){
+  usethis::use_directory("experiments", ignore = TRUE)
   # add jspsychr template example
   rmarkdown::draft(
     file = "experiments/Experiment_1.Rmd",

### 2019-12-19 R/vertical_project.R
improve documentation

@@ -490,10 +490,30 @@ build_vertical <- function(clean=TRUE,update_yml=FALSE,...) {
 #' 3. add `mydf.R` to the R folder with a roxygen skeleton for documenting the data
 #' 4. open `mydf.R` for editing.
 #'
+#' This function is intended for convenience because it wraps a few common tasks together when adding data to an R package. We recommend that users familiarize themselves with the underlying functions, which are not features of vertical. See the details section for some pointers.
+#'
 #' @section Usage:
 #' ```
 #' document_data(mydf)
 #' ```
+#' @details
+#'
+#' We highly recommend understanding the basics of adding functions and data to R packages by reading Hadley Wickham's excellent R package guide <http://r-pkgs.had.co.nz>. Below we briefly elaborate on the individual action of functions inside `document_data()`, which can be used individually to automate aspects of documenting data sets.
+#'
+#'  **export dataframe to data folder as .rda**
+#'  `usethis::use_data(mydf)` accepts an existing dataframe (e.g., mydf). It writes the `data` folder if it does not already exist and exports the dataframe as an .rda file to the `data` folder (eg., `data/mydf.rda`)
+#'
+#'  **create documentation .R file**
+#'  `usethis::use_r("mydf")` creates a new .R file in the `R/` folder (e.g., `R/mydf.R`). This is a blank file where `roxygen2` comments can be added to document the data set.
+#'
+#'  **roxygen comments for documentation**
+#'  `sinew::makeOxygen(mydf)` generates a skeleton roxygen comment block that can be edited to document the data. This function prints the skeleton to the console. The user then copies it to `mydf.R`, and edits as necessary.
+#'
+#'  **`document_data()`** is a wrapper to the above three main functions. After the roxygen comments have been added to the `.R` file, they should be compiled...
+#'
+#'  **compiling roxygen comments**
+#'  `devtools::document()` can be run to compile any .R functions in `R/` with roxygen comments.
+#'
 #' @export
 document_data <- function(...){
   usethis::use_data(...)

### 2019-12-20 R/vertical_project.R
delete document_data

@@ -477,49 +477,4 @@ build_vertical <- function(clean=TRUE,update_yml=FALSE,...) {
 
 
 
-#' Add R dataframe and documentation to R Package
-#'
-#' @param ... the name of a single dataframe in the global environment
-#'
-#' @return files, adds dataframe as .rda to data folder, adds .R to R for documentation
-#'
-#' @description `document_data()` is a helper function for including a dataframe in an R package, and documenting the data set. It is a wrapper to `usethis::use_data()`, `usethis::use_r()`, and `sinew::makeOxygen()`. For example, if `mydf` was an existing data frame, `document_data(mydf)` would:
-#'
-#' 1. create the `data`` folder (if it doesn't exist)
-#' 2. add `mydf.rda` to the data folder
-#' 3. add `mydf.R` to the R folder with a roxygen skeleton for documenting the data
-#' 4. open `mydf.R` for editing.
-#'
-#' This function is intended for convenience because it wraps a few common tasks together when adding data to an R package. We recommend that users familiarize themselves with the underlying functions, which are not features of vertical. See the details section for some pointers.
-#'
-#' @section Usage:
-#' ```
-#' document_data(mydf)
-#' ```
-#' @details
-#'
-#' We highly recommend understanding the basics of adding functions and data to R packages by reading Hadley Wickham's excellent R package guide <http://r-pkgs.had.co.nz>. Below we briefly elaborate on the individual action of functions inside `document_data()`, which can be used individually to automate aspects of documenting data sets.
-#'
-#'  **export dataframe to data folder as .rda**
-#'  `usethis::use_data(mydf)` accepts an existing dataframe (e.g., mydf). It writes the `data` folder if it does not already exist and exports the dataframe as an .rda file to the `data` folder (eg., `data/mydf.rda`)
-#'
-#'  **create documentation .R file**
-#'  `usethis::use_r("mydf")` creates a new .R file in the `R/` folder (e.g., `R/mydf.R`). This is a blank file where `roxygen2` comments can be added to document the data set.
-#'
-#'  **roxygen comments for documentation**
-#'  `sinew::makeOxygen(mydf)` generates a skeleton roxygen comment block that can be edited to document the data. This function prints the skeleton to the console. The user then copies it to `mydf.R`, and edits as necessary.
-#'
[…]

### 2019-12-20 R/vertical_project.R
fix Supplementary_1.html
and remove documentation for vertical_project()

@@ -1,14 +1,3 @@
-#' Initialize **vertical** project from RStudio project template
-#'
-#' **NOT for use in the console**
-#'
-#' `vertical_project()` is triggered by loading a new vertical project template from RStudio, and is a wrapper function to create a vertical project structure in the new project folder. The inputs  to this function are selections from the new project template window. Checkbox option for the template are defined in `inst/rstudio/templates/project/vertical_project.dcf`.
-#'
-#'
-#' @param path Where to create the project. This must be a valid R package name.
-#' @param ...  Not used.
-#'
-#' @export
 vertical_project <- function(path, ...) {
 
   dots <- list(...)

### 2019-12-20 R/build.R
Separate source code to multiple files #37

@@ -1,0 +1,57 @@
+#' Build vertical project
+#'
+#' Build the website associated with a vertical project. This is mostly a wrapper to `pkgdown::build_site()`, but extended to render content from vertical folders.
+#'
+#' @param clean logical, when clean=TRUE (the default), the `docs` folder is cleaned (e.g., completely wiped) using `pkgdown::clean_site()`, otherwise when clean=FALSE `clean_site()` will not be run.
+#' @param update_yml logical, update_yml=FALSE is the default so existing yml **is not overwritten**, updates yml components in `_pkgdown.yml` to list all .Rmds in folders and subfolders of vertical components (manuscript, posters, slides, vignettes) in associated navigation bar tabs on the website. Set to TRUE to overwrite.
+#' @param ... params, pass additional parameters to `pkgdown::build_site(...)`
+#' @section Usage:
+#' ```
+#' build_vertical()
+#' ```
+#'
+#' @export
+build_vertical <- function(clean=TRUE,update_yml=FALSE,...) {
+  if(update_yml == TRUE) {
+    update_yml("vignettes","articles","articles","Supplementary")
+    update_yml("manuscript","manuscript","manuscript","PDF")
+    update_yml("posters","posters","posters","Poster")
+    update_yml("slides","slides","slides","Slides")
+  }
+  if(clean == TRUE) pkgdown::clean_site()
+  pkgdown::build_site(...)
+
+  if (dir.exists("posters")) {
+    file_names <- list.files("posters", pattern = "\\.Rmd$")
+    for(i in file_names){
+      rmarkdown::render(paste("posters",i,sep="/"),
+                        output_dir = "docs/posters/")
+    }
+  }
+
+  if (dir.exists("slides")) {
+    file_names <- list.files("slides", pattern = "\\.Rmd$")
+    for(i in file_names){
+      rmarkdown::render(paste("slides",i,sep="/"),
+                        output_dir = "docs/slides/")
+    }
+  }
+
+  if (dir.exists("manuscript")) {
+    # Works, but throws error if file not created first
+    file.create("manuscript/r-references.bib")
+    file_names <- list.files("manuscript", pattern = "\\.Rmd$")
+    for(i in file_names){
+      rmarkdown::render(paste("manuscript",i,sep="/"),
+                        output_dir = "docs/manuscript/")
+    }
+  }
+  if (dir.exists("experiments")) {
+    dir.create("docs/experiments")
+    file.copy("experiments", "docs", recursive = TRUE)
+  }
+}
+
+
+
+

### 2019-12-20 R/initialize.R

@@ -325,145 +325,3 @@ init_jspsychr <- function(){
     create_dir = TRUE
   )
 }
-
-#' Update _pkgdown.yml for Rmds in vertical folders
-#'
-#' Write yml for the website navigation bar; run automatically by `build_vertical()`.
-#'
-#' @param vertical_folder character, name of a vertical folder, e.g., "vignettes"
-#' @param docs_folder character, name of folder in docs, e.g., "articles"
-#' @param yml_component character, name of the corresponding yml component, e.g., "articles"
-#' @param tab_title character, the title of the component tab written to the website, e.g., "Supplementary"
-#'
-#' @return file, an updated _pkgdown.yml file
-#'
-#' @details `update_yml()` is a helper function for adding vertical content to the website navigation bar by upating `_pkgdown.yml`.
-#'
-#' Specifically, the titles of .Rmds in  `vertical_folder` and its subfolders are listed under the associated navbar component in `_pkgdown.yml`. If subfolders exist, then the name of the subfolder is used as a section header in the list. The sections and titles are shown in the dropdown navigation tab on the website.
-#'
-#' `_pkgdown.yml` can be further modified by hand to achieve various customizations to the website. See the [pkgdown documentation](https://pkgdown.r-lib.org/reference/build_site.html) for additional information.
-#'
-#' @export
-#'
-
-update_yml <- function(vertical_folder,
-                       docs_folder,
-                       yml_component,
-                       tab_title){
-
-  if (dir.exists(vertical_folder)){
-    # get rmds in folder
-    top_rmds <- list.files(vertical_folder, pattern = "\\.Rmd$", full.names = TRUE)
-    top_rmds_yml <- lapply(top_rmds,rmarkdown::yaml_front_matter)
-    top_rmds_title <- unlist(sapply(top_rmds_yml,"[","title"))
-    top_rmds_html <- list.files(vertical_folder, pattern = "\\.Rmd$")
-    top_rmds_html <- gsub(".Rmd",".html",top_rmds_html)
-
-    if (vertical_folder == "manuscript"){
-      top_rmds_html <- list.files(vertical_folder, pattern = "\\.Rmd$")
-      top_rmds_html <- gsub(".Rmd",".pdf",top_rmds_html)
-    }
-
-    # create temporary articles list
-    new_articles <- list()
-    new_articles <- list(text = tab_title,
[…]

### 2019-12-20 R/initialize.R
Don't delete documentation of internal function

@@ -1,3 +1,12 @@
+#' Initialize **vertical** project from RStudio project template
+#'
+#' **NOT for use in the console**
+#'
+#' `vertical_project()` is triggered by loading a new vertical project template from RStudio, and is a wrapper function to create a vertical project structure in the new project folder. The inputs  to this function are selections from the new project template window. Checkbox option for the template are defined in `inst/rstudio/templates/project/vertical_project.dcf`.
+#'
+#'
+#' @param path Where to create the project. This must be a valid R package name.
+#' @param ...  Not used.
 vertical_project <- function(path, ...) {
 
   dots <- list(...)

### 2019-12-20 R/vertical_project.R
yml update

@@ -429,6 +429,7 @@ build_vertical <- function(clean=TRUE,update_yml=FALSE,...) {
     update_yml("manuscript","manuscript","manuscript","PDF")
     update_yml("posters","posters","posters","Poster")
     update_yml("slides","slides","slides","Slides")
+    update_yml("experiments","experiments","experiments","Experiments")
   }
   if(clean == TRUE) pkgdown::clean_site()
   pkgdown::build_site(...)

### 2019-12-20 R/build.R
add experiments to yml update

@@ -17,6 +17,7 @@ build_vertical <- function(clean=TRUE,update_yml=FALSE,...) {
     update_yml("manuscript","manuscript","manuscript","PDF")
     update_yml("posters","posters","posters","Poster")
     update_yml("slides","slides","slides","Slides")
+    update_yml("experiments","experiments","experiments","Experiments")
   }
   if(clean == TRUE) pkgdown::clean_site()
   pkgdown::build_site(...)

### 2019-12-20 R/initialize.R
refactor vertical_project()

@@ -9,34 +9,37 @@
 #' @param ...  Not used.
 vertical_project <- function(path, ...) {
 
-  dots <- list(...)
-  #dir.create(path, recursive = TRUE, showWarnings = FALSE)
-  p_name <- path
-  path <- file.path(getwd(), path)
-
-  # This is a package
-  usethis::create_package(path, open = FALSE)
-  usethis::proj_set(path)
-  setwd(path) # [TODO] improve this hack
-
-
-  # Git?
-  if (dots$init_git) {
-    git2r::init(usethis::proj_get())
-    usethis::use_git_ignore(c(".Rhistory", ".RData", ".Rproj.user"))
-  }
-
-  # pkgdown template
-  usethis::use_template(template = "_pkgdown.yml",
-                        package = "vertical")
-
-  # additional content modules
-  usethis::use_data_raw(open = FALSE)
-  if (dots$init_ms) init_papaja()
-  if (dots$init_som) init_supplemental(prjct_name = p_name)
-  if (dots$init_slides) init_slides()
-  if (dots$init_poster) init_poster()
-  if (dots$init_exp) init_jspsych()
+  init_vertical_project(project_name = path,
+                        project_path = getwd(),
+                        ...)
+  # dots <- list(...)
+  # #dir.create(path, recursive = TRUE, showWarnings = FALSE)
+  # p_name <- path
+  # path <- file.path(getwd(), path)
+  #
+  # # This is a package
+  # usethis::create_package(path, open = FALSE)
+  # usethis::proj_set(path)
+  # setwd(path) # [TODO] improve this hack
+  #
+  #
+  # # Git?
+  # if (dots$init_git) {
+  #   git2r::init(usethis::proj_get())
+  #   usethis::use_git_ignore(c(".Rhistory", ".RData", ".Rproj.user"))
+  # }
+  #
+  # # pkgdown template
+  # usethis::use_template(template = "_pkgdown.yml",
+  #                       package = "vertical")
+  #
+  # # additional content modules
+  # usethis::use_data_raw(open = FALSE)
+  # if (dots$init_ms) init_papaja()
+  # if (dots$init_som) init_supplemental(prjct_name = p_name)
+  # if (dots$init_slides) init_slides()
+  # if (dots$init_poster) init_poster()
+  # if (dots$init_exp) init_jspsych()
 }
 
 #' Initialize vertical project from command line

### 2019-12-20 R/initialize.R
Simplify build; one function to rule them all

@@ -1,194 +1,47 @@
-#' Initialize **vertical** project from RStudio project template
+#' Initialize a **vertical** project
 #'
-#' **NOT for use in the console**
+#' This function is called when a **vertical** R project is created.
 #'
-#' `vertical_project()` is triggered by loading a new vertical project template from RStudio, and is a wrapper function to create a vertical project structure in the new project folder. The inputs  to this function are selections from the new project template window. Checkbox option for the template are defined in `inst/rstudio/templates/project/vertical_project.dcf`.
-#'
-#'
-#' @param path Where to create the project. This must be a valid R package name.
+#' @param path Name (and location) of package. Must be a valid R package name.
 #' @param ...  Not used.
-vertical_project <- function(path, ...) {
+#'
+#' @export
+vertical_project <- function(path,
+                             init_git = TRUE,
+                             init_data = TRUE,
+                             init_ms = TRUE,
+                             init_som = TRUE,
+                             init_slides = TRUE,
+                             init_poster = TRUE,
+                             init_exp = FALSE,
+                             ...) {
 
   dots <- list(...)
-  #dir.create(path, recursive = TRUE, showWarnings = FALSE)
-  p_name <- path
-  path <- file.path(getwd(), path)
-
-  # This is a package
-  usethis::create_package(path, open = FALSE)
-  usethis::proj_set(path)
-  setwd(path) # [TODO] improve this hack
-
-
-  # Git?
-  if (dots$init_git) {
-    git2r::init(usethis::proj_get())
-    usethis::use_git_ignore(c(".Rhistory", ".RData", ".Rproj.user"))
+  if (length(dots) == 0) {
+    dots <- append(dots, list(
+      init_git = init_git,
+      init_data = init_data,
+      init_ms = init_ms,
+      init_som = init_som,
+      init_slides = init_slides,
+      init_poster = init_poster,
+      init_exp = init_exp
+    ))
   }
-
-  # pkgdown template
+  usethis::create_package(path)
+  setwd(path)
   usethis::use_template(template = "_pkgdown.yml",
                         package = "vertical")
-
-  # additional content modules
-  usethis::use_data_raw(open = FALSE)
+  if (dots$init_git) {
+    git2r::init()
+    usethis::use_git_ignore(c(".Rhistory", ".RData", ".Rproj.user"))
+  }
+  if (dots$init_data) usethis::use_data_raw(open = FALSE)
   if (dots$init_ms) init_papaja()
[…]

### 2019-12-20 R/initialize.R
add basename

@@ -38,7 +38,7 @@ vertical_project <- function(path,
   }
   if (dots$init_data) usethis::use_data_raw(open = FALSE)
   if (dots$init_ms) init_papaja()
-  if (dots$init_som) init_supplemental(prjct_name = path)
+  if (dots$init_som) init_supplemental(prjct_name = basename(path))
   if (dots$init_slides) init_slides()
   if (dots$init_poster) init_poster()
   if (dots$init_exp) init_jspsych()

### 2019-12-20 R/initialize.R
Simplify init SOM

@@ -38,10 +38,10 @@ vertical_project <- function(path,
   }
   if (dots$init_data) usethis::use_data_raw(open = FALSE)
   if (dots$init_ms) init_papaja()
-  if (dots$init_som) init_supplemental(prjct_name = basename(path))
   if (dots$init_slides) init_slides()
   if (dots$init_poster) init_poster()
   if (dots$init_exp) init_jspsych()
+  if (dots$init_som) init_som()
 
 }
 
@@ -77,21 +77,15 @@ init_papaja <- function() {
 #' See [usethis documentation](https://usethis.r-lib.org/reference/use_vignette.html) for more information.
 #'
 #' @export
-init_supplemental <- function(prjct_name=NULL) {
-
-  if(is.null(prjct_name) == FALSE){
-    usethis::use_directory("vignettes")
-    usethis::use_template(template = "article.Rmd",
-                          save_as = "vignettes/Supplementary_1.Rmd",
-                          data = list(vignette_title="Supplementary analyses",
-                                      Package = prjct_name),
-                          package = "usethis")
-  } else{
-    usethis::use_article(
-      "Supplemental_1",
-      title = "Supplementary analyses"
-    )
-  }
+init_som <- function() {
+  usethis::use_directory("vignettes")
+  usethis::use_template(
+    template = "article.Rmd",
+    save_as = "vignettes/som.Rmd",
+    data = list(vignette_title="Supplementary analyses",
+                Package = basename(getwd())),
+    package = "usethis"
+  )
 }
 
 #' Initialize slides

### 2019-12-20 R/initialize.R
Provide data.R template and processing script that actually works

@@ -36,13 +36,28 @@ vertical_project <- function(path,
     git2r::init()
     usethis::use_git_ignore(c(".Rhistory", ".RData", ".Rproj.user"))
   }
-  if (dots$init_data) usethis::use_data_raw(open = FALSE)
+  if (dots$init_data) init_data()
   if (dots$init_ms) init_papaja()
   if (dots$init_slides) init_slides()
   if (dots$init_poster) init_poster()
   if (dots$init_exp) init_jspsych()
   if (dots$init_som) init_som()
 
+}
+
+init_data <- function() {
+  usethis::use_directory("data-raw", ignore = TRUE)
+  usethis::use_template(
+    template = "preprocess.R",
+    save_as = "data-raw/preprocess.R",
+    package = "vertical"
+  )
+  usethis::use_template(
+    template = "data.R",
+    save_as = "R/data.R",
+    data = list(dataname = "mydata"),
+    package = "vertical"
+  )
 }
 
 #' Initialize manuscript

### 2019-12-20 inst/templates/data.R

@@ -1,0 +1,9 @@
+#' @title A dataset
+#' @description Example data frame
+#' @format A data.frame with . rows and . variables:
+#' \describe{
+#'   \item{\code{a}}{Variable a; integer.}
+#'   \item{\code{b}}{Variable b; factor with 2 levels (x, y).}
+#'}
+#' @details This dataset is an example.
+"{{{dataname}}}"

### 2019-12-20 inst/templates/preprocess.R

@@ -1,0 +1,8 @@
+## Example data preprocessing script
+
+# Load data
+mydata <- data.frame(a = 1, b = 2)
+
+# Processing ...
+
+usethis::use_data(mydata)

### 2019-12-20 R/initialize.R
update vertical_project

@@ -6,7 +6,7 @@
 #' @param ...  Not used.
 #'
 #' @export
-vertical_project <- function(path,
+vertical_project <- function(path=NULL,
                              init_git = TRUE,
                              init_data = TRUE,
                              init_ms = TRUE,
@@ -15,6 +15,11 @@ vertical_project <- function(path,
                              init_poster = TRUE,
                              init_exp = FALSE,
                              ...) {
+  if(is.null(path)) {
+    rstudioapi::executeCommand('newProject', quiet = FALSE)
+    return(usethis::ui_info("Opening RStudio Project Template, select New Directory, then Vertical Research Project"))
+
+  }
 
   dots <- list(...)
   if (length(dots) == 0) {

### 2019-12-20 R/initialize.R
Update initialize.R

@@ -18,7 +18,6 @@ vertical_project <- function(path=NULL,
   if(is.null(path)) {
     rstudioapi::executeCommand('newProject', quiet = FALSE)
     return(usethis::ui_info("Opening RStudio Project Template, select New Directory, then Vertical Research Project"))
-
   }
 
   dots <- list(...)

### 2019-12-20 R/initialize.R
update instructions for vertical_project()

@@ -6,7 +6,7 @@
 #' @param ...  Not used.
 #'
 #' @export
-vertical_project <- function(path,
+vertical_project <- function(path=NULL,
                              init_git = TRUE,
                              init_data = TRUE,
                              init_ms = TRUE,
@@ -15,6 +15,10 @@ vertical_project <- function(path,
                              init_poster = TRUE,
                              init_exp = FALSE,
                              ...) {
+  if(is.null(path)) {
+    rstudioapi::executeCommand('newProject', quiet = FALSE)
+    return(usethis::ui_info("Opening RStudio Project Template, select New Directory, then Vertical Research Project"))
+  }
 
   dots <- list(...)
   if (length(dots) == 0) {

### 2019-12-23 R/initialize.R
fix build warnings

@@ -42,7 +42,9 @@ vertical_project <- function(path = NULL,
   usethis::create_package(path)
   setwd(path)
   usethis::use_template(template = "_pkgdown.yml",
+                        ignore = TRUE,
                         package = "vertical")
+  usethis::use_mit_license()
   if (dots$init_git) {
     git2r::init()
     usethis::use_git_ignore(c(".Rhistory", ".RData", ".Rproj.user"))
@@ -108,6 +110,7 @@ init_som <- function() {
   usethis::use_template(
     template = "article.Rmd",
     save_as = "vignettes/som.Rmd",
+    ignore = TRUE,
     data = list(vignette_title="Supplementary analyses",
                 Package = basename(getwd())),
     package = "usethis"

### 2019-12-23 R/build.R
WIP #47

@@ -22,31 +22,13 @@ build_vertical <- function(clean=TRUE,update_yml=FALSE,...) {
   if(clean == TRUE) pkgdown::clean_site()
   pkgdown::build_site(...)
 
-  if (dir.exists("posters")) {
-    file_names <- list.files("posters", pattern = "\\.Rmd$")
-    for(i in file_names){
-      rmarkdown::render(paste("posters",i,sep="/"),
-                        output_dir = "docs/posters/")
+  for (i in list.files(pattern = "\\.Rmd", recursive = TRUE)) {
+    rmarkdown::render(i, output_dir = paste0("docs/", i))
+    if (!any(grepl(tools::file_path_sans_ext(i), readLines("_pkgdown.yml")))) {
+      warning(i, " is not linked to in navbar. Please edit _pkgdown.yml")
     }
   }
 
-  if (dir.exists("slides")) {
-    file_names <- list.files("slides", pattern = "\\.Rmd$")
-    for(i in file_names){
-      rmarkdown::render(paste("slides",i,sep="/"),
-                        output_dir = "docs/slides/")
-    }
-  }
-
-  if (dir.exists("manuscript")) {
-    # Works, but throws error if file not created first
-    file.create("manuscript/r-references.bib")
-    file_names <- list.files("manuscript", pattern = "\\.Rmd$")
-    for(i in file_names){
-      rmarkdown::render(paste("manuscript",i,sep="/"),
-                        output_dir = "docs/manuscript/")
-    }
-  }
   if (dir.exists("experiments")) {
     dir.create("docs/experiments")
     file.copy("experiments", "docs", recursive = TRUE)

### 2019-12-23 R/build.R
Update build.R
- don't build vignettes
- usethis::ui_info rather than warnings for _pkgdown.yml
- report output file names for warnings, not .rmd

@@ -23,9 +23,13 @@ build_vertical <- function(clean=TRUE,update_yml=FALSE,...) {
   pkgdown::build_site(...)
 
   for (i in list.files(pattern = "\\.Rmd", recursive = TRUE)) {
-    rmarkdown::render(i, output_dir = paste0("docs/", i))
-    if (!any(grepl(tools::file_path_sans_ext(i), readLines("_pkgdown.yml")))) {
-      warning(i, " is not linked to in navbar. Please edit _pkgdown.yml")
+    if(dirname(i) != "vignettes"){
+      out.file <- rmarkdown::render(i, output_dir = paste0("docs/", dirname(i)), quiet=TRUE)
+      if (!any(grepl(tools::file_path_sans_ext(i), readLines("_pkgdown.yml")))) {
+        #warning(i, " is not linked to in navbar. Please edit _pkgdown.yml")
+        usethis::ui_info(paste(file.path(dirname(i),basename(out.file)),
+                               " is not linked to in navbar. Please edit _pkgdown.yml"))
+      }
     }
   }
 

### 2019-12-24 R/build.R
suggest_yml
- deleted update_yml()
- removed update_yml() from build_vertical()
- added suggest_yml() as a standalone function to suggest yml for _pkgdown.yml

@@ -1,9 +1,8 @@
 #' Build vertical project
 #'
-#' Build the website associated with a vertical project. This is mostly a wrapper to `pkgdown::build_site()`, but extended to render content from vertical folders.
+#' Build the website associated with a vertical project. This is mostly a wrapper to `pkgdown::build_site()`, but extended to render .Rmd content from other folders.
 #'
 #' @param clean logical, when clean=TRUE (the default), the `docs` folder is cleaned (e.g., completely wiped) using `pkgdown::clean_site()`, otherwise when clean=FALSE `clean_site()` will not be run.
-#' @param update_yml logical, update_yml=FALSE is the default so existing yml **is not overwritten**, updates yml components in `_pkgdown.yml` to list all .Rmds in folders and subfolders of vertical components (manuscript, posters, slides, vignettes) in associated navigation bar tabs on the website. Set to TRUE to overwrite.
 #' @param ... params, pass additional parameters to `pkgdown::build_site(...)`
 #' @section Usage:
 #' ```
@@ -11,14 +10,8 @@
 #' ```
 #'
 #' @export
-build_vertical <- function(clean=TRUE,update_yml=FALSE,...) {
-  if(update_yml == TRUE) {
-    update_yml("vignettes","articles","articles","Supplementary")
-    update_yml("manuscript","manuscript","manuscript","PDF")
-    update_yml("posters","posters","posters","Poster")
-    update_yml("slides","slides","slides","Slides")
-    update_yml("experiments","experiments","experiments","Experiments")
-  }
+build_vertical <- function(clean=TRUE,...) {
+
   if(clean == TRUE) pkgdown::clean_site()
   pkgdown::build_site(...)
 
@@ -26,7 +19,6 @@ build_vertical <- function(clean=TRUE,update_yml=FALSE,...) {
     if(dirname(i) != "vignettes"){
       out.file <- rmarkdown::render(i, output_dir = paste0("docs/", dirname(i)), quiet=TRUE)
       if (!any(grepl(tools::file_path_sans_ext(i), readLines("_pkgdown.yml")))) {
-        #warning(i, " is not linked to in navbar. Please edit _pkgdown.yml")
         usethis::ui_info(paste(file.path(dirname(i),basename(out.file)),
                                " is not linked to in navbar. Please edit _pkgdown.yml"))
       }

### 2019-12-24 R/maintain.R

@@ -1,81 +1,50 @@
-#' Update _pkgdown.yml for Rmds in vertical folders
+#' Suggest yml for _pkgdown.yml
 #'
-#' Write yml for the website navigation bar; run automatically by `build_vertical()`.
+#' Run `vertical::suggest_yml` to generate a suggestions for populating `_pkgdown.yml`. The suggested yml is written to the console. Copy this to `_pkgdown.yml` and modify as necessary.
 #'
-#' @param vertical_folder character, name of a vertical folder, e.g., "vignettes"
-#' @param docs_folder character, name of folder in docs, e.g., "articles"
-#' @param yml_component character, name of the corresponding yml component, e.g., "articles"
-#' @param tab_title character, the title of the component tab written to the website, e.g., "Supplementary"
+#' @return yml copied to console
 #'
-#' @return file, an updated _pkgdown.yml file
-#'
-#' @details `update_yml()` is a helper function for adding vertical content to the website navigation bar by upating `_pkgdown.yml`.
-#'
-#' Specifically, the titles of .Rmds in  `vertical_folder` and its subfolders are listed under the associated navbar component in `_pkgdown.yml`. If subfolders exist, then the name of the subfolder is used as a section header in the list. The sections and titles are shown in the dropdown navigation tab on the website.
-#'
-#' `_pkgdown.yml` can be further modified by hand to achieve various customizations to the website. See the [pkgdown documentation](https://pkgdown.r-lib.org/reference/build_site.html) for additional information.
+#' @details `_pkgdown.yml` can be further modified by hand to achieve various customizations to the website. See the [pkgdown documentation](https://pkgdown.r-lib.org/reference/build_site.html) for additional information.
 #'
 #' @export
-update_yml <- function(vertical_folder,
-                       docs_folder,
-                       yml_component,
-                       tab_title){
+#'
+suggest_yml <- function(){
+  temp_yml <- yaml::read_yaml("_pkgdown.yml")
+  temp_yml$navbar$components$articles <- NULL
 
-  if (dir.exists(vertical_folder)){
-    # get rmds in folder
-    top_rmds <- list.files(vertical_folder, pattern = "\\.Rmd$", full.names = TRUE)
[…]

### 2019-12-24 R/maintain.R
add user tip about .pdf

@@ -46,5 +46,6 @@ suggest_yml <- function(){
   }
 
   usethis::ui_info("Copy to _pkgdown.yml then modify as needed")
+  usethis::ui_info("tip: change .html to .pdf in hrefs for papaja manuscripts, or other .pdf assets")
   ymlthis::as_yml(temp_yml)
 }

### 2020-01-08 R/build.R
Update build.R

@@ -16,7 +16,7 @@ build_vertical <- function(clean=TRUE,...) {
   pkgdown::build_site(...)
 
   for (i in list.files(pattern = "\\.Rmd", recursive = TRUE)) {
-    if(dirname(i) != "vignettes"){
+    if(gtools::split_path(i,FALSE)[1] %in% c("experiments","vignettes") == FALSE){
       out.file <- rmarkdown::render(i, output_dir = paste0("docs/", dirname(i)), quiet=TRUE)
       if (!any(grepl(tools::file_path_sans_ext(i), readLines("_pkgdown.yml")))) {
         usethis::ui_info(paste(file.path(dirname(i),basename(out.file)),
@@ -27,7 +27,7 @@ build_vertical <- function(clean=TRUE,...) {
 
   if (dir.exists("experiments")) {
     dir.create("docs/experiments")
-    file.copy("experiments", "docs", recursive = TRUE)
+    #file.copy("experiments", "docs", recursive = TRUE)
   }
 }
 

### 2020-01-08 R/initialize.R
clarify imports and remove license
added new imports
deleted old imports
added new jspsychr remote
commented out mit license

@@ -44,7 +44,7 @@ vertical_project <- function(path = NULL,
   usethis::use_template(template = "_pkgdown.yml",
                         ignore = TRUE,
                         package = "vertical")
-  usethis::use_mit_license()
+ # usethis::use_mit_license()
   if (dots$init_git) {
     git2r::init()
     usethis::use_git_ignore(c(".Rhistory", ".RData", ".Rproj.user"))

### 2020-01-08 R/build.R
reduce dependencies
refactored suggest_yml and build_vertical to remove package dependencies from gtools, ymlthis, and rlist

@@ -16,7 +16,7 @@ build_vertical <- function(clean=TRUE,...) {
   pkgdown::build_site(...)
 
   for (i in list.files(pattern = "\\.Rmd", recursive = TRUE)) {
-    if(gtools::split_path(i,FALSE)[1] %in% c("experiments","vignettes") == FALSE){
+    if(unlist(strsplit(i,split=.Platform$file.sep))[1] %in% c("experiments","vignettes") == FALSE){
       out.file <- rmarkdown::render(i, output_dir = paste0("docs/", dirname(i)), quiet=TRUE)
       if (!any(grepl(tools::file_path_sans_ext(i), readLines("_pkgdown.yml")))) {
         usethis::ui_info(paste(file.path(dirname(i),basename(out.file)),

### 2020-01-08 R/maintain.R

@@ -10,42 +10,40 @@
 #'
 suggest_yml <- function(){
   temp_yml <- yaml::read_yaml("_pkgdown.yml")
-  temp_yml$navbar$components$articles <- NULL
+  exclude_folders <- c("experiments","data","data-raw","docs","R","man","inst")
 
+  # write structure and components
   for (i in list.files(pattern = "\\.Rmd", recursive = TRUE)) {
-    #j <- gsub("vignettes","articles",i)
-    assets <- gtools::split_path(i,FALSE)
+    assets <- unlist(strsplit(i,split=.Platform$file.sep))
+    if(assets[1] %in% exclude_folders == FALSE ){
+      if (assets[1] == "vignettes") assets[1] <- "articles"
+      if(assets[1] %in% temp_yml$navbar$structure$left == FALSE) temp_yml$navbar$structure$left <- c(temp_yml$navbar$structure$left,assets[1])
+      temp_yml$navbar$components[[assets[1]]] <- list(text = assets[1],menu = list())
+    }
+  }
+
+  # write titles and hrefs
+  for (i in list.files(pattern = "\\.Rmd", recursive = TRUE)) {
+    assets <- unlist(strsplit(i,split=.Platform$file.sep))
     j <- assets
-    if(assets[1] %in% c("experiments","data","data-raw","docs","R","man","inst") == FALSE ){
-
-      # handle vignettes exception
-      if (assets[1] == "vignettes") {
-        assets[1] <- "articles"
-        j[1] <- "articles"
-      }
-
-      # write structure to left
-      if(assets[1] %in% temp_yml$navbar$structure$left == FALSE) {
-        temp_yml$navbar$structure$left <- c(temp_yml$navbar$structure$left,assets[1])
-      }
-
-      # write top level component tab text and menu
-      if( "menu" %in% names(temp_yml$navbar$components[[assets[1]]]) == FALSE) {
-        temp_yml$navbar$components[[assets[1]]] <- list(text = assets[1],
-                                                        menu = list())
-      }
-
-      # write titles and hrefs
-      temp_yml$navbar$components[[assets[1]]]$menu <- rlist::list.append(
+    if(assets[1] %in% exclude_folders == FALSE ){
+      if (assets[1] == "vignettes") assets[1] <- "articles"; j[1] <- "articles"
+      temp_yml$navbar$components[[assets[1]]]$menu <- c(
         temp_yml$navbar$components[[assets[1]]]$menu,
[…]

### 2020-01-08 R/initialize.R
fix license problem

@@ -44,7 +44,7 @@ vertical_project <- function(path = NULL,
   usethis::use_template(template = "_pkgdown.yml",
                         ignore = TRUE,
                         package = "vertical")
- # usethis::use_mit_license()
+  if (!is.null(getOption("usethis.full_name"))) usethis::use_mit_license()
   if (dots$init_git) {
     git2r::init()
     usethis::use_git_ignore(c(".Rhistory", ".RData", ".Rproj.user"))

### 2020-01-09 R/initialize.R
remove license on init

@@ -44,7 +44,6 @@ vertical_project <- function(path = NULL,
   usethis::use_template(template = "_pkgdown.yml",
                         ignore = TRUE,
                         package = "vertical")
-  if (!is.null(getOption("usethis.full_name"))) usethis::use_mit_license()
   if (dots$init_git) {
     git2r::init()
     usethis::use_git_ignore(c(".Rhistory", ".RData", ".Rproj.user"))

### 2020-01-13 docs/experiments/Experiment 1/run.R
move /doc/dev to /doc

@@ -1,17 +1,0 @@
-# Run all of this to run the experiment in a browser
-# critical randomization is done by R right now
-# so this should be run to initiate the experiment for each new subject
-# After the experiment is complete, the data should be stored in the data folder
-# when finished, "stop" the server that will be be started by this code
-
-library(xprmntr)
-
-rmarkdown::render("experiments/Experiment 1/experiment/index.Rmd", "html_document")
-
-run_locally(path="experiments/Experiment 1/",
-            show_in = "browser",
-            xprmntr_host = "127.0.0.1",
-            xprmntr_port = 8000)
-
-
-

### 2020-01-13 docs/manuscript/manuscript.tex
include manuscript folder

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  jou,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Sharing and organizing research products as R packages},
  pdflang={en-EN},
  pdfkeywords={reproducibility; research methods; R; open data; open science},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Manuscript styling
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{Research products as R packages}
\author{Matti Vuorre\textsuperscript{1}\ \& Matthew J. C. Crump\textsuperscript{2}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Columbia University, New York, USA\\\textsuperscript{2} Department of Psychology, Brooklyn College of CUNY, New York USA}
\authornote{The authors declare equal authorship.

Correspondence concerning this article should be addressed to Matti Vuorre, 406 Schermerhorn Hall, 1190 Amsterdam Avenue MC 5501, New York, NY 10027. E-mail: mv2521@columbia.edu}
\keywords{reproducibility; research methods; R; open data; open science\newline\indent Word count: 4787}
\usepackage{dblfloatfix}

\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[shorthands=off,main=english]{babel}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {\setlength{\parindent}{0pt}%
  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
  {\par}

\title{Sharing and organizing research products as R packages}

\date{}

\abstract{
A consensus on the importance of open data and reproducible code is emerging. How should data and code be shared to maximize the key desiderata of reproducibility, permanence, and accessibility? Research assets should be stored persistently in formats that are not software restrictive, and documented so that others can reproduce and extend the required computations. The sharing method should be easy to adopt by already busy researchers. We suggest the R package standard as a solution for creating, curating, and communicating research assets. The R package standard, with extensions discussed herein, provides a format for assets and metadata that satisfies the above desiderata, facilitates reproducibility, open access, and sharing of materials through online platforms like GitHub and Open Science Framework. We discuss a stack of R resources that help users create reproducible collections of research assets, from experiments to manuscripts, in the RStudio interface. We created an R package, \emph{vertical}, to help researchers incorporate these tools into their workflows, and discuss its functionality at length in an online supplement. Together, these tools may increase the reproducibility and openness of psychological science.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Research projects produce many assets, such as experiments, data, analyses, manuscripts, posters, slides, stimulus sets and materials, computational models, and more. Although creating and communicating research assets are mainstay activities in science, there is a trend (Vanpaemel, Vermorgen, Deriemaecker, \& Storms, 2015; Wicherts, Borsboom, Kats, \& Molenaar, 2006) encouraging more open and transparent sharing practices (Houtkoop et al., 2018; Lindsay, 2017). However, the value of shared assets is often limited by poor documentation, incompatible file formats, or lack of organization, resulting in low rates of reproducibility (Hardwicke et al., 2018). Some data sharing guidelines and good practices have been suggested (e.g., Rouder, Haaf, \& Snyder, 2019; Rouder, 2016); along with related issues of ethics and privacy, ownership of data, attitudes about sharing, and where to share data (Klein et al., 2018; Martone, Garcia-Castro, \& VandenBos, 2018). Nevertheless, curating standards for research assets have not emerged, possibly due to the variance of research in psychology. Instead of developing another standard, we suggest borrowing existing standards and practices from software engineering. Specifically, the R package standard, with additional R authoring tools, provides a robust framework for organizing and sharing reproducible research products.

Some advances in data-sharing standards have emerged: In much of psychological science it is now customary to share data on Open Science Framework (OSF). However, those materials often contain idiosyncratic file organization and minimal or missing documentation for raw data. In specific areas, organization and documentation standards have emerged, (e.g., the BIDS framework in neuroscience, Gorgolewski et al., 2016), but they usually only consider data and code instead of the project as a whole. More comprehensive proposals are described in the Transparency and Openness Promotion (Nosek et al., 2015), and Peer Reviewers' Openness initiative guidelines (Morey et al., 2016), but these fall short of describing detailed standards for organization and metadata. Additionally, these standards (if they exist) may not be widely known, recognized, agreed upon, and/or adopted at large.

We sought for a standard for organization and sharing that would adhere to the FAIR (Findable, Accessible, Interoperable, Reusable) guidelines to maximize the reuse potential of data and support ``discovery through good data management'' (Wilkinson et al., 2016, see also general discussion). Additionally, we recognized the added value of including other research outputs (``products''; e.g., manuscripts) in a reproducible collection of materials that could be shared with the data. We identified the R package standard as a solution that doesn't present overwhelming overhead for already busy researchers. Here, we discuss the R package standard for creating reproducible research material containers for data, analysis code, and metadata. Then, we introduce a collection of additional R packages for incorporating other research products into the project. These additional tools are easily accessible through our \emph{vertical} package, thus named because an entire research project can be completed from top to bottom as a stack of R processes from within the RStudio software platform.

\hypertarget{r-packages}{%
\section{R Packages}\label{r-packages}}

R is a programming language commonly used for statistical analyses (R Core Team, 2019), and is available at \url{https://www.r-project.org}. RStudio is an integrated development environment (IDE; RStudio Team, 2016), and is available at \url{https://www.rstudio.com/}. Both are free, open source and work on Windows, Mac, and Linux operating systems. We chose these tools because they are already widely used among psychology researchers: In one crowd-sourced analysis, 16 of 29 analysis teams used R (Silberzahn et al., 2018; see also Yee \& Debbie, 2017, and http://r4stats.com/articles/popularity/). Further, many psychologists have already developed R packages that share analysis tools and data.

A cornerstone of R are user-created packages, which contain R functions, data, metadata, and documentation written in a standardized format that allows seamless sharing across users and operating systems. When data and functions are wrapped in R packages, they are immediately available to other R users through the R console, and their documentation can be easily viewed in R or online. Next, we outline how data and functions are included in R packages, and then turn to including other assets, such as manuscripts and posters. We present a brief outline, a complete guide to creating R packages is outside the scope of this manuscript, and we refer readers to Wickham (2015), which remains an authoritative and in-depth guide on creating R packages.

\begin{figure*}

{\centering \includegraphics[width=\textwidth]{images/r-package-skeleton}

}

\caption{Creating an R package using the RStudio graphical user interface.}\label{fig:creating-package-rstudio}
\end{figure*}

Creating a package through RStudio is straightforward (see Figure \ref{fig:creating-package-rstudio}). When a new package is created with the method shown in Figure \ref{fig:creating-package-rstudio}, the required files and directories are automatically created. First, the ``DESCRIPTION'' file includes basic information about the package (e.g., authors, description) in a machine readable format. To edit it, users can click on the file in the RStudio File browser. The ``blank.Rproj'', ``man'' and ``NAMESPACE'' files/directories shown in Figure \ref{fig:creating-package-rstudio} should not be edited by users, and we thus turn to R functions.

\hypertarget{functions}{%
\subsection{Functions}\label{functions}}

Because R packages are primarily intended for R functions, an ``R'' directory is automatically created to store source code for defining functions. Researchers often develop and use custom functions in their analysis. Although it is possible to declare them directly in the analysis scripts, we suggest declaring functions using the R package standard, by placing the function's source code into a file in the ``R'' directory. Functions in R packages are portable, such that others can install the package from their R console, load it, and start using the functions immediately. Packages can also depend on other packages (and be depended on), such that R will automatically include any other packages required for your functions to run appropriately. Functions within R packages are documented in a standardized manner, and the documentation for a function can be viewed within R or online.

Additionally, learning and following R conventions for declaring functions has a pedagogical benefit to the researcher and may improve their practices. There is also a reuse benefit to the researcher: Functions can be difficult to find in old scripts, but easy to find and load if they are already packaged and called from an existing package. Thus, formally including one's functions in R packages facilitates reproducibility and sharing.

\hypertarget{data}{%
\subsection{Data}\label{data}}

Broadly, there are 3 steps to including data in an R package: 1. Placing raw data in the ``data-raw'' directory, 2. creating an R script that processes the raw data and creates an R data object into the ``data'' directory, and 3. documenting the final data object.

First, raw data in any format is stored in the ``data-raw'' directory to ensure that the method of sharing remains software-agnostic. Importantly, the raw data files should never be modified themselves. Second, instructions for converting and cleaning the raw data into analyzable format should also be included in this directory. These pre-processing steps should be declared so that the entire analysis is transparent and reproducible, preferably in an R script in the ``data-raw'' directory. That script should end with creating an R data object in the project's ``data'' directory. Placing the R data object there ensures that it is included in the resulting R package, which makes using the data effortless for any R user, and eschews the need to download additional files. Loading the R package makes the included data immediately available in the R console.

Finally, the resulting data object in ``data'' should be documented in a standardized format by placing a \texttt{data.R} file in the ``R'' directory, a process discussed in more detail in our complete online supplementary tutorial, and in Wickham (2015). That standard format of documentation is especially useful because it allows datasets' documentation to be viewed in R (e.g., type \texttt{?attitude} in the R console) and online (see below).

To recap, R package standards offer a systematic method for creating and curating reproducible research assets. Researchers could use the R package infrastructure to store and document their data and functions, and, as we will see, analyses, manuscripts, posters, and presentations as well. The entire research project is then easily shared online, and showcased as a website, as we will show below. To enable these additional features, many R packages have been developed (e.g., writing manuscripts within R). To facilitate their use, we have created an R package, \emph{vertical}, that, when installed, also installs these other packages, and makes it easier to create projects that use them.

\hypertarget{reproducible-research-projects-with-vertical}{%
\section{\texorpdfstring{Reproducible research projects with \emph{vertical}}{Reproducible research projects with vertical}}\label{reproducible-research-projects-with-vertical}}

\emph{vertical} extends the R package template with a set of additional files and directories for products such as manuscripts and posters. These assets are organized such that they can then be easily shared online on GitHub or OSF, installed to R as an R package, and showcased online as a website. Figure \ref{fig:vertical-workflow} illustrates the workflow of creating a \emph{vertical} research project.

\begin{figure*}

{\centering \includegraphics[width=\textwidth]{images/vertical-workflow}

}

\caption{An overview of the vertical workflow. The project template suggests R Markdown modules for asset creation (manuscripts, slides, posters), that are compiled to multiple formats. Content is curated follow R package standards, then compiled and communicated in the form of an online repository (GitHub/OSF), project website, and R package.}\label{fig:vertical-workflow}
\end{figure*}

Additional benefits of packaging everything together in this manner are that every computation supporting manuscripts, supplementary analyses, data preprocessing, etc. is reproducible and uses the same data in the same computational environment. Because sharing of materials is built-in, no additional work is needed after-the-fact in organizing the materials for sharing them. We begin with an overview of this process, but a more exhaustive step-by-step tutorial is provided online (Crump \& Vuorre, n.d.)\footnote{\url{https://crumplab.github.io/vertical}}. Before using \emph{vertical}, or any packages contained therein, users must first install it in R (the \emph{devtools} (Wickham et al., 2019) package is necessary for installing R packages from GitHub):

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{)}
\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install\_github}\NormalTok{(}\StringTok{"CrumpLab/vertical"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

After installing the package, users should restart RStudio to make all its features available. First, similar to creating an R package as shown above, a new \emph{vertical} research project is created through the RStudio GUI (Figure \ref{fig:vertical-project-structure}.1). Figure \ref{fig:vertical-project-structure}.2 shows the resulting file and directory structure, which is an extension of the R package structure shown in Figure \ref{fig:creating-package-rstudio}. Users may enable or disable components (as indicated by directories like ``experiments'') as needed. Next, we describe the components, and then how to share the resulting research projects and showcase them as websites.

\begin{figure*}

{\centering \includegraphics[width=\textwidth]{images/vertical-project}

}

\caption{Creating a *vertical* project with the RStudio *vertical* project template, and description of the file-organization and structure of a *vertical* project.}\label{fig:vertical-project-structure}
\end{figure*}

In addition to data and functions, almost any kind of content can be included within the project directory structure. Most importantly, to create reproducible assets, such as data analyses or manuscripts, these materials should be created with the R Markdown language. R Markdown is a simple plain text format that allows mixing prose with code (R, Python, etc.), and thus enables including one's text and analyses in a single document. Because of its importance, we describe R Markdown first.

\hypertarget{reproducible-documents-with-r-markdown}{%
\subsection{Reproducible documents with R Markdown}\label{reproducible-documents-with-r-markdown}}

R Markdown is a plain text file format and an R package that allows executable code snippets to be embedded alongside regular text (Xie, Allaire, \& Grolemund, 2018). For example, an R Markdown file can include the typical prose found in manuscripts, the analysis code (R, Python, etc.), and code snippets that display numerical results and figures within the text. R Markdown documents are transparent and reproducible. In principle, recipients can see the analysis scripts, and reproduce them by compiling the document on their own machine. R Markdown documents separate content and style, and can be compiled to multiple output formats, such as PDF, HTML, and Microsoft Word. Importantly, R Markdown is basically plain text, and thus effortless to learn. Nevertheless, a tutorial on R Markdown is beyond our scope (see Xie et al., 2018)\footnote{See also \url{https://rmarkdown.rstudio.com/authoring_quick_tour.html} for a quick overview of R Markdown.}.

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

Data analysis can be made fully transparent and reproducible with R Markdown. Above, we suggested placing any preprocessing steps in a separate file in the ``data-raw'' directory, and creating a documented R data object in ``data''. If the data object is called \texttt{mydata}, once the package is installed and loaded, simply calling \texttt{mydata} in an R script will call the data object, ready to be used in analyses. Our favored approach is to write any analyses relevant to a product (e.g., manuscript or a poster) directly into the product's R Markdown source file. Typically, we would include data analysis code in the R Markdown source of the APA style manuscript (see below).

Nevertheless, sometimes supplementary analyses are conducted that are not included in any other document. These are most conveniently placed in R Markdown files into the ``vignettes'' directory. Then, when the project is compiled into a website (see below), those analyses are viewable as part of the resulting website. When a \emph{vertical} project is created, the ``vignettes'' directory is automatically created with an example data analysis R Markdown script.

\hypertarget{apa-manuscripts}{%
\subsection{APA manuscripts}\label{apa-manuscripts}}

Reproducible APA formatted manuscripts can easily be created with the \emph{papaja} R package, that provides an R Markdown template file, and additional helper functions for creating tables and figures (Aust \& Barth, 2018). For example, the R Markdown file can contain the text and code to generate the manuscript and all results. \emph{papaja} provides several functions for reporting results, especially from common statistical analyses such as ANOVAs, t-tests, and regressions. For example, \emph{papaja}'s \texttt{apa\_print()} function embeds statistical results directly into the manuscript, obviating the need to copy them by hand and removing human error in reporting. References can be automatically populated from Zotero using the \texttt{citr} plugin (Aust, 2019). We refer readers to the \emph{papaja} documentation website for more information.\footnote{\url{https://crsh.github.io/papaja_man/introduction.html}}

When a new \emph{vertical} project is created, a \emph{papaja} manuscript template is automatically included in the ``manuscript'' directory. Then, once the project is compiled, the manuscript PDF is copied to the ``docs'' folder, from where it can be viewed as part of the resulting website (see below).

\hypertarget{posters}{%
\subsection{Posters}\label{posters}}

\emph{vertical} includes a posters folder where poster documents can be deposited and shared on the website. Additionally, \emph{vertical} suggests using the \emph{posterdown} package (Thorne, 2019) for poster creation through an R Markdown document. When the provided template is used, the resulting reproducible document is viewable as part of the website.

\hypertarget{slides}{%
\subsection{Slides}\label{slides}}

\emph{vertical} includes a slides directory for slide decks in any format, which can then be included for downloading or viewing on the resulting website. There are several R Markdown options for creating reproducible slide decks in various formats. When a new \emph{vertical} project is created, a \emph{slidy} template is included to create web-ready presentations.\footnote{\url{https://bookdown.org/yihui/rmarkdown/slidy-presentation.html}} Slidy presentations are viewable in a browser like typical PowerPoint or keynote presentations, but have some additional features like long scroll-able slides, interactive figures and animations, and rendering full websites within individual slides.

\hypertarget{other-materials}{%
\subsection{Other materials}\label{other-materials}}

Other research assets may include supplementary materials, stimulus sets, code for a computational model, additional analyses, or other documents, like a public-facing blog. \emph{vertical} is flexible with respect to the inclusion of additional assets not specified by the template. For example, arbitrary R Markdown documents can be included in the vignettes directory. When the project is compiled (see below), any R Markdown files in that directory are compiled and served on the website.

\hypertarget{experiments}{%
\subsection{Experiments}\label{experiments}}

Finally, \emph{vertical} projects can include experiments in the experiments directory. These materials could be packaged as a compressed .zip file, which would then be accessible from an online repository. \emph{vertical} suggests using \emph{jsPsych} (De Leeuw, 2015), a JavaScript library for building browser-based experiments that can be served to the web or run locally. Selecting the \emph{jsPsych} module during project initialization downloads the most recent version of \emph{jsPsych} to the experiments folder. A basic html template is also provided, which loads the jsPsych library and presents some basic word stimuli. To further edit the experiment, users can simply edit the html file as they would when creating any other jsPsych experiment. When the \emph{vertical} project is compiled, this experiment is immediately available from the website navigation bar.

For users who wish to work entirely within R Markdown, we suggest \emph{jspsychr} (Crump, 2019),\footnote{\url{https://crumplab.github.io/jspsychr/}} an R package that helps creating jsPsych experiments with R Markdown in RStudio. This option allows using RStudio's ability to pass objects between R and JavaScript when constructing experiment materials. The template creates a sample \emph{jsPsych} experiment produced by \emph{jspsychr}, which contains the R Markdown script to generate the experiment file, the html file to run the experiment, and a folder to save the data and document the experiment.

Browser-friendly experiments have several benefits over local-only experiments. First, scripts are open-source and easily reusable (no installation is required to run an experiment). Second, scripts can be shared in a self-contained manner such that the experiment is reproducible and the methods are transparent and verifiable. Third, because the experiment is an html file it can be shared and run as part of the project website, facilitating understanding of the experimental procedures. Sharing experiment source code with a demonstration is not yet common, but could benefit the review process and readers' understanding of the procedures by making methods more transparent.

\hypertarget{sharing-and-communicating}{%
\section{Sharing and communicating}\label{sharing-and-communicating}}

Above, we discussed using R Markdown, R package standards, and other R resources for creating research products. Then next step is sharing and communicating these products. When the products are created as suggested above, \emph{vertical} facilitates sharing the entire project as an R package for easy access to the data and functions, as a version controlled repository for access to the complete source code for reproducing every component of the project, and as a website, where the components are easily viewed by others.

\begin{figure*}

{\centering \includegraphics[width=\textwidth]{images/vertical-website}

}

\caption{Research assets from a \emph{vertical} project displayed on a website. The navigation tab bar on the top links to the content pages depicted in the body of the webpage.}\label{fig:vertical-project-website}
\end{figure*}

\hypertarget{sharing-the-source-code-with-version-control}{%
\subsection{Sharing the source code with version control}\label{sharing-the-source-code-with-version-control}}

An organizing principle of our approach is that all project materials are stored in a single directory on the user's local filesystem. When version control, such as Git, is enabled in this directory, the entire project can be seamlessly shared with others through online platforms such as GitHub and GitLab. Version control systems provide filesystems with additional functioning for keeping track of files' versions, improved options for collaborating, and seamless integration with online collaboration platforms (GitHub and GitLab); however, the details are outside the scope of this manuscript and we refer readers to Vuorre \& Curley (2018). By default, \emph{vertical} projects enable Git for new projects. Further, projects developed with Git (on a sharing platform such as GitHub or GitLab) can be seamlessly integrated with the Open Science Framework for long-term curation and, for example, to enable digital object identifiers (DOI).

\hypertarget{sharing-the-project-as-a-website}{%
\subsection{Sharing the project as a website}\label{sharing-the-project-as-a-website}}

An additional benefit of the R package organizing principles, and using the \emph{vertical} helper tools, is that the entire project can be showcased as a website (see Figure \ref{fig:vertical-project-website}. The website is compiled by \emph{pkgdown} (Wickham \& Hesselberth, 2019) in one command (\texttt{vertical::build\_vertical()}), or by clicking Addins -\textgreater{} Build Vertical in RStudio, and expertise with web standards is not required. For instructions on customizing the resulting website, see the \emph{pkgdown} instructions (\url{https://usethis.r-lib.org/}).

Building a \emph{vertical} project as shown above creates the website in the docs folder, from where it can be viewed. During this process, all the R Markdown files within the project are evaluated while the documents are compiled, thus ensuring that all analyses, and thus products, are up to date. Importantly, the website can be easily made available on the internet. To do so, users should integrate their project with a GitHub repository (Vuorre \& Curley, 2018), and then enable GitHub pages under the repository's settings. An example of the resulting website can be viewed at \url{https://mvuorre.github.io/rparp/}.

\hypertarget{sharing-as-an-r-package}{%
\subsection{Sharing as an R Package}\label{sharing-as-an-r-package}}

If the project contains R functions, data, and documentation following the R package standard (as discussed above), then others can install the package on their computers for fast access to the data and functions from within R. Again, integrating the project with GitHub is beneficial, because then others can install the package from GitHub using the \emph{devtools} package in R (\texttt{devtools::install\_github("account/repository")}). Sharing as an R package is very convenient for other R users, as any packaged data and functions are immediately available for use, natively in R, after package installation. We expand on the extended benefits of using R packages for transparency and reproducibility in the general discussion.

\hypertarget{general-discussion}{%
\section{General Discussion}\label{general-discussion}}

We presented \emph{vertical}, a stack of R tools for a single-platform solution for creating, curating, and communicating psychological research projects in a reproducible manner. It might seem that the benefits of this approach apply only to R users. We believe this is not the case; the workflow creates a single package that includes all the research products related to a product, which can then be viewed in a public (or private, if users so choose) website and source code repository. Of course, to get the most out of the proposal, some R use is required (from both the creator and subsequent users.) However, any standardized computational procedure assumes some computational environment, our choice of R is natural due to its already widespread adoption in the psychological researcher community. Furthermore, by requiring raw data inclusion, our proposal does not strictly require using R. And, the software we have discussed, RStudio, works well with other languages, including JavaScript, Python, and C++.

\hypertarget{validating-research-assets}{%
\subsection{Validating research assets}\label{validating-research-assets}}

Our suggested approach has wider implications for making research practices in psychology more transparent and reproducible. Consider the implied truth claims associated with any peer-reviewed manuscript. The act of publishing can be seen as a validity assertion: that the truth claims made by the authors are verifiably true, or more weakly that the research assets are verifiable. Unlike a set of conclusions from premises that may be tested for logical validity, conclusions and claims from a research project are often not as easily scrutinized and tested. Reviewers extend trust that methods and results were as written when assets can not be verified directly. Sharing research assets in the manner discussed here would greatly facilitate scrutinizing and testing all components of the research project, and thus potentially increase the reliability of the resulting scientific products.

Furthermore, organizing research assets in the manner discussed here allow its components to be formally tested, a process referred to as unit testing in software engineering. It refers to the practice of verifying that functions perform as intended in a variety of cases. We make a loose analogy between building a \emph{vertical} project and unit testing. Building is a call to compile individual components of the project. If a user adheres to the \emph{vertical} workflow, then R must be able to perform all of the scripted operations. Successfully building a \emph{vertical} project is loosely analogous to performing unit tests that confirm and validate the process of asset creation. So, compiling a vertical project allows a researcher to unit test their own research project and verify its assets. By sharing a \emph{vertical} project, other researchers can perform the same verification. We note that \emph{vertical} does not conduct proper unit tests on functions included in an R package. However, because these projects are R packages, proper unit tests can be easily included (see Wickham, 2011).

\hypertarget{collaborating-with-vertical}{%
\subsection{Collaborating with vertical}\label{collaborating-with-vertical}}

Software developers routinely use version control platforms, such as GitHub and GitLab to coordinate software development across a team of collaborators (see, Vuorre \& Curley, 2018). We recognize that scientific research is \emph{not} software development. However, some components of the research workflow readily conform to the software development model---e.g., developing experimental programs, data analysis---and as such stand to benefit from borrowing best practices and tools from that domain.

Research projects can be hosted on GitHub throughout the lifespan of the project to make use of tools for collaboration. With this in mind, \emph{vertical} strongly suggests using Git (along with its online platforms). For example, we used GitHub as a collaboration tool to create \emph{vertical} and to write this manuscript (using \emph{papaja}). Among other features, Git provides functionality similar to track changes in Microsoft Word that allow contributions to be reviewed, accepted, and modified. It makes this process robust even when many collaborators are involved. We also made extensive use of the GitHub issues tab, which is normally used to report bugs or feature requests to a software developer. Many users may find this feature very useful. In our case, we opened project related issues for discussion and then closed the issues once they were addressed. A side-effect is that the GitHub issues tab, and the Git history, may double as a project development narrative, or a lab journaling system preserving issues and solutions for posterity.

\hypertarget{teaching-r-with-vertical}{%
\subsection{Teaching R with vertical}\label{teaching-r-with-vertical}}

\emph{vertical} may be most beneficial to users who are already well-versed in R; and, although the website tutorial is intended for R beginners, it is not a tutorial on the R language. Nevertheless, \emph{vertical} could be used for pedagogical purposes, like a museum guide offering an orientation to collections in a large museum. For example, an R programming class for psychologists (or statistics/methods course using R) could be structured around the components of a vertical project, such as how to program functions, import and analyse data, write manuscripts, posters, slide decks, and websites in R. Moreover, \emph{vertical} provides a convenient and well-organized workflow for students to save their work as they learn each component.

\hypertarget{extending-vertical}{%
\subsection{Extending vertical}\label{extending-vertical}}

\emph{vertical} is a modular open-source tool that can be extended by anyone. We added content modules that we think are common in many psychology research projects, and can incorporate new modules as they are identified and suggested by a user base. Users can fork the \emph{vertical} source code on GitHub, make suggested changes, and submit pull requests for review and inclusion.

\hypertarget{fair}{%
\subsection{FAIR}\label{fair}}

To maximize the reuse potential of open data and support ``discovery through good data management'', Wilkinson et al. (2016) proposed the FAIR (Findable, Accessible, Interoperable, Reusable) guiding principles. Below, we describe how a \emph{vertical} workflow targets FAIR principles and facilitates accessibility and data reanalyses (FAIR principles paraphrased from, Wilkinson et al., 2016; and Martone et al., 2018)

\textbf{F}indable data are discoverable through persistent identifiers (e.g., permanent URLs / DOIs):

\begin{itemize}
\tightlist
\item
  Projects are accessible through the internet at OSF or GitHub
\item
  DOIs for a \emph{vertical} project can be minted through OSF or Zenodo
\end{itemize}

\textbf{A}ccessible data is available to approved researchers (e.g., anyone or lab members only) through a standardized communications protocol (e.g., Internet):

\begin{itemize}
\tightlist
\item
  Projects on GitHub and OSF can be made accessible to a specific set of users, or completely private, if desired
\item
  Most data sharing guides emphasize manually uploading heterogeneous data files to a repository (e.g., OSF). R packages allow that, but also facilitate programmatic access to the data, and therefore are more accessible than methods relying on manual transfer of files
\end{itemize}

\textbf{I}nteroperable data is described (via metadata and file organization) in a common, openly available, non-proprietary language:

\begin{itemize}
\tightlist
\item
  R packages describe a strict organization of source files, resulting in a software product that is directly usable in a programming environment
\item
  R packages describe a common language for documenting the data, its use, and dependencies
\item
  R package (meta)data is human- and machine-readable and in a non-proprietary format
\item
  Data wrapped in an R package is accessible in other programming environments due to standardized formatting
\end{itemize}

\textbf{R}eusable data are richly described and licensed to facilitate reuse:

\begin{itemize}
\tightlist
\item
  R packages are described in a standard format
\item
  R packages facilitate use of appropriate licenses
\end{itemize}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Researchers may improve the reproducibility of their work by borrowing standards from software development, such as the R package standard, discussed here for sharing research products. By organizing research assets in a standardized format, their reuse will be less time consuming and error-prone. Standardized formats also facilitate large-scale meta-scientific investigations (Gorgolewski et al., 2016), and inclusion of data sets for meta-analyses.

More generally, in addition to the convenience of a single-platform \emph{vertical} workflow, adopting R standards during asset creation has the side effect of demanding some amount of attention on how research products are curated, stored, documented, and shared. Drawing attention to these details may serve to increase the reproducibility of our work.

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-R-citr}{}%
Aust, F. (2019). \emph{Citr: 'RStudio' add-in to insert markdown citations}. Retrieved from \url{https://CRAN.R-project.org/package=citr}

\leavevmode\hypertarget{ref-R-papaja}{}%
Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}

\leavevmode\hypertarget{ref-R-jspsychr}{}%
Crump, M. J. C. (2019). \emph{Jspsychr: Templates and functions for writing and running jspsych experiments from r-studio}. Retrieved from \url{https://github.com/CrumpLab/jspsychr}

\leavevmode\hypertarget{ref-R-vertical}{}%
Crump, M. J. C., \& Vuorre, M. (n.d.). \emph{Vertical: Reproducible worfklow for psychological science research asset creation and communication}. Retrieved from \url{https://github.com/CrumpLab/vertical}

\leavevmode\hypertarget{ref-de2015jspsych}{}%
De Leeuw, J. R. (2015). JsPsych: A javascript library for creating behavioral experiments in a web browser. \emph{Behavior Research Methods}, \emph{47}(1), 1--12.

\leavevmode\hypertarget{ref-gorgolewski_brain_2016}{}%
Gorgolewski, K. J., Auer, T., Calhoun, V. D., Craddock, R. C., Das, S., Duff, E. P., \ldots{} Poldrack, R. A. (2016). The brain imaging data structure, a format for organizing and describing outputs of neuroimaging experiments. \emph{Scientific Data}, \emph{3}, 160044. \url{https://doi.org/10.1038/sdata.2016.44}

\leavevmode\hypertarget{ref-HardwickeDataavailabilityreusability2018}{}%
Hardwicke, T. E., Mathur, M. B., MacDonald, K., Nilsonne, G., Banks, G. C., Kidwell, M. C., \ldots{} Frank, M. C. (2018). Data availability, reusability, and analytic reproducibility: Evaluating the impact of a mandatory open data policy at the journal Cognition. \emph{Royal Society Open Science}, \emph{5}(8), 180448. \url{https://doi.org/10.1098/rsos.180448}

\leavevmode\hypertarget{ref-HoutkoopDataSharingPsychology2018}{}%
Houtkoop, B. L., Chambers, C., Macleod, M., Bishop, D. V. M., Nichols, T. E., \& Wagenmakers, E.-J. (2018). Data Sharing in Psychology: A Survey on Barriers and Preconditions. \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(1), 70--85. \url{https://doi.org/10.1177/2515245917751886}

\leavevmode\hypertarget{ref-KleinPracticalGuideTransparency2018}{}%
Klein, O., Hardwicke, T. E., Aust, F., Breuer, J., Danielsson, H., Mohr, A. H., \ldots{} Frank, M. C. (2018). A Practical Guide for Transparency in Psychological Science. \emph{Collabra: Psychology}, \emph{4}(1), 20. \url{https://doi.org/10.1525/collabra.158}

\leavevmode\hypertarget{ref-LindsaySharingDataMaterials2017}{}%
Lindsay, D. S. (2017). Sharing Data and Materials in Psychological Science. \emph{Psychological Science}, \emph{28}(6), 699--702. \url{https://doi.org/10.1177/0956797617704015}

\leavevmode\hypertarget{ref-MartoneDatasharingpsychology2018}{}%
Martone, M. E., Garcia-Castro, A., \& VandenBos, G. R. (2018). Data sharing in psychology. \emph{American Psychologist}, \emph{73}(2), 111--125. \url{https://doi.org/10.1037/amp0000242}

\leavevmode\hypertarget{ref-morey_peer_2016}{}%
Morey, R. D., Chambers, C. D., Etchells, P. J., Harris, C. R., Hoekstra, R., Lakens, D., \ldots{} Zwaan, R. A. (2016). The Peer Reviewers Openness Initiative: Incentivizing open research practices through peer review. \emph{Royal Society Open Science}, \emph{3}(1), 150547. \url{https://doi.org/10.1098/rsos.150547}

\leavevmode\hypertarget{ref-nosek_promoting_2015}{}%
Nosek, B. A., Alter, G., Banks, G. C., Borsboom, D., Bowman, S. D., Breckler, S. J., \ldots{} Yarkoni, T. (2015). Promoting an open research culture. \emph{Science}, \emph{348}(6242), 1422--1425. \url{https://doi.org/10.1126/science.aab2374}

\leavevmode\hypertarget{ref-R-base}{}%
R Core Team. (2019). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}

\leavevmode\hypertarget{ref-Rouderwhatwhyhow2016}{}%
Rouder, J. N. (2016). The what, why, and how of born-open data. \emph{Behavior Research Methods}, \emph{48}(3), 1062--1069. \url{https://doi.org/10.3758/s13428-015-0630-z}

\leavevmode\hypertarget{ref-RouderMinimizingMistakesPsychological2019}{}%
Rouder, J. N., Haaf, J. M., \& Snyder, H. K. (2019). Minimizing Mistakes in Psychological Science. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(1), 3--11. \url{https://doi.org/10.1177/2515245918801915}

\leavevmode\hypertarget{ref-RStudioTeamRStudioIntegratedDevelopment2016}{}%
RStudio Team. (2016). \emph{RStudio: Integrated Development Environment for R}. Boston, MA: RStudio, Inc. Retrieved from \url{http://www.rstudio.com/}

\leavevmode\hypertarget{ref-SilberzahnManyAnalystsOne2018}{}%
Silberzahn, R., Uhlmann, E. L., Martin, D. P., Anselmi, P., Aust, F., Awtrey, E., \ldots{} Nosek, B. A. (2018). Many Analysts, One Data Set: Making Transparent How Variations in Analytic Choices Affect Results. \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(3), 337--356. \url{https://doi.org/10.1177/2515245917747646}

\leavevmode\hypertarget{ref-R-posterdown}{}%
Thorne, W. B. (2019). \emph{Posterdown: An r package built to generate reproducible conference posters for the academic and professional world where powerpoint and pages just won't cut it}. Retrieved from \url{https://github.com/brentthorne/posterdown}

\leavevmode\hypertarget{ref-VanpaemelAreWeWasting2015}{}%
Vanpaemel, W., Vermorgen, M., Deriemaecker, L., \& Storms, G. (2015). Are We Wasting a Good Crisis? The Availability of Psychological Research Data after the Storm. \emph{Collabra: Psychology}, \emph{1}(1), Art. 3. \url{https://doi.org/10.1525/collabra.13}

\leavevmode\hypertarget{ref-VuorreCuratingResearchAssets2018}{}%
Vuorre, M., \& Curley, J. P. (2018). Curating Research Assets: A Tutorial on the Git Version Control System. \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 219--236. \url{https://doi.org/10.1177/2515245918754826}

\leavevmode\hypertarget{ref-Wichertspooravailabilitypsychological2006}{}%
Wicherts, J. M., Borsboom, D., Kats, J., \& Molenaar, D. (2006). The poor availability of psychological research data for reanalysis. \emph{American Psychologist}, \emph{61}(7), 726--728. \url{https://doi.org/10.1037/0003-066X.61.7.726}

\leavevmode\hypertarget{ref-R-testthat}{}%
Wickham, H. (2011). Testthat: Get started with testing. \emph{The R Journal}, \emph{3}, 5--10. Retrieved from \url{https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf}

\leavevmode\hypertarget{ref-WickhamPackagesOrganizeTest2015}{}%
Wickham, H. (2015). \emph{R Packages: Organize, Test, Document, and Share Your Code}. "O'Reilly Media, Inc.". Retrieved from \url{http://r-pkgs.had.co.nz/}

\leavevmode\hypertarget{ref-R-pkgdown}{}%
Wickham, H., \& Hesselberth, J. (2019). \emph{Pkgdown: Make static html documentation for a package}. Retrieved from \url{https://CRAN.R-project.org/package=pkgdown}

\leavevmode\hypertarget{ref-R-devtools}{}%
Wickham, H., Hester, J., \& Chang, W. (2019). \emph{Devtools: Tools to make developing r packages easier}. Retrieved from \url{https://CRAN.R-project.org/package=devtools}

\leavevmode\hypertarget{ref-WilkinsonFAIRGuidingPrinciples2016}{}%
Wilkinson, M. D., Dumontier, M., Aalbersberg, I. J., Appleton, G., Axton, M., Baak, A., \ldots{} Mons, B. (2016). The FAIR Guiding Principles for scientific data management and stewardship. \emph{Scientific Data}, \emph{3}, 160018. \url{https://doi.org/10.1038/sdata.2016.18}

\leavevmode\hypertarget{ref-R-rmarkdown}{}%
Xie, Y., Allaire, J. J., \& Grolemund, G. (2018). \emph{R markdown: The definitive guide}. Boca Raton, Florida: Chapman; Hall/CRC. Retrieved from \url{https://bookdown.org/yihui/rmarkdown}

\leavevmode\hypertarget{ref-yee_why_2017}{}%
Yee, S. J. W., \& Debbie. (2017). Why You Should Become a UseR: A Brief Introduction to R. \emph{APS Observer}, \emph{30}(3). Retrieved from \url{https://www.psychologicalscience.org/observer/why-you-should-become-a-user-a-brief-introduction-to-r}

\leavevmode\hypertarget{ref-R-citr}{}%
Aust, F. (2019). \emph{Citr: 'RStudio' add-in to insert markdown citations}. Retrieved from \url{https://CRAN.R-project.org/package=citr}

\leavevmode\hypertarget{ref-R-papaja}{}%
Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}

\leavevmode\hypertarget{ref-R-jspsychr}{}%
Crump, M. J. C. (2019). \emph{Jspsychr: Templates and functions for writing and running jspsych experiments from r-studio}. Retrieved from \url{https://github.com/CrumpLab/jspsychr}

\leavevmode\hypertarget{ref-R-vertical}{}%
Crump, M. J. C., \& Vuorre, M. (n.d.). \emph{Vertical: Reproducible worfklow for psychological science research asset creation and communication}. Retrieved from \url{https://github.com/CrumpLab/vertical}

\leavevmode\hypertarget{ref-de2015jspsych}{}%
De Leeuw, J. R. (2015). JsPsych: A javascript library for creating behavioral experiments in a web browser. \emph{Behavior Research Methods}, \emph{47}(1), 1--12.

\leavevmode\hypertarget{ref-gorgolewski_brain_2016}{}%
Gorgolewski, K. J., Auer, T., Calhoun, V. D., Craddock, R. C., Das, S., Duff, E. P., \ldots{} Poldrack, R. A. (2016). The brain imaging data structure, a format for organizing and describing outputs of neuroimaging experiments. \emph{Scientific Data}, \emph{3}, 160044. \url{https://doi.org/10.1038/sdata.2016.44}

\leavevmode\hypertarget{ref-HardwickeDataavailabilityreusability2018}{}%
Hardwicke, T. E., Mathur, M. B., MacDonald, K., Nilsonne, G., Banks, G. C., Kidwell, M. C., \ldots{} Frank, M. C. (2018). Data availability, reusability, and analytic reproducibility: Evaluating the impact of a mandatory open data policy at the journal Cognition. \emph{Royal Society Open Science}, \emph{5}(8), 180448. \url{https://doi.org/10.1098/rsos.180448}

\leavevmode\hypertarget{ref-HoutkoopDataSharingPsychology2018}{}%
Houtkoop, B. L., Chambers, C., Macleod, M., Bishop, D. V. M., Nichols, T. E., \& Wagenmakers, E.-J. (2018). Data Sharing in Psychology: A Survey on Barriers and Preconditions. \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(1), 70--85. \url{https://doi.org/10.1177/2515245917751886}

\leavevmode\hypertarget{ref-KleinPracticalGuideTransparency2018}{}%
Klein, O., Hardwicke, T. E., Aust, F., Breuer, J., Danielsson, H., Mohr, A. H., \ldots{} Frank, M. C. (2018). A Practical Guide for Transparency in Psychological Science. \emph{Collabra: Psychology}, \emph{4}(1), 20. \url{https://doi.org/10.1525/collabra.158}

\leavevmode\hypertarget{ref-LindsaySharingDataMaterials2017}{}%
Lindsay, D. S. (2017). Sharing Data and Materials in Psychological Science. \emph{Psychological Science}, \emph{28}(6), 699--702. \url{https://doi.org/10.1177/0956797617704015}

\leavevmode\hypertarget{ref-MartoneDatasharingpsychology2018}{}%
Martone, M. E., Garcia-Castro, A., \& VandenBos, G. R. (2018). Data sharing in psychology. \emph{American Psychologist}, \emph{73}(2), 111--125. \url{https://doi.org/10.1037/amp0000242}

\leavevmode\hypertarget{ref-morey_peer_2016}{}%
Morey, R. D., Chambers, C. D., Etchells, P. J., Harris, C. R., Hoekstra, R., Lakens, D., \ldots{} Zwaan, R. A. (2016). The Peer Reviewers Openness Initiative: Incentivizing open research practices through peer review. \emph{Royal Society Open Science}, \emph{3}(1), 150547. \url{https://doi.org/10.1098/rsos.150547}

\leavevmode\hypertarget{ref-nosek_promoting_2015}{}%
Nosek, B. A., Alter, G., Banks, G. C., Borsboom, D., Bowman, S. D., Breckler, S. J., \ldots{} Yarkoni, T. (2015). Promoting an open research culture. \emph{Science}, \emph{348}(6242), 1422--1425. \url{https://doi.org/10.1126/science.aab2374}

\leavevmode\hypertarget{ref-R-base}{}%
R Core Team. (2019). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}

\leavevmode\hypertarget{ref-Rouderwhatwhyhow2016}{}%
Rouder, J. N. (2016). The what, why, and how of born-open data. \emph{Behavior Research Methods}, \emph{48}(3), 1062--1069. \url{https://doi.org/10.3758/s13428-015-0630-z}

\leavevmode\hypertarget{ref-RouderMinimizingMistakesPsychological2019}{}%
Rouder, J. N., Haaf, J. M., \& Snyder, H. K. (2019). Minimizing Mistakes in Psychological Science. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(1), 3--11. \url{https://doi.org/10.1177/2515245918801915}

\leavevmode\hypertarget{ref-RStudioTeamRStudioIntegratedDevelopment2016}{}%
RStudio Team. (2016). \emph{RStudio: Integrated Development Environment for R}. Boston, MA: RStudio, Inc. Retrieved from \url{http://www.rstudio.com/}

\leavevmode\hypertarget{ref-SilberzahnManyAnalystsOne2018}{}%
Silberzahn, R., Uhlmann, E. L., Martin, D. P., Anselmi, P., Aust, F., Awtrey, E., \ldots{} Nosek, B. A. (2018). Many Analysts, One Data Set: Making Transparent How Variations in Analytic Choices Affect Results. \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(3), 337--356. \url{https://doi.org/10.1177/2515245917747646}

\leavevmode\hypertarget{ref-R-posterdown}{}%
Thorne, W. B. (2019). \emph{Posterdown: An r package built to generate reproducible conference posters for the academic and professional world where powerpoint and pages just won't cut it}. Retrieved from \url{https://github.com/brentthorne/posterdown}

\leavevmode\hypertarget{ref-VanpaemelAreWeWasting2015}{}%
Vanpaemel, W., Vermorgen, M., Deriemaecker, L., \& Storms, G. (2015). Are We Wasting a Good Crisis? The Availability of Psychological Research Data after the Storm. \emph{Collabra: Psychology}, \emph{1}(1), Art. 3. \url{https://doi.org/10.1525/collabra.13}

\leavevmode\hypertarget{ref-VuorreCuratingResearchAssets2018}{}%
Vuorre, M., \& Curley, J. P. (2018). Curating Research Assets: A Tutorial on the Git Version Control System. \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 219--236. \url{https://doi.org/10.1177/2515245918754826}

\leavevmode\hypertarget{ref-Wichertspooravailabilitypsychological2006}{}%
Wicherts, J. M., Borsboom, D., Kats, J., \& Molenaar, D. (2006). The poor availability of psychological research data for reanalysis. \emph{American Psychologist}, \emph{61}(7), 726--728. \url{https://doi.org/10.1037/0003-066X.61.7.726}

\leavevmode\hypertarget{ref-R-testthat}{}%
Wickham, H. (2011). Testthat: Get started with testing. \emph{The R Journal}, \emph{3}, 5--10. Retrieved from \url{https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf}

\leavevmode\hypertarget{ref-WickhamPackagesOrganizeTest2015}{}%
Wickham, H. (2015). \emph{R Packages: Organize, Test, Document, and Share Your Code}. "O'Reilly Media, Inc.". Retrieved from \url{http://r-pkgs.had.co.nz/}

\leavevmode\hypertarget{ref-R-pkgdown}{}%
Wickham, H., \& Hesselberth, J. (2019). \emph{Pkgdown: Make static html documentation for a package}. Retrieved from \url{https://CRAN.R-project.org/package=pkgdown}

\leavevmode\hypertarget{ref-R-devtools}{}%
Wickham, H., Hester, J., \& Chang, W. (2019). \emph{Devtools: Tools to make developing r packages easier}. Retrieved from \url{https://CRAN.R-project.org/package=devtools}

\leavevmode\hypertarget{ref-WilkinsonFAIRGuidingPrinciples2016}{}%
Wilkinson, M. D., Dumontier, M., Aalbersberg, I. J., Appleton, G., Axton, M., Baak, A., \ldots{} Mons, B. (2016). The FAIR Guiding Principles for scientific data management and stewardship. \emph{Scientific Data}, \emph{3}, 160018. \url{https://doi.org/10.1038/sdata.2016.18}

\leavevmode\hypertarget{ref-R-rmarkdown}{}%
Xie, Y., Allaire, J. J., \& Grolemund, G. (2018). \emph{R markdown: The definitive guide}. Boca Raton, Florida: Chapman; Hall/CRC. Retrieved from \url{https://bookdown.org/yihui/rmarkdown}

\leavevmode\hypertarget{ref-yee_why_2017}{}%
Yee, S. J. W., \& Debbie. (2017). Why You Should Become a UseR: A Brief Introduction to R. \emph{APS Observer}, \emph{30}(3). Retrieved from \url{https://www.psychologicalscience.org/observer/why-you-should-become-a-user-a-brief-introduction-to-r}
\end{cslreferences}

\endgroup

\end{document}

### 2020-01-13 manuscript/manuscript.Rmd
include manuscript folder

--- 2020-01-13 docs/manuscript/manuscript.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-%
-\documentclass[
-  english,
-  jou,floatsintext]{apa6}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={Sharing and organizing research products as R packages},
-  pdflang={en-EN},
-  pdfkeywords={reproducibility; research methods; R; open data; open science},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-% Manuscript styling
-\usepackage{csquotes}
-\usepackage{upgreek}
-\captionsetup{font=singlespacing,justification=justified}
-
-% Table formatting
-\usepackage{longtable}
-\usepackage{lscape}
-% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
-\usepackage{multirow}		% Table styling
-\usepackage{tabularx}		% Control Column width
-\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
-\usepackage{threeparttablex}            % Lets threeparttable work with longtable
-
-% Create new environments so endfloat can handle them
-% \newenvironment{ltable}
-%   {\begin{landscape}\begin{center}\begin{threeparttable}}
-%   {\end{threeparttable}\end{center}\end{landscape}}
-\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
-
-% Enables adjusting longtable caption width to table width
-% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
-\makeatletter
-\newcommand\LastLTentrywidth{1em}
-\newlength\longtablewidth
-\setlength{\longtablewidth}{1in}
-\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
-
-% \setlength{\parindent}{0.5in}
-% \setlength{\parskip}{0pt plus 0pt minus 0pt}
-
-% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
-% See https://github.com/crsh/papaja/issues/292
-\makeatletter
-\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
-  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
-  {-1em}%
-  {\normalfont\normalsize\bfseries\itshape\typesectitle}}
-
-\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
-  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
-  {-\z@\relax}%
-  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
-\makeatother
-
-% \usepackage{etoolbox}
-\makeatletter
-\patchcmd{\HyOrg@maketitle}
-  {\section{\normalfont\normalsize\abstractname}}
-  {\section*{\normalfont\normalsize\abstractname}}
-  {}{\typeout{Failed to patch abstract.}}
-\makeatother
-\shorttitle{Research products as R packages}
-\author{Matti Vuorre\textsuperscript{1}\ \& Matthew J. C. Crump\textsuperscript{2}}
-\affiliation{
-\vspace{0.5cm}
-\textsuperscript{1} Department of Psychology, Columbia University, New York, USA\\\textsuperscript{2} Department of Psychology, Brooklyn College of CUNY, New York USA}
-\authornote{The authors declare equal authorship.
-
-Correspondence concerning this article should be addressed to Matti Vuorre, 406 Schermerhorn Hall, 1190 Amsterdam Avenue MC 5501, New York, NY 10027. E-mail: mv2521@columbia.edu}
-\keywords{reproducibility; research methods; R; open data; open science\newline\indent Word count: 4787}
-\usepackage{dblfloatfix}
-
-\ifxetex
-  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
-\else
-  \usepackage[shorthands=off,main=english]{babel}
-\fi
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newenvironment{cslreferences}%
-  {\setlength{\parindent}{0pt}%
-  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
-  {\par}
-
-\title{Sharing and organizing research products as R packages}
-
-\date{}
-
-\abstract{
+---
+title: "Sharing and organizing research products as R packages"
+shorttitle: Research products as R packages
+author: 
+  - name: Matti Vuorre
+    affiliation: "1"
+    corresponding: yes
+    address: 406 Schermerhorn Hall, 1190 Amsterdam Avenue MC 5501, New York, NY 10027
+    email: mv2521@columbia.edu
+  - name: Matthew J. C. Crump
+    affiliation: "2"
+affiliation:
+  - id: "1"
+    institution: Department of Psychology, Columbia University, New York, USA
+  - id: "2"
+    institution: Department of Psychology, Brooklyn College of CUNY, New York USA
+authornote: |
+  The authors declare equal authorship.
+abstract: |
   A consensus on the importance of open data and reproducible code is emerging.
@@
 We discuss a stack of R resources that help users create reproducible collections of research assets, from experiments to manuscripts, in the RStudio interface.
-We created an R package, \emph{vertical}, to help researchers incorporate these tools into their workflows, and discuss its functionality at length in an online supplement.
+We created an R package, *vertical*, to help researchers incorporate these tools into their workflows, and discuss its functionality at length in an online supplement.
 Together, these tools may increase the reproducibility and openness of psychological science.
-}
+keywords: "reproducibility; research methods; R; open data; open science"
+wordcount: "`r wordcountaddin::word_count('manuscript.Rmd')`"
+bibliography: ["references.bib","r-references.bib"]
+floatsintext: yes
+figurelist: no
+tablelist: no
+footnotelist: no
+linenumbers: no
+mask: no
+draft: no
+documentclass: apa6
+classoption: jou
+output: papaja::apa6_pdf
+pkgdown:
+  as_is: true
+  extension: pdf
+---
 
-\begin{document}
-\maketitle
+```{r include = FALSE}
+library(knitr)
+library(papaja)
+opts_chunk$set(fig.align = "center")
+```
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+# Introduction
 
 Research projects produce many assets, such as experiments, data, analyses, manuscripts, posters, slides, stimulus sets and materials, computational models, and more.
-Although creating and communicating research assets are mainstay activities in science, there is a trend (Vanpaemel, Vermorgen, Deriemaecker, \& Storms, 2015; Wicherts, Borsboom, Kats, \& Molenaar, 2006) encouraging more open and transparent sharing practices (Houtkoop et al., 2018; Lindsay, 2017).
-However, the value of shared assets is often limited by poor documentation, incompatible file formats, or lack of organization, resulting in low rates of reproducibility (Hardwicke et al., 2018).
-Some data sharing guidelines and good practices have been suggested (e.g., Rouder, Haaf, \& Snyder, 2019; Rouder, 2016); along with related issues of ethics and privacy, ownership of data, attitudes about sharing, and where to share data (Klein et al., 2018; Martone, Garcia-Castro, \& VandenBos, 2018).
+Although creating and communicating research assets are mainstay activities in science, there is a trend [@VanpaemelAreWeWasting2015; @Wichertspooravailabilitypsychological2006] encouraging more open and transparent sharing practices [@HoutkoopDataSharingPsychology2018; @LindsaySharingDataMaterials2017].
+However, the value of shared assets is often limited by poor documentation, incompatible file formats, or lack of organization, resulting in low rates of reproducibility [@HardwickeDataavailabilityreusability2018].
+Some data sharing guidelines and good practices have been suggested [e.g., @RouderMinimizingMistakesPsychological2019; @Rouderwhatwhyhow2016]; along with related issues of ethics and privacy, ownership of data, attitudes about sharing, and where to share data [@KleinPracticalGuideTransparency2018; @MartoneDatasharingpsychology2018].
 Nevertheless, curating standards for research assets have not emerged, possibly due to the variance of research in psychology.
@@
 However, those materials often contain idiosyncratic file organization and minimal or missing documentation for raw data.
-In specific areas, organization and documentation standards have emerged, (e.g., the BIDS framework in neuroscience, Gorgolewski et al., 2016), but they usually only consider data and code instead of the project as a whole.
-More comprehensive proposals are described in the Transparency and Openness Promotion (Nosek et al., 2015), and Peer Reviewers' Openness initiative guidelines (Morey et al., 2016), but these fall short of describing detailed standards for organization and metadata.
+In specific areas, organization and documentation standards have emerged, [e.g., the BIDS framework in neuroscience, @gorgolewski_brain_2016], but they usually only consider data and code instead of the project as a whole.
+More comprehensive proposals are described in the Transparency and Openness Promotion [@nosek_promoting_2015], and Peer Reviewers' Openness initiative guidelines [@morey_peer_2016], but these fall short of describing detailed standards for organization and metadata.
 Additionally, these standards (if they exist) may not be widely known, recognized, agreed upon, and/or adopted at large.
 
-We sought for a standard for organization and sharing that would adhere to the FAIR (Findable, Accessible, Interoperable, Reusable) guidelines to maximize the reuse potential of data and support ``discovery through good data management'' (Wilkinson et al., 2016, see also general discussion).
-Additionally, we recognized the added value of including other research outputs (``products''; e.g., manuscripts) in a reproducible collection of materials that could be shared with the data.
+We sought for a standard for organization and sharing that would adhere to the FAIR (Findable, Accessible, Interoperable, Reusable) guidelines to maximize the reuse potential of data and support "discovery through good data management" [@WilkinsonFAIRGuidingPrinciples2016, see also general discussion].
+Additionally, we recognized the added value of including other research outputs ("products"; e.g., manuscripts) in a reproducible collection of materials that could be shared with the data.
 We identified the R package standard as a solution that doesn't present overwhelming overhead for already busy researchers.
@@
 Then, we introduce a collection of additional R packages for incorporating other research products into the project.
-These additional tools are easily accessible through our \emph{vertical} package, thus named because an entire research project can be completed from top to bottom as a stack of R processes from within the RStudio software platform.
+These additional tools are easily accessible through our *vertical* package, thus named because an entire research project can be completed from top to bottom as a stack of R processes from within the RStudio software platform.
 
-\hypertarget{r-packages}{%
-\section{R Packages}\label{r-packages}}
+# R Packages
 
-R is a programming language commonly used for statistical analyses (R Core Team, 2019), and is available at \url{https://www.r-project.org}.
-RStudio is an integrated development environment (IDE; RStudio Team, 2016), and is available at \url{https://www.rstudio.com/}.
+R is a programming language commonly used for statistical analyses [@R-base], and is available at <https://www.r-project.org>.
+RStudio is an integrated development environment [IDE; @RStudioTeamRStudioIntegratedDevelopment2016], and is available at <https://www.rstudio.com/>.
 Both are free, open source and work on Windows, Mac, and Linux operating systems.
-We chose these tools because they are already widely used among psychology researchers: In one crowd-sourced analysis, 16 of 29 analysis teams used R (Silberzahn et al., 2018; see also Yee \& Debbie, 2017, and http://r4stats.com/articles/popularity/).
+We chose these tools because they are already widely used among psychology researchers: In one crowd-sourced analysis, 16 of 29 analysis teams used R [@SilberzahnManyAnalystsOne2018; see also @yee_why_2017 and <http://r4stats.com/articles/popularity/>].
 Further, many psychologists have already developed R packages that share analysis tools and data.
@@
 Next, we outline how data and functions are included in R packages, and then turn to including other assets, such as manuscripts and posters.
-We present a brief outline, a complete guide to creating R packages is outside the scope of this manuscript, and we refer readers to Wickham (2015), which remains an authoritative and in-depth guide on creating R packages.
+We present a brief outline, a complete guide to creating R packages is outside the scope of this manuscript, and we refer readers to Wickham [-@WickhamPackagesOrganizeTest2015], which remains an authoritative and in-depth guide on creating R packages.
 
-\begin{figure*}
+```{r creating-package-rstudio, echo = F, eval = T, fig.cap = "Creating an R package using the RStudio graphical user interface.", out.width="\\textwidth", fig.env="figure*"}
+include_graphics("images/r-package-skeleton.png", dpi = 300)
+```
 
-{\centering \includegraphics[width=\textwidth]{images/r-package-skeleton}
+Creating a package through RStudio is straightforward (see Figure \@ref(fig:creating-package-rstudio)).
+When a new package is created with the method shown in Figure \@ref(fig:creating-package-rstudio), the required files and directories are automatically created.
+First, the "DESCRIPTION" file includes basic information about the package (e.g., authors, description) in a machine readable format.
+To edit it, users can click on the file in the RStudio File browser.
+The "blank.Rproj", "man" and "NAMESPACE" files/directories shown in Figure \@ref(fig:creating-package-rstudio) should not be edited by users, and we thus turn to R functions.
 
-}
+## Functions
 
-\caption{Creating an R package using the RStudio graphical user interface.}\label{fig:creating-package-rstudio}
-\end{figure*}
-
-Creating a package through RStudio is straightforward (see Figure \ref{fig:creating-package-rstudio}).
-When a new package is created with the method shown in Figure \ref{fig:creating-package-rstudio}, the required files and directories are automatically created.
-First, the ``DESCRIPTION'' file includes basic information about the package (e.g., authors, description) in a machine readable format.
-To edit it, users can click on the file in the RStudio File browser.
-The ``blank.Rproj'', ``man'' and ``NAMESPACE'' files/directories shown in Figure \ref{fig:creating-package-rstudio} should not be edited by users, and we thus turn to R functions.
-
-\hypertarget{functions}{%
-\subsection{Functions}\label{functions}}
-
-Because R packages are primarily intended for R functions, an ``R'' directory is automatically created to store source code for defining functions.
+Because R packages are primarily intended for R functions, an "R" directory is automatically created to store source code for defining functions.
 Researchers often develop and use custom functions in their analysis.
-Although it is possible to declare them directly in the analysis scripts, we suggest declaring functions using the R package standard, by placing the function's source code into a file in the ``R'' directory.
+Although it is possible to declare them directly in the analysis scripts, we suggest declaring functions using the R package standard, by placing the function's source code into a file in the "R" directory.
 Functions in R packages are portable, such that others can install the package from their R console, load it, and start using the functions immediately.
@@
 
-\hypertarget{data}{%
-\subsection{Data}\label{data}}
+## Data
 
 Broadly, there are 3 steps to including data in an R package: 1.
-Placing raw data in the ``data-raw'' directory, 2.
-creating an R script that processes the raw data and creates an R data object into the ``data'' directory, and 3.
+Placing raw data in the "data-raw" directory, 2.
+creating an R script that processes the raw data and creates an R data object into the "data" directory, and 3.
 documenting the final data object.
 
-First, raw data in any format is stored in the ``data-raw'' directory to ensure that the method of sharing remains software-agnostic.
+First, raw data in any format is stored in the "data-raw" directory to ensure that the method of sharing remains software-agnostic.
 Importantly, the raw data files should never be modified themselves.
 Second, instructions for converting and cleaning the raw data into analyzable format should also be included in this directory.
-These pre-processing steps should be declared so that the entire analysis is transparent and reproducible, preferably in an R script in the ``data-raw'' directory.
-That script should end with creating an R data object in the project's ``data'' directory.
+These pre-processing steps should be declared so that the entire analysis is transparent and reproducible, preferably in an R script in the "data-raw" directory.
+That script should end with creating an R data object in the project's "data" directory.
 Placing the R data object there ensures that it is included in the resulting R package, which makes using the data effortless for any R user, and eschews the need to download additional files.
@@
 
-Finally, the resulting data object in ``data'' should be documented in a standardized format by placing a \texttt{data.R} file in the ``R'' directory, a process discussed in more detail in our complete online supplementary tutorial, and in Wickham (2015).
-That standard format of documentation is especially useful because it allows datasets' documentation to be viewed in R (e.g., type \texttt{?attitude} in the R console) and online (see below).
+Finally, the resulting data object in "data" should be documented in a standardized format by placing a `data.R` file in the "R" directory, a process discussed in more detail in our complete online supplementary tutorial, and in @WickhamPackagesOrganizeTest2015.
+That standard format of documentation is especially useful because it allows datasets' documentation to be viewed in R (e.g., type `?attitude` in the R console) and online (see below).
 
@@
 To enable these additional features, many R packages have been developed (e.g., writing manuscripts within R).
-To facilitate their use, we have created an R package, \emph{vertical}, that, when installed, also installs these other packages, and makes it easier to create projects that use them.
+To facilitate their use, we have created an R package, *vertical*, that, when installed, also installs these other packages, and makes it easier to create projects that use them.
 
-\hypertarget{reproducible-research-projects-with-vertical}{%
-\section{\texorpdfstring{Reproducible research projects with \emph{vertical}}{Reproducible research projects with vertical}}\label{reproducible-research-projects-with-vertical}}
+# Reproducible research projects with *vertical*
 
-\emph{vertical} extends the R package template with a set of additional files and directories for products such as manuscripts and posters.
+*vertical* extends the R package template with a set of additional files and directories for products such as manuscripts and posters.
 These assets are organized such that they can then be easily shared online on GitHub or OSF, installed to R as an R package, and showcased online as a website.
-Figure \ref{fig:vertical-workflow} illustrates the workflow of creating a \emph{vertical} research project.
+Figure \@ref(fig:vertical-workflow) illustrates the workflow of creating a *vertical* research project.
 
-\begin{figure*}
-
-{\centering \includegraphics[width=\textwidth]{images/vertical-workflow}
-
-}
-
-\caption{An overview of the vertical workflow.
+```{r vertical-workflow, echo = F, eval = T, fig.cap = "An overview of the vertical workflow.
 The project template suggests R Markdown modules for asset creation (manuscripts, slides, posters), that are compiled to multiple formats.
-Content is curated follow R package standards, then compiled and communicated in the form of an online repository (GitHub/OSF), project website, and R package.}\label{fig:vertical-workflow}
-\end{figure*}
+Content is curated follow R package standards, then compiled and communicated in the form of an online repository (GitHub/OSF), project website, and R package.", out.width="\\textwidth", fig.env="figure*"}
+include_graphics("images/vertical-workflow.png", dpi = 300)
+```
 
@@
 Because sharing of materials is built-in, no additional work is needed after-the-fact in organizing the materials for sharing them.
-We begin with an overview of this process, but a more exhaustive step-by-step tutorial is provided online (Crump \& Vuorre, n.d.)\footnote{\url{https://crumplab.github.io/vertical}}.
-Before using \emph{vertical}, or any packages contained therein, users must first install it in R (the \emph{devtools} (Wickham et al., 2019) package is necessary for installing R packages from GitHub):
+We begin with an overview of this process, but a more exhaustive step-by-step tutorial is provided online [@R-vertical][^v1].
+Before using *vertical*, or any packages contained therein, users must first install it in R (the *devtools* [@R-devtools] package is necessary for installing R packages from GitHub):
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{)}
-\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install\_github}\NormalTok{(}\StringTok{"CrumpLab/vertical"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+```{r, eval=FALSE, echo=TRUE}
+install.packages("devtools")
+devtools::install_github("CrumpLab/vertical")
+```
+
+[^v1]: <https://crumplab.github.io/vertical>
 
 After installing the package, users should restart RStudio to make all its features available.
-First, similar to creating an R package as shown above, a new \emph{vertical} research project is created through the RStudio GUI (Figure \ref{fig:vertical-project-structure}.1).
-Figure \ref{fig:vertical-project-structure}.2 shows the resulting file and directory structure, which is an extension of the R package structure shown in Figure \ref{fig:creating-package-rstudio}.
-Users may enable or disable components (as indicated by directories like ``experiments'') as needed.
+First, similar to creating an R package as shown above, a new *vertical* research project is created through the RStudio GUI (Figure \@ref(fig:vertical-project-structure).1).
+Figure \@ref(fig:vertical-project-structure).2 shows the resulting file and directory structure, which is an extension of the R package structure shown in Figure \@ref(fig:creating-package-rstudio).
+Users may enable or disable components (as indicated by directories like "experiments") as needed.
 Next, we describe the components, and then how to share the resulting research projects and showcase them as websites.
 
-\begin{figure*}
-
-{\centering \includegraphics[width=\textwidth]{images/vertical-project}
-
-}
-
-\caption{Creating a *vertical* project with the RStudio *vertical* project template, and description of the file-organization and structure of a *vertical* project.}\label{fig:vertical-project-structure}
-\end{figure*}
+```{r vertical-project-structure, echo = F, eval = T, fig.cap = "Creating a *vertical* project with the RStudio *vertical* project template, and description of the file-organization and structure of a *vertical* project.", out.width="\\textwidth", fig.env="figure*"}
+include_graphics("images/vertical-project.png", dpi = 300)
+```
 
@@
 
-\hypertarget{reproducible-documents-with-r-markdown}{%
-\subsection{Reproducible documents with R Markdown}\label{reproducible-documents-with-r-markdown}}
+## Reproducible documents with R Markdown
 
-R Markdown is a plain text file format and an R package that allows executable code snippets to be embedded alongside regular text (Xie, Allaire, \& Grolemund, 2018).
+R Markdown is a plain text file format and an R package that allows executable code snippets to be embedded alongside regular text [@R-rmarkdown].
 For example, an R Markdown file can include the typical prose found in manuscripts, the analysis code (R, Python, etc.), and code snippets that display numerical results and figures within the text.
@@
 Importantly, R Markdown is basically plain text, and thus effortless to learn.
-Nevertheless, a tutorial on R Markdown is beyond our scope (see Xie et al., 2018)\footnote{See also \url{https://rmarkdown.rstudio.com/authoring_quick_tour.html} for a quick overview of R Markdown.}.
+Nevertheless, a tutorial on R Markdown is beyond our scope [see @R-rmarkdown][^rmark].
 
-\hypertarget{data-analysis}{%
-\subsection{Data analysis}\label{data-analysis}}
+[^rmark]: See also <https://rmarkdown.rstudio.com/authoring_quick_tour.html> for a quick overview of R Markdown.
+
+## Data analysis
 
 Data analysis can be made fully transparent and reproducible with R Markdown.
-Above, we suggested placing any preprocessing steps in a separate file in the ``data-raw'' directory, and creating a documented R data object in ``data''.
-If the data object is called \texttt{mydata}, once the package is installed and loaded, simply calling \texttt{mydata} in an R script will call the data object, ready to be used in analyses.
+Above, we suggested placing any preprocessing steps in a separate file in the "data-raw" directory, and creating a documented R data object in "data".
+If the data object is called `mydata`, once the package is installed and loaded, simply calling `mydata` in an R script will call the data object, ready to be used in analyses.
 Our favored approach is to write any analyses relevant to a product (e.g., manuscript or a poster) directly into the product's R Markdown source file.
@@
 Nevertheless, sometimes supplementary analyses are conducted that are not included in any other document.
-These are most conveniently placed in R Markdown files into the ``vignettes'' directory.
+These are most conveniently placed in R Markdown files into the "vignettes" directory.
 Then, when the project is compiled into a website (see below), those analyses are viewable as part of the resulting website.
-When a \emph{vertical} project is created, the ``vignettes'' directory is automatically created with an example data analysis R Markdown script.
+When a *vertical* project is created, the "vignettes" directory is automatically created with an example data analysis R Markdown script.
 
-\hypertarget{apa-manuscripts}{%
-\subsection{APA manuscripts}\label{apa-manuscripts}}
+## APA manuscripts
 
-Reproducible APA formatted manuscripts can easily be created with the \emph{papaja} R package, that provides an R Markdown template file, and additional helper functions for creating tables and figures (Aust \& Barth, 2018).
+Reproducible APA formatted manuscripts can easily be created with the *papaja* R package, that provides an R Markdown template file, and additional helper functions for creating tables and figures [@R-papaja].
 For example, the R Markdown file can contain the text and code to generate the manuscript and all results.
-\emph{papaja} provides several functions for reporting results, especially from common statistical analyses such as ANOVAs, t-tests, and regressions.
-For example, \emph{papaja}'s \texttt{apa\_print()} function embeds statistical results directly into the manuscript, obviating the need to copy them by hand and removing human error in reporting.
-References can be automatically populated from Zotero using the \texttt{citr} plugin (Aust, 2019).
-We refer readers to the \emph{papaja} documentation website for more information.\footnote{\url{https://crsh.github.io/papaja_man/introduction.html}}
+*papaja* provides several functions for reporting results, especially from common statistical analyses such as ANOVAs, t-tests, and regressions.
+For example, *papaja*'s `apa_print()` function embeds statistical results directly into the manuscript, obviating the need to copy them by hand and removing human error in reporting.
+References can be automatically populated from Zotero using the `citr` plugin [@R-citr].
+We refer readers to the *papaja* documentation website for more information.[^ppj]
 
-When a new \emph{vertical} project is created, a \emph{papaja} manuscript template is automatically included in the ``manuscript'' directory.
-Then, once the project is compiled, the manuscript PDF is copied to the ``docs'' folder, from where it can be viewed as part of the resulting website (see below).
+[^ppj]: <https://crsh.github.io/papaja_man/introduction.html>
 
-\hypertarget{posters}{%
-\subsection{Posters}\label{posters}}
+When a new *vertical* project is created, a *papaja* manuscript template is automatically included in the "manuscript" directory.
+Then, once the project is compiled, the manuscript PDF is copied to the "docs" folder, from where it can be viewed as part of the resulting website (see below).
 
-\emph{vertical} includes a posters folder where poster documents can be deposited and shared on the website.
-Additionally, \emph{vertical} suggests using the \emph{posterdown} package (Thorne, 2019) for poster creation through an R Markdown document.
+## Posters
+
+*vertical* includes a posters folder where poster documents can be deposited and shared on the website.
+Additionally, *vertical* suggests using the *posterdown* package [@R-posterdown] for poster creation through an R Markdown document.
 When the provided template is used, the resulting reproducible document is viewable as part of the website.
 
-\hypertarget{slides}{%
-\subsection{Slides}\label{slides}}
+## Slides
 
-\emph{vertical} includes a slides directory for slide decks in any format, which can then be included for downloading or viewing on the resulting website.
+*vertical* includes a slides directory for slide decks in any format, which can then be included for downloading or viewing on the resulting website.
 There are several R Markdown options for creating reproducible slide decks in various formats.
-When a new \emph{vertical} project is created, a \emph{slidy} template is included to create web-ready presentations.\footnote{\url{https://bookdown.org/yihui/rmarkdown/slidy-presentation.html}} Slidy presentations are viewable in a browser like typical PowerPoint or keynote presentations, but have some additional features like long scroll-able slides, interactive figures and animations, and rendering full websites within individual slides.
+When a new *vertical* project is created, a *slidy* template is included to create web-ready presentations [^slidy].
+Slidy presentations are viewable in a browser like typical PowerPoint or keynote presentations, but have some additional features like long scroll-able slides, interactive figures and animations, and rendering full websites within individual slides.
 
-\hypertarget{other-materials}{%
-\subsection{Other materials}\label{other-materials}}
+[^slidy]: <https://bookdown.org/yihui/rmarkdown/slidy-presentation.html>
+
+## Other materials
 
 Other research assets may include supplementary materials, stimulus sets, code for a computational model, additional analyses, or other documents, like a public-facing blog.
-\emph{vertical} is flexible with respect to the inclusion of additional assets not specified by the template.
+*vertical* is flexible with respect to the inclusion of additional assets not specified by the template.
 For example, arbitrary R Markdown documents can be included in the vignettes directory.
@@
 
-\hypertarget{experiments}{%
-\subsection{Experiments}\label{experiments}}
+## Experiments
 
-Finally, \emph{vertical} projects can include experiments in the experiments directory.
+Finally, *vertical* projects can include experiments in the experiments directory.
 These materials could be packaged as a compressed .zip file, which would then be accessible from an online repository.
-\emph{vertical} suggests using \emph{jsPsych} (De Leeuw, 2015), a JavaScript library for building browser-based experiments that can be served to the web or run locally.
-Selecting the \emph{jsPsych} module during project initialization downloads the most recent version of \emph{jsPsych} to the experiments folder.
+*vertical* suggests using *jsPsych* [@de2015jspsych], a JavaScript library for building browser-based experiments that can be served to the web or run locally.
+Selecting the *jsPsych* module during project initialization downloads the most recent version of *jsPsych* to the experiments folder.
 A basic html template is also provided, which loads the jsPsych library and presents some basic word stimuli.
 To further edit the experiment, users can simply edit the html file as they would when creating any other jsPsych experiment.
-When the \emph{vertical} project is compiled, this experiment is immediately available from the website navigation bar.
+When the *vertical* project is compiled, this experiment is immediately available from the website navigation bar.
 
-For users who wish to work entirely within R Markdown, we suggest \emph{jspsychr} (Crump, 2019),\footnote{\url{https://crumplab.github.io/jspsychr/}} an R package that helps creating jsPsych experiments with R Markdown in RStudio.
+For users who wish to work entirely within R Markdown, we suggest *jspsychr* [@R-jspsychr] [^jspsychr], an R package that helps creating jsPsych experiments with R Markdown in RStudio.
 This option allows using RStudio's ability to pass objects between R and JavaScript when constructing experiment materials.
-The template creates a sample \emph{jsPsych} experiment produced by \emph{jspsychr}, which contains the R Markdown script to generate the experiment file, the html file to run the experiment, and a folder to save the data and document the experiment.
+The template creates a sample *jsPsych* experiment produced by *jspsychr*, which contains the R Markdown script to generate the experiment file, the html file to run the experiment, and a folder to save the data and document the experiment.
+
+[^jspsychr]: <https://crumplab.github.io/jspsychr/>
 
@@
 
-\hypertarget{sharing-and-communicating}{%
-\section{Sharing and communicating}\label{sharing-and-communicating}}
+# Sharing and communicating
 
@@
 Then next step is sharing and communicating these products.
-When the products are created as suggested above, \emph{vertical} facilitates sharing the entire project as an R package for easy access to the data and functions, as a version controlled repository for access to the complete source code for reproducing every component of the project, and as a website, where the components are easily viewed by others.
+When the products are created as suggested above, *vertical* facilitates sharing the entire project as an R package for easy access to the data and functions, as a version controlled repository for access to the complete source code for reproducing every component of the project, and as a website, where the components are easily viewed by others.
 
-\begin{figure*}
+```{r vertical-project-website, echo = F, eval = T, fig.cap = "Research assets from a \\emph{vertical} project displayed on a website.
+The navigation tab bar on the top links to the content pages depicted in the body of the webpage.", out.width="\\textwidth", fig.env="figure*"}
+include_graphics("images/vertical-website.png", dpi = 300)
+```
 
-{\centering \includegraphics[width=\textwidth]{images/vertical-website}
-
-}
-
-\caption{Research assets from a \emph{vertical} project displayed on a website.
-The navigation tab bar on the top links to the content pages depicted in the body of the webpage.}\label{fig:vertical-project-website}
-\end{figure*}
-
-\hypertarget{sharing-the-source-code-with-version-control}{%
-\subsection{Sharing the source code with version control}\label{sharing-the-source-code-with-version-control}}
+## Sharing the source code with version control
 
@@
 When version control, such as Git, is enabled in this directory, the entire project can be seamlessly shared with others through online platforms such as GitHub and GitLab.
-Version control systems provide filesystems with additional functioning for keeping track of files' versions, improved options for collaborating, and seamless integration with online collaboration platforms (GitHub and GitLab); however, the details are outside the scope of this manuscript and we refer readers to Vuorre \& Curley (2018).
-By default, \emph{vertical} projects enable Git for new projects.
+Version control systems provide filesystems with additional functioning for keeping track of files' versions, improved options for collaborating, and seamless integration with online collaboration platforms (GitHub and GitLab); however, the details are outside the scope of this manuscript and we refer readers to Vuorre & Curley [-@VuorreCuratingResearchAssets2018].
+By default, *vertical* projects enable Git for new projects.
 Further, projects developed with Git (on a sharing platform such as GitHub or GitLab) can be seamlessly integrated with the Open Science Framework for long-term curation and, for example, to enable digital object identifiers (DOI).
 
-\hypertarget{sharing-the-project-as-a-website}{%
-\subsection{Sharing the project as a website}\label{sharing-the-project-as-a-website}}
+## Sharing the project as a website
 
-An additional benefit of the R package organizing principles, and using the \emph{vertical} helper tools, is that the entire project can be showcased as a website (see Figure \ref{fig:vertical-project-website}.
-The website is compiled by \emph{pkgdown} (Wickham \& Hesselberth, 2019) in one command (\texttt{vertical::build\_vertical()}), or by clicking Addins -\textgreater{} Build Vertical in RStudio, and expertise with web standards is not required.
-For instructions on customizing the resulting website, see the \emph{pkgdown} instructions (\url{https://usethis.r-lib.org/}).
+An additional benefit of the R package organizing principles, and using the *vertical* helper tools, is that the entire project can be showcased as a website (see Figure \@ref(fig:vertical-project-website).
+The website is compiled by *pkgdown* [@R-pkgdown] in one command (`vertical::build_vertical()`), or by clicking Addins -> Build Vertical in RStudio, and expertise with web standards is not required.
+For instructions on customizing the resulting website, see the *pkgdown* instructions (<https://usethis.r-lib.org/>).
 
-Building a \emph{vertical} project as shown above creates the website in the docs folder, from where it can be viewed.
+Building a *vertical* project as shown above creates the website in the docs folder, from where it can be viewed.
 During this process, all the R Markdown files within the project are evaluated while the documents are compiled, thus ensuring that all analyses, and thus products, are up to date.
 Importantly, the website can be easily made available on the internet.
-To do so, users should integrate their project with a GitHub repository (Vuorre \& Curley, 2018), and then enable GitHub pages under the repository's settings.
-An example of the resulting website can be viewed at \url{https://mvuorre.github.io/rparp/}.
+To do so, users should integrate their project with a GitHub repository [@VuorreCuratingResearchAssets2018], and then enable GitHub pages under the repository's settings.
+An example of the resulting website can be viewed at <https://mvuorre.github.io/rparp/>.
 
-\hypertarget{sharing-as-an-r-package}{%
-\subsection{Sharing as an R Package}\label{sharing-as-an-r-package}}
+## Sharing as an R Package
 
 If the project contains R functions, data, and documentation following the R package standard (as discussed above), then others can install the package on their computers for fast access to the data and functions from within R.
-Again, integrating the project with GitHub is beneficial, because then others can install the package from GitHub using the \emph{devtools} package in R (\texttt{devtools::install\_github("account/repository")}).
+Again, integrating the project with GitHub is beneficial, because then others can install the package from GitHub using the *devtools* package in R (`devtools::install_github("account/repository")`).
 Sharing as an R package is very convenient for other R users, as any packaged data and functions are immediately available for use, natively in R, after package installation.
@@
 
-\hypertarget{general-discussion}{%
-\section{General Discussion}\label{general-discussion}}
+# General Discussion
 
-We presented \emph{vertical}, a stack of R tools for a single-platform solution for creating, curating, and communicating psychological research projects in a reproducible manner.
+We presented *vertical*, a stack of R tools for a single-platform solution for creating, curating, and communicating psychological research projects in a reproducible manner.
 It might seem that the benefits of this approach apply only to R users.
@@
 
-\hypertarget{validating-research-assets}{%
-\subsection{Validating research assets}\label{validating-research-assets}}
+## Validating research assets
 
@@
 It refers to the practice of verifying that functions perform as intended in a variety of cases.
-We make a loose analogy between building a \emph{vertical} project and unit testing.
+We make a loose analogy between building a *vertical* project and unit testing.
 Building is a call to compile individual components of the project.
-If a user adheres to the \emph{vertical} workflow, then R must be able to perform all of the scripted operations.
-Successfully building a \emph{vertical} project is loosely analogous to performing unit tests that confirm and validate the process of asset creation.
+If a user adheres to the *vertical* workflow, then R must be able to perform all of the scripted operations.
+Successfully building a *vertical* project is loosely analogous to performing unit tests that confirm and validate the process of asset creation.
 So, compiling a vertical project allows a researcher to unit test their own research project and verify its assets.
-By sharing a \emph{vertical} project, other researchers can perform the same verification.
-We note that \emph{vertical} does not conduct proper unit tests on functions included in an R package.
-However, because these projects are R packages, proper unit tests can be easily included (see Wickham, 2011).
+By sharing a *vertical* project, other researchers can perform the same verification.
+We note that *vertical* does not conduct proper unit tests on functions included in an R package.
+However, because these projects are R packages, proper unit tests can be easily included [see @R-testthat].
 
-\hypertarget{collaborating-with-vertical}{%
-\subsection{Collaborating with vertical}\label{collaborating-with-vertical}}
+## Collaborating with vertical
 
-Software developers routinely use version control platforms, such as GitHub and GitLab to coordinate software development across a team of collaborators (see, Vuorre \& Curley, 2018).
-We recognize that scientific research is \emph{not} software development.
+Software developers routinely use version control platforms, such as GitHub and GitLab to coordinate software development across a team of collaborators [see, @VuorreCuratingResearchAssets2018].
+We recognize that scientific research is *not* software development.
 However, some components of the research workflow readily conform to the software development model---e.g., developing experimental programs, data analysis---and as such stand to benefit from borrowing best practices and tools from that domain.
@@
 Research projects can be hosted on GitHub throughout the lifespan of the project to make use of tools for collaboration.
-With this in mind, \emph{vertical} strongly suggests using Git (along with its online platforms).
-For example, we used GitHub as a collaboration tool to create \emph{vertical} and to write this manuscript (using \emph{papaja}).
+With this in mind, *vertical* strongly suggests using Git (along with its online platforms).
+For example, we used GitHub as a collaboration tool to create *vertical* and to write this manuscript (using *papaja*).
 Among other features, Git provides functionality similar to track changes in Microsoft Word that allow contributions to be reviewed, accepted, and modified.
@@
 
-\hypertarget{teaching-r-with-vertical}{%
-\subsection{Teaching R with vertical}\label{teaching-r-with-vertical}}
+## Teaching R with vertical
 
-\emph{vertical} may be most beneficial to users who are already well-versed in R; and, although the website tutorial is intended for R beginners, it is not a tutorial on the R language.
-Nevertheless, \emph{vertical} could be used for pedagogical purposes, like a museum guide offering an orientation to collections in a large museum.
+*vertical* may be most beneficial to users who are already well-versed in R; and, although the website tutorial is intended for R beginners, it is not a tutorial on the R language.
+Nevertheless, *vertical* could be used for pedagogical purposes, like a museum guide offering an orientation to collections in a large museum.
 For example, an R programming class for psychologists (or statistics/methods course using R) could be structured around the components of a vertical project, such as how to program functions, import and analyse data, write manuscripts, posters, slide decks, and websites in R.
-Moreover, \emph{vertical} provides a convenient and well-organized workflow for students to save their work as they learn each component.
+Moreover, *vertical* provides a convenient and well-organized workflow for students to save their work as they learn each component.
 
-\hypertarget{extending-vertical}{%
-\subsection{Extending vertical}\label{extending-vertical}}
+## Extending vertical
 
-\emph{vertical} is a modular open-source tool that can be extended by anyone.
+*vertical* is a modular open-source tool that can be extended by anyone.
 We added content modules that we think are common in many psychology research projects, and can incorporate new modules as they are identified and suggested by a user base.
-Users can fork the \emph{vertical} source code on GitHub, make suggested changes, and submit pull requests for review and inclusion.
+Users can fork the *vertical* source code on GitHub, make suggested changes, and submit pull requests for review and inclusion.
 
-\hypertarget{fair}{%
-\subsection{FAIR}\label{fair}}
+## FAIR
 
-To maximize the reuse potential of open data and support ``discovery through good data management'', Wilkinson et al.
-(2016) proposed the FAIR (Findable, Accessible, Interoperable, Reusable) guiding principles.
-Below, we describe how a \emph{vertical} workflow targets FAIR principles and facilitates accessibility and data reanalyses (FAIR principles paraphrased from, Wilkinson et al., 2016; and Martone et al., 2018)
+To maximize the reuse potential of open data and support "discovery through good data management", @WilkinsonFAIRGuidingPrinciples2016 proposed the FAIR (Findable, Accessible, Interoperable, Reusable) guiding principles.
+Below, we describe how a *vertical* workflow targets FAIR principles and facilitates accessibility and data reanalyses [FAIR principles paraphrased from, @WilkinsonFAIRGuidingPrinciples2016; and @MartoneDatasharingpsychology2018]
 
-\textbf{F}indable data are discoverable through persistent identifiers (e.g., permanent URLs / DOIs):
+**F**indable data are discoverable through persistent identifiers (e.g., permanent URLs / DOIs):
 
-\begin{itemize}
-\tightlist
-\item
-  Projects are accessible through the internet at OSF or GitHub
-\item
-  DOIs for a \emph{vertical} project can be minted through OSF or Zenodo
-\end{itemize}
+  - Projects are accessible through the internet at OSF or GitHub
+  - DOIs for a *vertical* project can be minted through OSF or Zenodo
 
-\textbf{A}ccessible data is available to approved researchers (e.g., anyone or lab members only) through a standardized communications protocol (e.g., Internet):
+**A**ccessible data is available to approved researchers (e.g., anyone or lab members only) through a standardized communications protocol (e.g., Internet):
 
-\begin{itemize}
-\tightlist
-\item
-  Projects on GitHub and OSF can be made accessible to a specific set of users, or completely private, if desired
-\item
-  Most data sharing guides emphasize manually uploading heterogeneous data files to a repository (e.g., OSF).
+  - Projects on GitHub and OSF can be made accessible to a specific set of users, or completely private, if desired
+  - Most data sharing guides emphasize manually uploading heterogeneous data files to a repository (e.g., OSF).
 R packages allow that, but also facilitate programmatic access to the data, and therefore are more accessible than methods relying on manual transfer of files
-\end{itemize}
 
-\textbf{I}nteroperable data is described (via metadata and file organization) in a common, openly available, non-proprietary language:
+**I**nteroperable data is described (via metadata and file organization) in a common, openly available, non-proprietary language:
 
-\begin{itemize}
-\tightlist
-\item
-  R packages describe a strict organization of source files, resulting in a software product that is directly usable in a programming environment
-\item
-  R packages describe a common language for documenting the data, its use, and dependencies
-\item
-  R package (meta)data is human- and machine-readable and in a non-proprietary format
-\item
-  Data wrapped in an R package is accessible in other programming environments due to standardized formatting
-\end{itemize}
+  - R packages describe a strict organization of source files, resulting in a software product that is directly usable in a programming environment
+  - R packages describe a common language for documenting the data, its use, and dependencies
+  - R package (meta)data is human- and machine-readable and in a non-proprietary format
+  - Data wrapped in an R package is accessible in other programming environments due to standardized formatting
 
-\textbf{R}eusable data are richly described and licensed to facilitate reuse:
+**R**eusable data are richly described and licensed to facilitate reuse:
 
-\begin{itemize}
-\tightlist
-\item
-  R packages are described in a standard format
-\item
-  R packages facilitate use of appropriate licenses
-\end{itemize}
+  - R packages are described in a standard format
+  - R packages facilitate use of appropriate licenses
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
+# Conclusion
 
@@
 By organizing research assets in a standardized format, their reuse will be less time consuming and error-prone.
-Standardized formats also facilitate large-scale meta-scientific investigations (Gorgolewski et al., 2016), and inclusion of data sets for meta-analyses.
+Standardized formats also facilitate large-scale meta-scientific investigations [@gorgolewski_brain_2016], and inclusion of data sets for meta-analyses.
 
-More generally, in addition to the convenience of a single-platform \emph{vertical} workflow, adopting R standards during asset creation has the side effect of demanding some amount of attention on how research products are curated, stored, documented, and shared.
+More generally, in addition to the convenience of a single-platform *vertical* workflow, adopting R standards during asset creation has the side effect of demanding some amount of attention on how research products are curated, stored, documented, and shared.
 Drawing attention to these details may serve to increase the reproducibility of our work.
 
-\hypertarget{references}{%
-\section{References}\label{references}}
+# References
+
+```{r create_r-references}
+create_bib(
+  x = c(
+    "devtools", "posterdown", "pkgdown", 
+    "vertical", "rmarkdown", "jspsychr",
+    "base", "usethis", "papaja", "knitr",
+    "rio", "roxygen2", "testthat", "citr"
+  ),
+  append = FALSE,
+  file = "r-references.bib"
+)
+```
 
 \begingroup
-
-\hypertarget{refs}{}
-\begin{cslreferences}
-\leavevmode\hypertarget{ref-R-citr}{}%
-Aust, F. (2019). \emph{Citr: 'RStudio' add-in to insert markdown citations}. Retrieved from \url{https://CRAN.R-project.org/package=citr}
-
-\leavevmode\hypertarget{ref-R-papaja}{}%
-Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}
-
-\leavevmode\hypertarget{ref-R-jspsychr}{}%
-Crump, M. J. C. (2019). \emph{Jspsychr: Templates and functions for writing and running jspsych experiments from r-studio}. Retrieved from \url{https://github.com/CrumpLab/jspsychr}
-
-\leavevmode\hypertarget{ref-R-vertical}{}%
-Crump, M.
-J.
-C., \& Vuorre, M.
-(n.d.).
-\emph{Vertical: Reproducible worfklow for psychological science research asset creation and communication}.
-Retrieved from \url{https://github.com/CrumpLab/vertical}
-
-\leavevmode\hypertarget{ref-de2015jspsych}{}%
-De Leeuw, J. R. (2015). JsPsych: A javascript library for creating behavioral experiments in a web browser. \emph{Behavior Research Methods}, \emph{47}(1), 1--12.
-
-\leavevmode\hypertarget{ref-gorgolewski_brain_2016}{}%
-Gorgolewski, K.
-J., Auer, T., Calhoun, V.
-D., Craddock, R.
-C., Das, S., Duff, E.
-P., \ldots{} Poldrack, R.
-A.
-(2016).
-The brain imaging data structure, a format for organizing and describing outputs of neuroimaging experiments.
-\emph{Scientific Data}, \emph{3}, 160044.
-\url{https://doi.org/10.1038/sdata.2016.44}
-
-\leavevmode\hypertarget{ref-HardwickeDataavailabilityreusability2018}{}%
-Hardwicke, T.
-E., Mathur, M.
-B., MacDonald, K., Nilsonne, G., Banks, G.
-C., Kidwell, M.
-C., \ldots{} Frank, M.
-C.
-(2018).
-Data availability, reusability, and analytic reproducibility: Evaluating the impact of a mandatory open data policy at the journal Cognition.
-\emph{Royal Society Open Science}, \emph{5}(8), 180448.
-\url{https://doi.org/10.1098/rsos.180448}
-
-\leavevmode\hypertarget{ref-HoutkoopDataSharingPsychology2018}{}%
-Houtkoop, B.
-L., Chambers, C., Macleod, M., Bishop, D.
-V.
-M., Nichols, T.
-E., \& Wagenmakers, E.-J.
-(2018).
-Data Sharing in Psychology: A Survey on Barriers and Preconditions.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(1), 70--85.
-\url{https://doi.org/10.1177/2515245917751886}
-
-\leavevmode\hypertarget{ref-KleinPracticalGuideTransparency2018}{}%
-Klein, O., Hardwicke, T.
-E., Aust, F., Breuer, J., Danielsson, H., Mohr, A.
-H., \ldots{} Frank, M.
-C.
-(2018).
-A Practical Guide for Transparency in Psychological Science.
-\emph{Collabra: Psychology}, \emph{4}(1), 20.
-\url{https://doi.org/10.1525/collabra.158}
-
-\leavevmode\hypertarget{ref-LindsaySharingDataMaterials2017}{}%
-Lindsay, D. S. (2017). Sharing Data and Materials in Psychological Science. \emph{Psychological Science}, \emph{28}(6), 699--702. \url{https://doi.org/10.1177/0956797617704015}
-
-\leavevmode\hypertarget{ref-MartoneDatasharingpsychology2018}{}%
-Martone, M. E., Garcia-Castro, A., \& VandenBos, G. R. (2018). Data sharing in psychology. \emph{American Psychologist}, \emph{73}(2), 111--125. \url{https://doi.org/10.1037/amp0000242}
-
-\leavevmode\hypertarget{ref-morey_peer_2016}{}%
-Morey, R.
-D., Chambers, C.
-D., Etchells, P.
-J., Harris, C.
-R., Hoekstra, R., Lakens, D., \ldots{} Zwaan, R.
-A.
-(2016).
-The Peer Reviewers Openness Initiative: Incentivizing open research practices through peer review.
-\emph{Royal Society Open Science}, \emph{3}(1), 150547.
-\url{https://doi.org/10.1098/rsos.150547}
-
-\leavevmode\hypertarget{ref-nosek_promoting_2015}{}%
-Nosek, B.
-A., Alter, G., Banks, G.
-C., Borsboom, D., Bowman, S.
-D., Breckler, S.
-J., \ldots{} Yarkoni, T.
-(2015).
-Promoting an open research culture.
-\emph{Science}, \emph{348}(6242), 1422--1425.
-\url{https://doi.org/10.1126/science.aab2374}
-
-\leavevmode\hypertarget{ref-R-base}{}%
-R Core Team. (2019). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
-
-\leavevmode\hypertarget{ref-Rouderwhatwhyhow2016}{}%
-Rouder, J. N. (2016). The what, why, and how of born-open data. \emph{Behavior Research Methods}, \emph{48}(3), 1062--1069. \url{https://doi.org/10.3758/s13428-015-0630-z}
-
-\leavevmode\hypertarget{ref-RouderMinimizingMistakesPsychological2019}{}%
-Rouder, J.
-N., Haaf, J.
-M., \& Snyder, H.
-K.
-(2019).
-Minimizing Mistakes in Psychological Science.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(1), 3--11.
-\url{https://doi.org/10.1177/2515245918801915}
-
-\leavevmode\hypertarget{ref-RStudioTeamRStudioIntegratedDevelopment2016}{}%
-RStudio Team. (2016). \emph{RStudio: Integrated Development Environment for R}. Boston, MA: RStudio, Inc. Retrieved from \url{http://www.rstudio.com/}
-
-\leavevmode\hypertarget{ref-SilberzahnManyAnalystsOne2018}{}%
-Silberzahn, R., Uhlmann, E.
-L., Martin, D.
-P., Anselmi, P., Aust, F., Awtrey, E., \ldots{} Nosek, B.
-A.
-(2018).
-Many Analysts, One Data Set: Making Transparent How Variations in Analytic Choices Affect Results.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(3), 337--356.
-\url{https://doi.org/10.1177/2515245917747646}
-
-\leavevmode\hypertarget{ref-R-posterdown}{}%
-Thorne, W.
-B.
-(2019).
-\emph{Posterdown: An r package built to generate reproducible conference posters for the academic and professional world where powerpoint and pages just won't cut it}.
-Retrieved from \url{https://github.com/brentthorne/posterdown}
-
-\leavevmode\hypertarget{ref-VanpaemelAreWeWasting2015}{}%
-Vanpaemel, W., Vermorgen, M., Deriemaecker, L., \& Storms, G.
-(2015).
-Are We Wasting a Good Crisis?
-The Availability of Psychological Research Data after the Storm.
-\emph{Collabra: Psychology}, \emph{1}(1), Art.
-3.
-\url{https://doi.org/10.1525/collabra.13}
-
-\leavevmode\hypertarget{ref-VuorreCuratingResearchAssets2018}{}%
-Vuorre, M., \& Curley, J.
-P.
-(2018).
-Curating Research Assets: A Tutorial on the Git Version Control System.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 219--236.
-\url{https://doi.org/10.1177/2515245918754826}
-
-\leavevmode\hypertarget{ref-Wichertspooravailabilitypsychological2006}{}%
-Wicherts, J.
-M., Borsboom, D., Kats, J., \& Molenaar, D.
-(2006).
-The poor availability of psychological research data for reanalysis.
-\emph{American Psychologist}, \emph{61}(7), 726--728.
-\url{https://doi.org/10.1037/0003-066X.61.7.726}
-
-\leavevmode\hypertarget{ref-R-testthat}{}%
-Wickham, H. (2011). Testthat: Get started with testing. \emph{The R Journal}, \emph{3}, 5--10. Retrieved from \url{https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf}
-
-\leavevmode\hypertarget{ref-WickhamPackagesOrganizeTest2015}{}%
-Wickham, H. (2015). \emph{R Packages: Organize, Test, Document, and Share Your Code}. "O'Reilly Media, Inc.". Retrieved from \url{http://r-pkgs.had.co.nz/}
-
-\leavevmode\hypertarget{ref-R-pkgdown}{}%
-Wickham, H., \& Hesselberth, J. (2019). \emph{Pkgdown: Make static html documentation for a package}. Retrieved from \url{https://CRAN.R-project.org/package=pkgdown}
-
-\leavevmode\hypertarget{ref-R-devtools}{}%
-Wickham, H., Hester, J., \& Chang, W. (2019). \emph{Devtools: Tools to make developing r packages easier}. Retrieved from \url{https://CRAN.R-project.org/package=devtools}
-
-\leavevmode\hypertarget{ref-WilkinsonFAIRGuidingPrinciples2016}{}%
-Wilkinson, M.
-D., Dumontier, M., Aalbersberg, I.
-J., Appleton, G., Axton, M., Baak, A., \ldots{} Mons, B.
-(2016).
-The FAIR Guiding Principles for scientific data management and stewardship.
-\emph{Scientific Data}, \emph{3}, 160018.
-\url{https://doi.org/10.1038/sdata.2016.18}
-
-\leavevmode\hypertarget{ref-R-rmarkdown}{}%
-Xie, Y., Allaire, J. J., \& Grolemund, G. (2018). \emph{R markdown: The definitive guide}. Boca Raton, Florida: Chapman; Hall/CRC. Retrieved from \url{https://bookdown.org/yihui/rmarkdown}
-
-\leavevmode\hypertarget{ref-yee_why_2017}{}%
-Yee, S.
-J.
-W., \& Debbie.
-(2017).
-Why You Should Become a UseR: A Brief Introduction to R.
-\emph{APS Observer}, \emph{30}(3).
-Retrieved from \url{https://www.psychologicalscience.org/observer/why-you-should-become-a-user-a-brief-introduction-to-r}
-
-\leavevmode\hypertarget{ref-R-citr}{}%
-Aust, F. (2019). \emph{Citr: 'RStudio' add-in to insert markdown citations}. Retrieved from \url{https://CRAN.R-project.org/package=citr}
-
-\leavevmode\hypertarget{ref-R-papaja}{}%
-Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}
-
-\leavevmode\hypertarget{ref-R-jspsychr}{}%
-Crump, M. J. C. (2019). \emph{Jspsychr: Templates and functions for writing and running jspsych experiments from r-studio}. Retrieved from \url{https://github.com/CrumpLab/jspsychr}
-
-\leavevmode\hypertarget{ref-R-vertical}{}%
-Crump, M.
-J.
-C., \& Vuorre, M.
-(n.d.).
-\emph{Vertical: Reproducible worfklow for psychological science research asset creation and communication}.
-Retrieved from \url{https://github.com/CrumpLab/vertical}
-
-\leavevmode\hypertarget{ref-de2015jspsych}{}%
-De Leeuw, J. R. (2015). JsPsych: A javascript library for creating behavioral experiments in a web browser. \emph{Behavior Research Methods}, \emph{47}(1), 1--12.
-
-\leavevmode\hypertarget{ref-gorgolewski_brain_2016}{}%
-Gorgolewski, K.
-J., Auer, T., Calhoun, V.
-D., Craddock, R.
-C., Das, S., Duff, E.
-P., \ldots{} Poldrack, R.
-A.
-(2016).
-The brain imaging data structure, a format for organizing and describing outputs of neuroimaging experiments.
-\emph{Scientific Data}, \emph{3}, 160044.
-\url{https://doi.org/10.1038/sdata.2016.44}
-
-\leavevmode\hypertarget{ref-HardwickeDataavailabilityreusability2018}{}%
-Hardwicke, T.
-E., Mathur, M.
-B., MacDonald, K., Nilsonne, G., Banks, G.
-C., Kidwell, M.
-C., \ldots{} Frank, M.
-C.
-(2018).
-Data availability, reusability, and analytic reproducibility: Evaluating the impact of a mandatory open data policy at the journal Cognition.
-\emph{Royal Society Open Science}, \emph{5}(8), 180448.
-\url{https://doi.org/10.1098/rsos.180448}
-
-\leavevmode\hypertarget{ref-HoutkoopDataSharingPsychology2018}{}%
-Houtkoop, B.
-L., Chambers, C., Macleod, M., Bishop, D.
-V.
-M., Nichols, T.
-E., \& Wagenmakers, E.-J.
-(2018).
-Data Sharing in Psychology: A Survey on Barriers and Preconditions.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(1), 70--85.
-\url{https://doi.org/10.1177/2515245917751886}
-
-\leavevmode\hypertarget{ref-KleinPracticalGuideTransparency2018}{}%
-Klein, O., Hardwicke, T.
-E., Aust, F., Breuer, J., Danielsson, H., Mohr, A.
-H., \ldots{} Frank, M.
-C.
-(2018).
-A Practical Guide for Transparency in Psychological Science.
-\emph{Collabra: Psychology}, \emph{4}(1), 20.
-\url{https://doi.org/10.1525/collabra.158}
-
-\leavevmode\hypertarget{ref-LindsaySharingDataMaterials2017}{}%
-Lindsay, D. S. (2017). Sharing Data and Materials in Psychological Science. \emph{Psychological Science}, \emph{28}(6), 699--702. \url{https://doi.org/10.1177/0956797617704015}
-
-\leavevmode\hypertarget{ref-MartoneDatasharingpsychology2018}{}%
-Martone, M. E., Garcia-Castro, A., \& VandenBos, G. R. (2018). Data sharing in psychology. \emph{American Psychologist}, \emph{73}(2), 111--125. \url{https://doi.org/10.1037/amp0000242}
-
-\leavevmode\hypertarget{ref-morey_peer_2016}{}%
-Morey, R.
-D., Chambers, C.
-D., Etchells, P.
-J., Harris, C.
-R., Hoekstra, R., Lakens, D., \ldots{} Zwaan, R.
-A.
-(2016).
-The Peer Reviewers Openness Initiative: Incentivizing open research practices through peer review.
-\emph{Royal Society Open Science}, \emph{3}(1), 150547.
-\url{https://doi.org/10.1098/rsos.150547}
-
-\leavevmode\hypertarget{ref-nosek_promoting_2015}{}%
-Nosek, B.
-A., Alter, G., Banks, G.
-C., Borsboom, D., Bowman, S.
-D., Breckler, S.
-J., \ldots{} Yarkoni, T.
-(2015).
-Promoting an open research culture.
-\emph{Science}, \emph{348}(6242), 1422--1425.
-\url{https://doi.org/10.1126/science.aab2374}
-
-\leavevmode\hypertarget{ref-R-base}{}%
-R Core Team. (2019). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
-
-\leavevmode\hypertarget{ref-Rouderwhatwhyhow2016}{}%
-Rouder, J. N. (2016). The what, why, and how of born-open data. \emph{Behavior Research Methods}, \emph{48}(3), 1062--1069. \url{https://doi.org/10.3758/s13428-015-0630-z}
-
-\leavevmode\hypertarget{ref-RouderMinimizingMistakesPsychological2019}{}%
-Rouder, J.
-N., Haaf, J.
-M., \& Snyder, H.
-K.
-(2019).
-Minimizing Mistakes in Psychological Science.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(1), 3--11.
-\url{https://doi.org/10.1177/2515245918801915}
-
-\leavevmode\hypertarget{ref-RStudioTeamRStudioIntegratedDevelopment2016}{}%
-RStudio Team. (2016). \emph{RStudio: Integrated Development Environment for R}. Boston, MA: RStudio, Inc. Retrieved from \url{http://www.rstudio.com/}
-
-\leavevmode\hypertarget{ref-SilberzahnManyAnalystsOne2018}{}%
-Silberzahn, R., Uhlmann, E.
-L., Martin, D.
-P., Anselmi, P., Aust, F., Awtrey, E., \ldots{} Nosek, B.
-A.
-(2018).
-Many Analysts, One Data Set: Making Transparent How Variations in Analytic Choices Affect Results.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(3), 337--356.
-\url{https://doi.org/10.1177/2515245917747646}
-
-\leavevmode\hypertarget{ref-R-posterdown}{}%
-Thorne, W.
-B.
-(2019).
-\emph{Posterdown: An r package built to generate reproducible conference posters for the academic and professional world where powerpoint and pages just won't cut it}.
-Retrieved from \url{https://github.com/brentthorne/posterdown}
-
-\leavevmode\hypertarget{ref-VanpaemelAreWeWasting2015}{}%
-Vanpaemel, W., Vermorgen, M., Deriemaecker, L., \& Storms, G.
-(2015).
-Are We Wasting a Good Crisis?
-The Availability of Psychological Research Data after the Storm.
-\emph{Collabra: Psychology}, \emph{1}(1), Art.
-3.
-\url{https://doi.org/10.1525/collabra.13}
-
-\leavevmode\hypertarget{ref-VuorreCuratingResearchAssets2018}{}%
-Vuorre, M., \& Curley, J.
-P.
-(2018).
-Curating Research Assets: A Tutorial on the Git Version Control System.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 219--236.
-\url{https://doi.org/10.1177/2515245918754826}
-
-\leavevmode\hypertarget{ref-Wichertspooravailabilitypsychological2006}{}%
-Wicherts, J.
-M., Borsboom, D., Kats, J., \& Molenaar, D.
-(2006).
-The poor availability of psychological research data for reanalysis.
-\emph{American Psychologist}, \emph{61}(7), 726--728.
-\url{https://doi.org/10.1037/0003-066X.61.7.726}
-
-\leavevmode\hypertarget{ref-R-testthat}{}%
-Wickham, H. (2011). Testthat: Get started with testing. \emph{The R Journal}, \emph{3}, 5--10. Retrieved from \url{https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf}
-
-\leavevmode\hypertarget{ref-WickhamPackagesOrganizeTest2015}{}%
-Wickham, H. (2015). \emph{R Packages: Organize, Test, Document, and Share Your Code}. "O'Reilly Media, Inc.". Retrieved from \url{http://r-pkgs.had.co.nz/}
-
-\leavevmode\hypertarget{ref-R-pkgdown}{}%
-Wickham, H., \& Hesselberth, J. (2019). \emph{Pkgdown: Make static html documentation for a package}. Retrieved from \url{https://CRAN.R-project.org/package=pkgdown}
-
-\leavevmode\hypertarget{ref-R-devtools}{}%
-Wickham, H., Hester, J., \& Chang, W. (2019). \emph{Devtools: Tools to make developing r packages easier}. Retrieved from \url{https://CRAN.R-project.org/package=devtools}
-
-\leavevmode\hypertarget{ref-WilkinsonFAIRGuidingPrinciples2016}{}%
-Wilkinson, M.
-D., Dumontier, M., Aalbersberg, I.
-J., Appleton, G., Axton, M., Baak, A., \ldots{} Mons, B.
-(2016).
-The FAIR Guiding Principles for scientific data management and stewardship.
-\emph{Scientific Data}, \emph{3}, 160018.
-\url{https://doi.org/10.1038/sdata.2016.18}
-
-\leavevmode\hypertarget{ref-R-rmarkdown}{}%
-Xie, Y., Allaire, J. J., \& Grolemund, G. (2018). \emph{R markdown: The definitive guide}. Boca Raton, Florida: Chapman; Hall/CRC. Retrieved from \url{https://bookdown.org/yihui/rmarkdown}
-
-\leavevmode\hypertarget{ref-yee_why_2017}{}%
-Yee, S.
-J.
-W., \& Debbie.
-(2017).
-Why You Should Become a UseR: A Brief Introduction to R.
-\emph{APS Observer}, \emph{30}(3).
-Retrieved from \url{https://www.psychologicalscience.org/observer/why-you-should-become-a-user-a-brief-introduction-to-r}
-\end{cslreferences}
-
+<div id = "refs"></div>
 \endgroup
-
-\end{document}

### 2020-01-13 R/build.R

@@ -16,7 +16,7 @@ build_vertical <- function(clean=TRUE,...) {
   pkgdown::build_site(...)
 
   for (i in list.files(pattern = "\\.Rmd", recursive = TRUE)) {
-    if(unlist(strsplit(i,split=.Platform$file.sep))[1] %in% c("experiments","vignettes") == FALSE){
+    if(unlist(strsplit(i,split=.Platform$file.sep))[1] %in% c("experiments","vignettes","inst") == FALSE){
       out.file <- rmarkdown::render(i, output_dir = paste0("docs/", dirname(i)), quiet=TRUE)
       if (!any(grepl(tools::file_path_sans_ext(i), readLines("_pkgdown.yml")))) {
         usethis::ui_info(paste(file.path(dirname(i),basename(out.file)),

### 2020-01-13 R/maintain.R

@@ -47,3 +47,5 @@ suggest_yml <- function(){
   writeLines(readLines(filename))
   unlink(filename)
 }
+
+

### 2020-01-14 manuscript/manuscript.Rmd
update manuscript

@@
 wordcount: "`r wordcountaddin::word_count('manuscript.Rmd')`"
-bibliography: ["references.bib","r-references.bib"]
+bibliography: ["references.bib", "r-references.bib"]
 floatsintext: yes
@@
 draft: no
-documentclass: apa6
-classoption: jou
-output: papaja::apa6_pdf
-pkgdown:
-  as_is: true
-  extension: pdf
+documentclass: "apa6"
+classoption: "jou"
+output: 
+  papaja::apa6_pdf:
+    keep_tex: false
 ---
@@
 
-Research projects produce many assets, such as experiments, data, analyses, manuscripts, posters, slides, stimulus sets and materials, computational models, and more.
-Although creating and communicating research assets are mainstay activities in science, there is a trend [@VanpaemelAreWeWasting2015; @Wichertspooravailabilitypsychological2006] encouraging more open and transparent sharing practices [@HoutkoopDataSharingPsychology2018; @LindsaySharingDataMaterials2017].
-However, the value of shared assets is often limited by poor documentation, incompatible file formats, or lack of organization, resulting in low rates of reproducibility [@HardwickeDataavailabilityreusability2018].
-Some data sharing guidelines and good practices have been suggested [e.g., @RouderMinimizingMistakesPsychological2019; @Rouderwhatwhyhow2016]; along with related issues of ethics and privacy, ownership of data, attitudes about sharing, and where to share data [@KleinPracticalGuideTransparency2018; @MartoneDatasharingpsychology2018].
-Nevertheless, curating standards for research assets have not emerged, possibly due to the variance of research in psychology.
+Research projects produce experiments, data, analyses, manuscripts, posters, slides, stimuli and materials, computational models, and more.
+However, the potential added value of these products is not fully realized due to limited sharing and curating practices.
+Although more transparent communication of these research products has recently been encouraged [@HoutkoopDataSharingPsychology2018; @LindsaySharingDataMaterials2017; @VanpaemelAreWeWasting2015; @Wichertspooravailabilitypsychological2006; @KleinPracticalGuideTransparency2018; @MartoneDatasharingpsychology2018; @RouderMinimizingMistakesPsychological2019; @Rouderwhatwhyhow2016], these efforts often focus narrowly on sharing data (and sometimes analysis code).
+Further, the practical value of sharing is often limited by poor documentation, incompatible file formats, and lack of organization, resulting in low rates of reproducibility [@HardwickeDataavailabilityreusability2018].
+Standardization of protocols for sharing would be beneficial; but, such standards have not emerged, possibly due to the variance of research in psychology.
 Instead of developing another standard, we suggest borrowing existing standards and practices from software engineering.
@@
 
-We sought for a standard for organization and sharing that would adhere to the FAIR (Findable, Accessible, Interoperable, Reusable) guidelines to maximize the reuse potential of data and support "discovery through good data management" [@WilkinsonFAIRGuidingPrinciples2016, see also general discussion].
-Additionally, we recognized the added value of including other research outputs ("products"; e.g., manuscripts) in a reproducible collection of materials that could be shared with the data.
-We identified the R package standard as a solution that doesn't present overwhelming overhead for already busy researchers.
+We sought for a standard for organizing and sharing that would adhere to the FAIR (Findable, Accessible, Interoperable, Reusable) guidelines to maximize the reuse potential of data and support "discovery through good data management" [@WilkinsonFAIRGuidingPrinciples2016, see also general discussion].
+Additionally, we recognized the added value of including other research outputs ("products"; e.g., manuscripts) beyond datasets in a reproducible collection of materials.
+We identified the R package standard with modern online-based workflows as a solution that doesn't present overwhelming overhead for already busy researchers.
 Here, we discuss the R package standard for creating reproducible research material containers for data, analysis code, and metadata.
-Then, we introduce a collection of additional R packages for incorporating other research products into the project.
+Then, we introduce additional R packages for incorporating other research products into the reproducible container.
 These additional tools are easily accessible through our *vertical* package, thus named because an entire research project can be completed from top to bottom as a stack of R processes from within the RStudio software platform.
@@
 
-R is a programming language commonly used for statistical analyses [@R-base], and is available at <https://www.r-project.org>.
-RStudio is an integrated development environment [IDE; @RStudioTeamRStudioIntegratedDevelopment2016], and is available at <https://www.rstudio.com/>.
-Both are free, open source and work on Windows, Mac, and Linux operating systems.
-We chose these tools because they are already widely used among psychology researchers: In one crowd-sourced analysis, 16 of 29 analysis teams used R [@SilberzahnManyAnalystsOne2018; see also @yee_why_2017 and <http://r4stats.com/articles/popularity/>].
-Further, many psychologists have already developed R packages that share analysis tools and data.
+R is a programming language for statistical analyses (@R-base; <https://www.r-project.org>), and RStudio is its associated integrated development environment (IDE; @RStudioTeamRStudioIntegratedDevelopment2016; <https://www.rstudio.com/>).
+Both are free, open source and work on Windows, Mac, and Linux systems.
+These tools are already widely used among psychology researchers: In one crowd-sourced analysis, 16 of 29 analysis teams used R [@SilberzahnManyAnalystsOne2018], and many psychologists have already developed R packages [see also @yee_why_2017 and <http://r4stats.com/articles/popularity/>].
 
-A cornerstone of R are user-created packages, which contain R functions, data, metadata, and documentation written in a standardized format that allows seamless sharing across users and operating systems.
-When data and functions are wrapped in R packages, they are immediately available to other R users through the R console, and their documentation can be easily viewed in R or online.
-Next, we outline how data and functions are included in R packages, and then turn to including other assets, such as manuscripts and posters.
-We present a brief outline, a complete guide to creating R packages is outside the scope of this manuscript, and we refer readers to Wickham [-@WickhamPackagesOrganizeTest2015], which remains an authoritative and in-depth guide on creating R packages.
+A cornerstone of R are user-created packages, which contain functions, (meta)data, and documentation in a standard format that allows seamless sharing across users and operating systems.
+Data and functions in R packages are immediately available to others through the R console, and documentation can be viewed in R or online.
+We first outline how data and functions are included in R packages, and then turn to including other assets, such as manuscripts and posters.
+A complete guide to R packages is outside our scope; for a complete treatment, see Wickham [-@WickhamPackagesOrganizeTest2015].
 
@@
 
-Creating a package through RStudio is straightforward (see Figure \@ref(fig:creating-package-rstudio)).
-When a new package is created with the method shown in Figure \@ref(fig:creating-package-rstudio), the required files and directories are automatically created.
-First, the "DESCRIPTION" file includes basic information about the package (e.g., authors, description) in a machine readable format.
-To edit it, users can click on the file in the RStudio File browser.
+When an R package is created as shown in Figure \@ref(fig:creating-package-rstudio), the required files and directories are automatically created.
+At this point, the package can be installed (click "Install and Restart" in RStudio's "Build" tab) and shared online (see below).
+However, to make it useful, content needs to be added, beginning with the package's description.
+That information (e.g., authors, dependencies) is written in the "DESCRIPTION" file in a machine readable format.
+To edit that file, click on its name in RStudio's File browser.
 The "blank.Rproj", "man" and "NAMESPACE" files/directories shown in Figure \@ref(fig:creating-package-rstudio) should not be edited by users, and we thus turn to R functions.
@@
 
-Because R packages are primarily intended for R functions, an "R" directory is automatically created to store source code for defining functions.
 Researchers often develop and use custom functions in their analysis.
 Although it is possible to declare them directly in the analysis scripts, we suggest declaring functions using the R package standard, by placing the function's source code into a file in the "R" directory.
+That directory is automatically created (see Figure \@ref(fig:creating-package-rstudio)).
 Functions in R packages are portable, such that others can install the package from their R console, load it, and start using the functions immediately.
-Packages can also depend on other packages (and be depended on), such that R will automatically include any other packages required for your functions to run appropriately.
-Functions within R packages are documented in a standardized manner, and the documentation for a function can be viewed within R or online.
+Packages can also depend on other packages (and be depended on), such that R automatically installs any requirements for your functions to work appropriately.
+Functions within R packages are documented in a standardized manner, and the documentation for a function can be viewed in R (e.g.
+try `?mean`) or online.
 
-Additionally, learning and following R conventions for declaring functions has a pedagogical benefit to the researcher and may improve their practices.
-There is also a reuse benefit to the researcher: Functions can be difficult to find in old scripts, but easy to find and load if they are already packaged and called from an existing package.
+Learning and following R conventions for declaring functions has a pedagogical benefit to the researcher and may improve their practices.
+There is also a reuse benefit: Functions can be difficult to find in old scripts, but easy to find and load if they are called from an existing package.
 Thus, formally including one's functions in R packages facilitates reproducibility and sharing.
@@
 First, raw data in any format is stored in the "data-raw" directory to ensure that the method of sharing remains software-agnostic.
-Importantly, the raw data files should never be modified themselves.
-Second, instructions for converting and cleaning the raw data into analyzable format should also be included in this directory.
-These pre-processing steps should be declared so that the entire analysis is transparent and reproducible, preferably in an R script in the "data-raw" directory.
-That script should end with creating an R data object in the project's "data" directory.
+Second, instructions for converting and cleaning the raw data into analyzable format should also be included in this directory (the raw data files should not themselves be modified).
+These pre-processing steps should be saved so that the complete path from raw data to results is transparent and reproducible, preferably in an R script in the "data-raw" directory.
+That script should end with creating an R data object in the "data" directory.
 Placing the R data object there ensures that it is included in the resulting R package, which makes using the data effortless for any R user, and eschews the need to download additional files.
@@
 
-To recap, R package standards offer a systematic method for creating and curating reproducible research assets.
-Researchers could use the R package infrastructure to store and document their data and functions, and, as we will see, analyses,  manuscripts, posters, and presentations as well.
-The entire research project is then easily shared online, and showcased as a website, as we will show below.
-To enable these additional features, many R packages have been developed (e.g., writing manuscripts within R).
+R packages thus provide a useful standard for functions and (meta)data.
+However, as we will see, the package structure can be extended to include analyses, manuscripts, posters, and presentations as well.
+The complete project is then easily shared online, and even showcased as a website, as we will show below.
+To enable these additional features, many R packages have been developed (e.g., writing manuscripts with R).
 To facilitate their use, we have created an R package, *vertical*, that, when installed, also installs these other packages, and makes it easier to create projects that use them.
@@
 *vertical* extends the R package template with a set of additional files and directories for products such as manuscripts and posters.
-These assets are organized such that they can then be easily shared online on GitHub or OSF, installed to R as an R package, and showcased online as a website.
+These assets are organized such that they can be easily shared online on GitHub or OSF, installed to R as a package, and showcased online as a website.
 Figure \@ref(fig:vertical-workflow) illustrates the workflow of creating a *vertical* research project.
@@
 
-Additional benefits of packaging everything together in this manner are that every computation supporting manuscripts, supplementary analyses, data preprocessing, etc.
-is reproducible and uses the same data in the same computational environment.
-Because sharing of materials is built-in, no additional work is needed after-the-fact in organizing the materials for sharing them.
+Additional benefits of packaging everything together in this manner are that the computations supporting manuscripts, supplementary analyses, data preprocessing, etc.
+are reproducible and use the same data in the same computational environment.
+Because sharing of materials is built-in, no additional work is needed after-the-fact in organizing the materials for sharing.
 We begin with an overview of this process, but a more exhaustive step-by-step tutorial is provided online [@R-vertical][^v1].
@@
 
-After installing the package, users should restart RStudio to make all its features available.
-First, similar to creating an R package as shown above, a new *vertical* research project is created through the RStudio GUI (Figure \@ref(fig:vertical-project-structure).1).
+Users should then restart RStudio to make the features available.
+A new *vertical* research project is created through RStudio (Figure \@ref(fig:vertical-project-structure).1).
 Figure \@ref(fig:vertical-project-structure).2 shows the resulting file and directory structure, which is an extension of the R package structure shown in Figure \@ref(fig:creating-package-rstudio).
-Users may enable or disable components (as indicated by directories like "experiments") as needed.
+Users may enable or disable components as needed, either by (un)checking boxes (Figure \@ref(fig:vertical-project-structure).1) or adding/deleting files/directories.
 Next, we describe the components, and then how to share the resulting research projects and showcase them as websites.
 
-```{r vertical-project-structure, echo = F, eval = T, fig.cap = "Creating a *vertical* project with the RStudio *vertical* project template, and description of the file-organization and structure of a *vertical* project.", out.width="\\textwidth", fig.env="figure*"}
+```{r vertical-project-structure, echo = F, eval = T, fig.cap = "Creating a \\emph{vertical} project in RStudio (1), and description of the file structure of a \\emph{vertical} project (2).", out.width="\\textwidth", fig.env="figure*"}
 include_graphics("images/vertical-project.png", dpi = 300)
@@
 
-In addition to data and functions, almost any kind of content can be included within the project directory structure.
-Most importantly, to create reproducible assets, such as data analyses or manuscripts, these materials should be created with the R Markdown language.
-R Markdown is a simple plain text format that allows mixing prose with code (R, Python, etc.), and thus enables including one's text and analyses in a single document.
+In addition to data and functions, almost any kind of content can be included within the project.
+Most importantly, to create reproducible assets, such as data analyses or manuscripts, these materials should be created with R Markdown, a simple language that allows mixing prose with code (R, Python, etc.), and thus enables including one's text and analyses in a single document.
 Because of its importance, we describe R Markdown first.
@@
 
-R Markdown is a plain text file format and an R package that allows executable code snippets to be embedded alongside regular text [@R-rmarkdown].
-For example, an R Markdown file can include the typical prose found in manuscripts, the analysis code (R, Python, etc.), and code snippets that display numerical results and figures within the text.
+R Markdown allows executable code snippets to be embedded alongside regular text, and simple markup for formatting code (e.g headers using `#`s; @R-rmarkdown).
 R Markdown documents are transparent and reproducible.
 In principle, recipients can see the analysis scripts, and reproduce them by compiling the document on their own machine.
-R Markdown documents separate content and style, and can be compiled to multiple output formats, such as PDF, HTML, and Microsoft Word.
-Importantly, R Markdown is basically plain text, and thus effortless to learn.
+R Markdown documents separate content and style, and can be compiled to multiple output formats, such as PDF, HTML, and Word.
+Importantly, R Markdown is plain text, and thus effortless to learn.
 Nevertheless, a tutorial on R Markdown is beyond our scope [see @R-rmarkdown][^rmark].
@@
 Above, we suggested placing any preprocessing steps in a separate file in the "data-raw" directory, and creating a documented R data object in "data".
-If the data object is called `mydata`, once the package is installed and loaded, simply calling `mydata` in an R script will call the data object, ready to be used in analyses.
+If the data object is called `mydata` and the package is installed and loaded, calling `mydata` in R will call the data object, ready to be used in analyses.
 Our favored approach is to write any analyses relevant to a product (e.g., manuscript or a poster) directly into the product's R Markdown source file.
@@
 
-Reproducible APA formatted manuscripts can easily be created with the *papaja* R package, that provides an R Markdown template file, and additional helper functions for creating tables and figures [@R-papaja].
+Reproducible APA formatted manuscripts can be created with the *papaja* R package, that provides an R Markdown template file, and additional helper functions for creating tables and figures [@R-papaja].
 For example, the R Markdown file can contain the text and code to generate the manuscript and all results.
-*papaja* provides several functions for reporting results, especially from common statistical analyses such as ANOVAs, t-tests, and regressions.
-For example, *papaja*'s `apa_print()` function embeds statistical results directly into the manuscript, obviating the need to copy them by hand and removing human error in reporting.
-References can be automatically populated from Zotero using the `citr` plugin [@R-citr].
+*papaja* provides several functions for reporting results: For example, *papaja*'s `apa_print(model)` embeds `model`'s statistical results directly into the manuscript, obviating the need to copy them by hand and removing human error in reporting.
+References can be automatically added from Zotero using the `citr` plugin [@R-citr].
 We refer readers to the *papaja* documentation website for more information.[^ppj]
@@
 
-*vertical* includes a posters folder where poster documents can be deposited and shared on the website.
-Additionally, *vertical* suggests using the *posterdown* package [@R-posterdown] for poster creation through an R Markdown document.
+*vertical* includes a posters folder where poster documents can be deposited and shared.
+*vertical* suggests using the *posterdown* package [@R-posterdown] for poster creation through an R Markdown document.
 When the provided template is used, the resulting reproducible document is viewable as part of the website.
@@
 When a new *vertical* project is created, a *slidy* template is included to create web-ready presentations [^slidy].
-Slidy presentations are viewable in a browser like typical PowerPoint or keynote presentations, but have some additional features like long scroll-able slides, interactive figures and animations, and rendering full websites within individual slides.
+Slidy presentations are viewable in a browser like typical PowerPoint or Keynote presentations, but have additional features like scrollable slides, interactive figures, and animations.
 
@@
 
-Other research assets may include supplementary materials, stimulus sets, code for a computational model, additional analyses, or other documents, like a public-facing blog.
-*vertical* is flexible with respect to the inclusion of additional assets not specified by the template.
+Other research assets may include supplementary materials, stimulus sets, code for a computational model, additional analyses, or other documents, such as blogs.
+Additional assets can be included by creating files in appropriate directories, and linking them to the resulting website (see below).
 For example, arbitrary R Markdown documents can be included in the vignettes directory.
-When the project is compiled (see below), any R Markdown files in that directory are compiled and served on the website.
+When the project is compiled (see below), R Markdown files in that directory are automatically compiled and served on the website (see @R-pkgdown).
 
@@
 
-Finally, *vertical* projects can include experiments in the experiments directory.
-These materials could be packaged as a compressed .zip file, which would then be accessible from an online repository.
+Finally, experiments should be placed in the experiments directory.
+They could be packaged as a compressed .zip file, which would then be accessible from an online repository.
 *vertical* suggests using *jsPsych* [@de2015jspsych], a JavaScript library for building browser-based experiments that can be served to the web or run locally.
 Selecting the *jsPsych* module during project initialization downloads the most recent version of *jsPsych* to the experiments folder.
-A basic html template is also provided, which loads the jsPsych library and presents some basic word stimuli.
-To further edit the experiment, users can simply edit the html file as they would when creating any other jsPsych experiment.
-When the *vertical* project is compiled, this experiment is immediately available from the website navigation bar.
+A basic HTML template is also provided, which loads the jsPsych library and presents some basic word stimuli.
+To further edit the experiment, users can simply edit the HTML file as they would when creating any other jsPsych experiment.
+When the *vertical* project is compiled, the experiment is available on the project website.
 
-For users who wish to work entirely within R Markdown, we suggest *jspsychr* [@R-jspsychr] [^jspsychr], an R package that helps creating jsPsych experiments with R Markdown in RStudio.
-This option allows using RStudio's ability to pass objects between R and JavaScript when constructing experiment materials.
+For working entirely with R Markdown, we suggest *jspsychr*, an R package that helps creating jsPsych experiments with R Markdown in RStudio [@R-jspsychr] [^jspsychr].
+RStudio enables passing objects between R and JavaScript, making this an attractive option for experienced R users.
 The template creates a sample *jsPsych* experiment produced by *jspsychr*, which contains the R Markdown script to generate the experiment file, the html file to run the experiment, and a folder to save the data and document the experiment.
@@
 
-Browser-friendly experiments have several benefits over local-only experiments.
-First, scripts are open-source and easily reusable (no installation is required to run an experiment).
-Second, scripts can be shared in a self-contained manner such that the experiment is reproducible and the methods are transparent and verifiable.
-Third, because the experiment is an html file it can be shared and run as part of the project website, facilitating understanding of the experimental procedures.
-Sharing experiment source code with a demonstration is not yet common, but could benefit the review process and readers' understanding of the procedures by making methods more transparent.
+Browser-friendly experiments have several benefits.
+First, scripts are open-source and easy to use and share (no installation is required to run an experiment).
+Scripts can be shared in a self-contained manner such that the experiment is reproducible and the methods are transparent and verifiable.
+Third, because the experiment is an HTML file it can be shared and run as part of the project website, facilitating understanding of the experimental procedures.
+Sharing experiment source code with a demonstration could benefit the review process and readers' understanding of the procedures.
 
@@
 Above, we discussed using R Markdown, R package standards, and other R resources for creating research products.
-Then next step is sharing and communicating these products.
-When the products are created as suggested above, *vertical* facilitates sharing the entire project as an R package for easy access to the data and functions, as a version controlled repository for access to the complete source code for reproducing every component of the project, and as a website, where the components are easily viewed by others.
+The next step is sharing and communicating these products.
+When the products are created as suggested above, *vertical* facilitates sharing the entire project as an R package for easy access to the data and functions, as a version controlled repository for the project's source code, and as a website, where the components can be viewed in their final format.
 
@@
 
-An organizing principle of our approach is that all project materials are stored in a single directory on the user's local filesystem.
-When version control, such as Git, is enabled in this directory, the entire project can be seamlessly shared with others through online platforms such as GitHub and GitLab.
-Version control systems provide filesystems with additional functioning for keeping track of files' versions, improved options for collaborating, and seamless integration with online collaboration platforms (GitHub and GitLab); however, the details are outside the scope of this manuscript and we refer readers to Vuorre & Curley [-@VuorreCuratingResearchAssets2018].
+The core of our proposal is that all project materials are stored in one directory on the user's local filesystem.
+When version control, such as Git, is enabled in this directory, the entire project can be seamlessly shared with others through online platforms such as GitHub, GitLab, and OSF.
+Version control systems provide filesystems with additional functioning for keeping track of files' versions, improved options for collaborating, and seamless integration with online collaboration platforms; however, the details are outside our scope, and we refer readers to Vuorre & Curley [-@VuorreCuratingResearchAssets2018].
 By default, *vertical* projects enable Git for new projects.
-Further, projects developed with Git (on a sharing platform such as GitHub or GitLab) can be seamlessly integrated with the Open Science Framework for long-term curation and, for example, to enable digital object identifiers (DOI).
 
@@
 
-An additional benefit of the R package organizing principles, and using the *vertical* helper tools, is that the entire project can be showcased as a website (see Figure \@ref(fig:vertical-project-website).
-The website is compiled by *pkgdown* [@R-pkgdown] in one command (`vertical::build_vertical()`), or by clicking Addins -> Build Vertical in RStudio, and expertise with web standards is not required.
-For instructions on customizing the resulting website, see the *pkgdown* instructions (<https://usethis.r-lib.org/>).
+An additional benefit of the R package organizing principles is that the entire project can be easily showcased as a website (see Figure \@ref(fig:vertical-project-website).
+The website is compiled by *pkgdown* [@R-pkgdown] (click "Addins" -> "Build Vertical" in RStudio), which is automatically included with *vertical* projects.
+For instructions on customizing the website, see <https://pkgdown.r-lib.org/>.
 
 Building a *vertical* project as shown above creates the website in the docs folder, from where it can be viewed.
-During this process, all the R Markdown files within the project are evaluated while the documents are compiled, thus ensuring that all analyses, and thus products, are up to date.
-Importantly, the website can be easily made available on the internet.
-To do so, users should integrate their project with a GitHub repository [@VuorreCuratingResearchAssets2018], and then enable GitHub pages under the repository's settings.
-An example of the resulting website can be viewed at <https://mvuorre.github.io/rparp/>.
+During this process, all the R Markdown files within the project are evaluated, ensuring that all analyses, and thus products, are up to date and use the same data in the same computational environment.
+The website is easily made available on the internet by integrating the project with a GitHub repository [@VuorreCuratingResearchAssets2018], and then enabling GitHub pages under the repository's settings.
+See <https://crumplab.github.io/vertical/> for an example.
 
@@
 
-If the project contains R functions, data, and documentation following the R package standard (as discussed above), then others can install the package on their computers for fast access to the data and functions from within R.
-Again, integrating the project with GitHub is beneficial, because then others can install the package from GitHub using the *devtools* package in R (`devtools::install_github("account/repository")`).
+If the project contains R functions, data, and documentation following the R package standard, then others can install the package on their computers for fast access to the data and functions from R.
+Again, integrating the project with GitHub is beneficial, because then others can install the package from GitHub using the *devtools* package, as was shown above in how to install *vertical*.
 Sharing as an R package is very convenient for other R users, as any packaged data and functions are immediately available for use, natively in R, after package installation.
@@
 It might seem that the benefits of this approach apply only to R users.
-We believe this is not the case; the workflow creates a single package that includes all the research products related to a product, which can then be viewed in a public (or private, if users so choose) website and source code repository.
-Of course, to get the most out of the proposal, some R use is required (from both the creator and subsequent users.) However, any standardized computational procedure assumes some computational environment, our choice of R is natural due to its already widespread adoption in the psychological researcher community.
+We believe this is not the case; the workflow creates a single repository that includes all the research products related to a project, which can then be viewed as a public (or private, if desired) website and source code repository.
+Of course, to get the most out of the proposal, some R use is required.
+However, any standardized computational procedure assumes some computational environment, and our choice of R is natural due to its already widespread adoption in the psychological researcher community.
 Furthermore, by requiring raw data inclusion, our proposal does not strictly require using R.
@@
 
-Our suggested approach has wider implications for making research practices in psychology more transparent and reproducible.
+Our suggestion has wider implications for research transparency and reproducibility.
 Consider the implied truth claims associated with any peer-reviewed manuscript.
 The act of publishing can be seen as a validity assertion: that the truth claims made by the authors are verifiably true, or more weakly that the research assets are verifiable.
-Unlike a set of conclusions from premises that may be tested for logical validity, conclusions and claims from a research project are often not as easily scrutinized and tested.
-Reviewers extend trust that methods and results were as written when assets can not be verified directly.
-Sharing research assets in the manner discussed here would greatly facilitate scrutinizing and testing all components of the research project, and thus potentially increase the reliability of the resulting scientific products.
+Unlike a set of conclusions from premises that may be tested for logical validity, research conclusions and claims are not as easily scrutinized and tested.
+Reviewers trust that methods and results were as written when assets can not be verified directly.
+Sharing in the manner discussed here would facilitate scrutinizing and testing research projects, and thus potentially increase the reliability of the resulting scientific products.
 
-Furthermore, organizing research assets in the manner discussed here allow its components to be formally tested, a process referred to as unit testing in software engineering.
-It refers to the practice of verifying that functions perform as intended in a variety of cases.
+Furthermore, organizing research assets in the manner discussed here allows computations to be formally tested, a process referred to as unit testing in software engineering.
+It is the practice of verifying that functions perform as intended in a variety of cases.
 We make a loose analogy between building a *vertical* project and unit testing.
@@
 Successfully building a *vertical* project is loosely analogous to performing unit tests that confirm and validate the process of asset creation.
-So, compiling a vertical project allows a researcher to unit test their own research project and verify its assets.
-By sharing a *vertical* project, other researchers can perform the same verification.
+So, compiling a project is a test of the research project and verification its assets.
+Other researchers can perform the same verification, if the project's source is shared.
 We note that *vertical* does not conduct proper unit tests on functions included in an R package.
@@
 Software developers routinely use version control platforms, such as GitHub and GitLab to coordinate software development across a team of collaborators [see, @VuorreCuratingResearchAssets2018].
-We recognize that scientific research is *not* software development.
-However, some components of the research workflow readily conform to the software development model---e.g., developing experimental programs, data analysis---and as such stand to benefit from borrowing best practices and tools from that domain.
+We recognize that scientific research is not software development.
+However, some components of the research workflow readily conform to the software development model---e.g., developing experiments and data analyses---and as such stand to benefit from borrowing best practices and tools from that domain.
 
-Research projects can be hosted on GitHub throughout the lifespan of the project to make use of tools for collaboration.
-With this in mind, *vertical* strongly suggests using Git (along with its online platforms).
+Research projects can be hosted on GitHub throughout the lifespan of the project to make use its collaboration tools.
+With this in mind, *vertical* suggests using Git (along with its online platforms).
 For example, we used GitHub as a collaboration tool to create *vertical* and to write this manuscript (using *papaja*).
@@
 Many users may find this feature very useful.
-In our case, we opened project related issues for discussion and then closed the issues once they were addressed.
+In our case, we opened issues for discussion and then closed the issues once they were addressed.
 A side-effect is that the GitHub issues tab, and the Git history, may double as a project development narrative, or a lab journaling system preserving issues and solutions for posterity.
@@
 
-*vertical* may be most beneficial to users who are already well-versed in R; and, although the website tutorial is intended for R beginners, it is not a tutorial on the R language.
+*vertical* may be most beneficial to users who already use R.
+Although our website tutorial is intended for R beginners, it is not a tutorial on the R language.
 Nevertheless, *vertical* could be used for pedagogical purposes, like a museum guide offering an orientation to collections in a large museum.
@@
 *vertical* is a modular open-source tool that can be extended by anyone.
-We added content modules that we think are common in many psychology research projects, and can incorporate new modules as they are identified and suggested by a user base.
+We included modules that we think are common, and can incorporate new modules as they are identified and suggested.
 Users can fork the *vertical* source code on GitHub, make suggested changes, and submit pull requests for review and inclusion.
@@
 
-Researchers may improve the reproducibility of their work by borrowing standards from software development, such as the R package standard, discussed here for sharing research products.
-By organizing research assets in a standardized format, their reuse will be less time consuming and error-prone.
-Standardized formats also facilitate large-scale meta-scientific investigations [@gorgolewski_brain_2016], and inclusion of data sets for meta-analyses.
-
-More generally, in addition to the convenience of a single-platform *vertical* workflow, adopting R standards during asset creation has the side effect of demanding some amount of attention on how research products are curated, stored, documented, and shared.
-Drawing attention to these details may serve to increase the reproducibility of our work.
+Borrowing standards and best practices from software development, such as the R package standard discussed here, may serve to dramatically improve reproducibility.
+By organizing research assets in a standard format, their reuse will be less time consuming and error-prone.
+Standard formats also facilitate large-scale meta-scientific investigations [@gorgolewski_brain_2016], and inclusion of data sets for meta-analyses.
+More generally, in addition to the convenience of a single-platform *vertical* workflow, adopting R standards during asset creation demands some amount of attention on how research products are curated, stored, documented, and shared.
+Drawing attention to these details may also serve to increase the reproducibility of our work.
 

### 2020-01-14 manuscript/manuscript.Rmd
fixed duplicate references

@@
 keywords: "reproducibility; research methods; R; open data; open science"
-wordcount: "`r wordcountaddin::word_count('manuscript.Rmd')`"
+wordcount: "$`r wordcountaddin::word_count('manuscript.Rmd')`$"
 bibliography: ["references.bib", "r-references.bib"]

### 2020-01-15 manuscript/manuscript.tex
update manuscript
fix word count

--- 2020-01-13 docs/manuscript/manuscript.tex
@@
 Correspondence concerning this article should be addressed to Matti Vuorre, 406 Schermerhorn Hall, 1190 Amsterdam Avenue MC 5501, New York, NY 10027. E-mail: mv2521@columbia.edu}
-\keywords{reproducibility; research methods; R; open data; open science\newline\indent Word count: 4787}
+\keywords{reproducibility; research methods; R; open data; open science\newline\indent Word count: `r wordcountaddin::word_count('manuscript.Rmd')`}
 \usepackage{dblfloatfix}
@@
 
-Research projects produce many assets, such as experiments, data, analyses, manuscripts, posters, slides, stimulus sets and materials, computational models, and more.
-Although creating and communicating research assets are mainstay activities in science, there is a trend (Vanpaemel, Vermorgen, Deriemaecker, \& Storms, 2015; Wicherts, Borsboom, Kats, \& Molenaar, 2006) encouraging more open and transparent sharing practices (Houtkoop et al., 2018; Lindsay, 2017).
-However, the value of shared assets is often limited by poor documentation, incompatible file formats, or lack of organization, resulting in low rates of reproducibility (Hardwicke et al., 2018).
-Some data sharing guidelines and good practices have been suggested (e.g., Rouder, Haaf, \& Snyder, 2019; Rouder, 2016); along with related issues of ethics and privacy, ownership of data, attitudes about sharing, and where to share data (Klein et al., 2018; Martone, Garcia-Castro, \& VandenBos, 2018).
-Nevertheless, curating standards for research assets have not emerged, possibly due to the variance of research in psychology.
+Research projects produce experiments, data, analyses, manuscripts, posters, slides, stimuli and materials, computational models, and more.
+However, the potential added value of these products is not fully realized due to limited sharing and curating practices.
+Although more transparent communication of these research products has recently been encouraged (Houtkoop et al., 2018; Klein et al., 2018; Lindsay, 2017; Martone, Garcia-Castro, \& VandenBos, 2018; Rouder, 2016; Rouder, Haaf, \& Snyder, 2019; Vanpaemel, Vermorgen, Deriemaecker, \& Storms, 2015; Wicherts, Borsboom, Kats, \& Molenaar, 2006), these efforts often focus narrowly on sharing data (and sometimes analysis code).
+Further, the practical value of sharing is often limited by poor documentation, incompatible file formats, and lack of organization, resulting in low rates of reproducibility (Hardwicke et al., 2018).
+Standardization of protocols for sharing would be beneficial; but, such standards have not emerged, possibly due to the variance of research in psychology.
 Instead of developing another standard, we suggest borrowing existing standards and practices from software engineering.
@@
 
-We sought for a standard for organization and sharing that would adhere to the FAIR (Findable, Accessible, Interoperable, Reusable) guidelines to maximize the reuse potential of data and support ``discovery through good data management'' (Wilkinson et al., 2016, see also general discussion).
-Additionally, we recognized the added value of including other research outputs (``products''; e.g., manuscripts) in a reproducible collection of materials that could be shared with the data.
-We identified the R package standard as a solution that doesn't present overwhelming overhead for already busy researchers.
+We sought for a standard for organizing and sharing that would adhere to the FAIR (Findable, Accessible, Interoperable, Reusable) guidelines to maximize the reuse potential of data and support ``discovery through good data management'' (Wilkinson et al., 2016, see also general discussion).
+Additionally, we recognized the added value of including other research outputs (``products''; e.g., manuscripts) beyond datasets in a reproducible collection of materials.
+We identified the R package standard with modern online-based workflows as a solution that doesn't present overwhelming overhead for already busy researchers.
 Here, we discuss the R package standard for creating reproducible research material containers for data, analysis code, and metadata.
-Then, we introduce a collection of additional R packages for incorporating other research products into the project.
+Then, we introduce additional R packages for incorporating other research products into the reproducible container.
 These additional tools are easily accessible through our \emph{vertical} package, thus named because an entire research project can be completed from top to bottom as a stack of R processes from within the RStudio software platform.
@@
 
-R is a programming language commonly used for statistical analyses (R Core Team, 2019), and is available at \url{https://www.r-project.org}.
-RStudio is an integrated development environment (IDE; RStudio Team, 2016), and is available at \url{https://www.rstudio.com/}.
-Both are free, open source and work on Windows, Mac, and Linux operating systems.
-We chose these tools because they are already widely used among psychology researchers: In one crowd-sourced analysis, 16 of 29 analysis teams used R (Silberzahn et al., 2018; see also Yee \& Debbie, 2017, and http://r4stats.com/articles/popularity/).
-Further, many psychologists have already developed R packages that share analysis tools and data.
+R is a programming language for statistical analyses (R Core Team (2019); \url{https://www.r-project.org}), and RStudio is its associated integrated development environment (IDE; RStudio Team (2016); \url{https://www.rstudio.com/}).
+Both are free, open source and work on Windows, Mac, and Linux systems.
+These tools are already widely used among psychology researchers: In one crowd-sourced analysis, 16 of 29 analysis teams used R (Silberzahn et al., 2018), and many psychologists have already developed R packages (see also Yee \& Debbie, 2017, and http://r4stats.com/articles/popularity/).
 
-A cornerstone of R are user-created packages, which contain R functions, data, metadata, and documentation written in a standardized format that allows seamless sharing across users and operating systems.
-When data and functions are wrapped in R packages, they are immediately available to other R users through the R console, and their documentation can be easily viewed in R or online.
-Next, we outline how data and functions are included in R packages, and then turn to including other assets, such as manuscripts and posters.
-We present a brief outline, a complete guide to creating R packages is outside the scope of this manuscript, and we refer readers to Wickham (2015), which remains an authoritative and in-depth guide on creating R packages.
+A cornerstone of R are user-created packages, which contain functions, (meta)data, and documentation in a standard format that allows seamless sharing across users and operating systems.
+Data and functions in R packages are immediately available to others through the R console, and documentation can be viewed in R or online.
+We first outline how data and functions are included in R packages, and then turn to including other assets, such as manuscripts and posters.
+A complete guide to R packages is outside our scope; for a complete treatment, see Wickham (2015).
 
@@
 
-Creating a package through RStudio is straightforward (see Figure \ref{fig:creating-package-rstudio}).
-When a new package is created with the method shown in Figure \ref{fig:creating-package-rstudio}, the required files and directories are automatically created.
-First, the ``DESCRIPTION'' file includes basic information about the package (e.g., authors, description) in a machine readable format.
-To edit it, users can click on the file in the RStudio File browser.
+When an R package is created as shown in Figure \ref{fig:creating-package-rstudio}, the required files and directories are automatically created.
+At this point, the package can be installed (click ``Install and Restart'' in RStudio's ``Build'' tab) and shared online (see below).
+However, to make it useful, content needs to be added, beginning with the package's description.
+That information (e.g., authors, dependencies) is written in the ``DESCRIPTION'' file in a machine readable format.
+To edit that file, click on its name in RStudio's File browser.
 The ``blank.Rproj'', ``man'' and ``NAMESPACE'' files/directories shown in Figure \ref{fig:creating-package-rstudio} should not be edited by users, and we thus turn to R functions.
@@
 
-Because R packages are primarily intended for R functions, an ``R'' directory is automatically created to store source code for defining functions.
 Researchers often develop and use custom functions in their analysis.
 Although it is possible to declare them directly in the analysis scripts, we suggest declaring functions using the R package standard, by placing the function's source code into a file in the ``R'' directory.
+That directory is automatically created (see Figure \ref{fig:creating-package-rstudio}).
 Functions in R packages are portable, such that others can install the package from their R console, load it, and start using the functions immediately.
-Packages can also depend on other packages (and be depended on), such that R will automatically include any other packages required for your functions to run appropriately.
-Functions within R packages are documented in a standardized manner, and the documentation for a function can be viewed within R or online.
+Packages can also depend on other packages (and be depended on), such that R automatically installs any requirements for your functions to work appropriately.
+Functions within R packages are documented in a standardized manner, and the documentation for a function can be viewed in R (e.g.~try \texttt{?mean}) or online.
 
-Additionally, learning and following R conventions for declaring functions has a pedagogical benefit to the researcher and may improve their practices.
-There is also a reuse benefit to the researcher: Functions can be difficult to find in old scripts, but easy to find and load if they are already packaged and called from an existing package.
+Learning and following R conventions for declaring functions has a pedagogical benefit to the researcher and may improve their practices.
+There is also a reuse benefit: Functions can be difficult to find in old scripts, but easy to find and load if they are called from an existing package.
 Thus, formally including one's functions in R packages facilitates reproducibility and sharing.
@@
 First, raw data in any format is stored in the ``data-raw'' directory to ensure that the method of sharing remains software-agnostic.
-Importantly, the raw data files should never be modified themselves.
-Second, instructions for converting and cleaning the raw data into analyzable format should also be included in this directory.
-These pre-processing steps should be declared so that the entire analysis is transparent and reproducible, preferably in an R script in the ``data-raw'' directory.
-That script should end with creating an R data object in the project's ``data'' directory.
+Second, instructions for converting and cleaning the raw data into analyzable format should also be included in this directory (the raw data files should not themselves be modified).
+These pre-processing steps should be saved so that the complete path from raw data to results is transparent and reproducible, preferably in an R script in the ``data-raw'' directory.
+That script should end with creating an R data object in the ``data'' directory.
 Placing the R data object there ensures that it is included in the resulting R package, which makes using the data effortless for any R user, and eschews the need to download additional files.
@@
 
-To recap, R package standards offer a systematic method for creating and curating reproducible research assets.
-Researchers could use the R package infrastructure to store and document their data and functions, and, as we will see, analyses, manuscripts, posters, and presentations as well.
-The entire research project is then easily shared online, and showcased as a website, as we will show below.
-To enable these additional features, many R packages have been developed (e.g., writing manuscripts within R).
+R packages thus provide a useful standard for functions and (meta)data.
+However, as we will see, the package structure can be extended to include analyses, manuscripts, posters, and presentations as well.
+The complete project is then easily shared online, and even showcased as a website, as we will show below.
+To enable these additional features, many R packages have been developed (e.g., writing manuscripts with R).
 To facilitate their use, we have created an R package, \emph{vertical}, that, when installed, also installs these other packages, and makes it easier to create projects that use them.
@@
 \emph{vertical} extends the R package template with a set of additional files and directories for products such as manuscripts and posters.
-These assets are organized such that they can then be easily shared online on GitHub or OSF, installed to R as an R package, and showcased online as a website.
+These assets are organized such that they can be easily shared online on GitHub or OSF, installed to R as a package, and showcased online as a website.
 Figure \ref{fig:vertical-workflow} illustrates the workflow of creating a \emph{vertical} research project.
@@
 
-Additional benefits of packaging everything together in this manner are that every computation supporting manuscripts, supplementary analyses, data preprocessing, etc.
-is reproducible and uses the same data in the same computational environment.
-Because sharing of materials is built-in, no additional work is needed after-the-fact in organizing the materials for sharing them.
+Additional benefits of packaging everything together in this manner are that the computations supporting manuscripts, supplementary analyses, data preprocessing, etc.
+are reproducible and use the same data in the same computational environment.
+Because sharing of materials is built-in, no additional work is needed after-the-fact in organizing the materials for sharing.
 We begin with an overview of this process, but a more exhaustive step-by-step tutorial is provided online (Crump \& Vuorre, n.d.)\footnote{\url{https://crumplab.github.io/vertical}}.
@@
 
-After installing the package, users should restart RStudio to make all its features available.
-First, similar to creating an R package as shown above, a new \emph{vertical} research project is created through the RStudio GUI (Figure \ref{fig:vertical-project-structure}.1).
+Users should then restart RStudio to make the features available.
+A new \emph{vertical} research project is created through RStudio (Figure \ref{fig:vertical-project-structure}.1).
 Figure \ref{fig:vertical-project-structure}.2 shows the resulting file and directory structure, which is an extension of the R package structure shown in Figure \ref{fig:creating-package-rstudio}.
-Users may enable or disable components (as indicated by directories like ``experiments'') as needed.
+Users may enable or disable components as needed, either by (un)checking boxes (Figure \ref{fig:vertical-project-structure}.1) or adding/deleting files/directories.
 Next, we describe the components, and then how to share the resulting research projects and showcase them as websites.
@@
 
-\caption{Creating a *vertical* project with the RStudio *vertical* project template, and description of the file-organization and structure of a *vertical* project.}\label{fig:vertical-project-structure}
+\caption{Creating a \emph{vertical} project in RStudio (1), and description of the file structure of a \emph{vertical} project (2).}\label{fig:vertical-project-structure}
 \end{figure*}
 
-In addition to data and functions, almost any kind of content can be included within the project directory structure.
-Most importantly, to create reproducible assets, such as data analyses or manuscripts, these materials should be created with the R Markdown language.
-R Markdown is a simple plain text format that allows mixing prose with code (R, Python, etc.), and thus enables including one's text and analyses in a single document.
+In addition to data and functions, almost any kind of content can be included within the project.
+Most importantly, to create reproducible assets, such as data analyses or manuscripts, these materials should be created with R Markdown, a simple language that allows mixing prose with code (R, Python, etc.), and thus enables including one's text and analyses in a single document.
 Because of its importance, we describe R Markdown first.
@@
 
-R Markdown is a plain text file format and an R package that allows executable code snippets to be embedded alongside regular text (Xie, Allaire, \& Grolemund, 2018).
-For example, an R Markdown file can include the typical prose found in manuscripts, the analysis code (R, Python, etc.), and code snippets that display numerical results and figures within the text.
+R Markdown allows executable code snippets to be embedded alongside regular text, and simple markup for formatting code (e.g headers using \texttt{\#}s; Xie, Allaire, and Grolemund (2018)).
 R Markdown documents are transparent and reproducible.
 In principle, recipients can see the analysis scripts, and reproduce them by compiling the document on their own machine.
-R Markdown documents separate content and style, and can be compiled to multiple output formats, such as PDF, HTML, and Microsoft Word.
-Importantly, R Markdown is basically plain text, and thus effortless to learn.
+R Markdown documents separate content and style, and can be compiled to multiple output formats, such as PDF, HTML, and Word.
+Importantly, R Markdown is plain text, and thus effortless to learn.
 Nevertheless, a tutorial on R Markdown is beyond our scope (see Xie et al., 2018)\footnote{See also \url{https://rmarkdown.rstudio.com/authoring_quick_tour.html} for a quick overview of R Markdown.}.
@@
 Above, we suggested placing any preprocessing steps in a separate file in the ``data-raw'' directory, and creating a documented R data object in ``data''.
-If the data object is called \texttt{mydata}, once the package is installed and loaded, simply calling \texttt{mydata} in an R script will call the data object, ready to be used in analyses.
+If the data object is called \texttt{mydata} and the package is installed and loaded, calling \texttt{mydata} in R will call the data object, ready to be used in analyses.
 Our favored approach is to write any analyses relevant to a product (e.g., manuscript or a poster) directly into the product's R Markdown source file.
@@
 
-Reproducible APA formatted manuscripts can easily be created with the \emph{papaja} R package, that provides an R Markdown template file, and additional helper functions for creating tables and figures (Aust \& Barth, 2018).
+Reproducible APA formatted manuscripts can be created with the \emph{papaja} R package, that provides an R Markdown template file, and additional helper functions for creating tables and figures (Aust \& Barth, 2018).
 For example, the R Markdown file can contain the text and code to generate the manuscript and all results.
-\emph{papaja} provides several functions for reporting results, especially from common statistical analyses such as ANOVAs, t-tests, and regressions.
-For example, \emph{papaja}'s \texttt{apa\_print()} function embeds statistical results directly into the manuscript, obviating the need to copy them by hand and removing human error in reporting.
-References can be automatically populated from Zotero using the \texttt{citr} plugin (Aust, 2019).
+\emph{papaja} provides several functions for reporting results: For example, \emph{papaja}'s \texttt{apa\_print(model)} embeds \texttt{model}'s statistical results directly into the manuscript, obviating the need to copy them by hand and removing human error in reporting.
+References can be automatically added from Zotero using the \texttt{citr} plugin (Aust, 2019).
 We refer readers to the \emph{papaja} documentation website for more information.\footnote{\url{https://crsh.github.io/papaja_man/introduction.html}}
@@
 
-\emph{vertical} includes a posters folder where poster documents can be deposited and shared on the website.
-Additionally, \emph{vertical} suggests using the \emph{posterdown} package (Thorne, 2019) for poster creation through an R Markdown document.
+\emph{vertical} includes a posters folder where poster documents can be deposited and shared.
+\emph{vertical} suggests using the \emph{posterdown} package (Thorne, 2019) for poster creation through an R Markdown document.
 When the provided template is used, the resulting reproducible document is viewable as part of the website.
@@
 There are several R Markdown options for creating reproducible slide decks in various formats.
-When a new \emph{vertical} project is created, a \emph{slidy} template is included to create web-ready presentations.\footnote{\url{https://bookdown.org/yihui/rmarkdown/slidy-presentation.html}} Slidy presentations are viewable in a browser like typical PowerPoint or keynote presentations, but have some additional features like long scroll-able slides, interactive figures and animations, and rendering full websites within individual slides.
+When a new \emph{vertical} project is created, a \emph{slidy} template is included to create web-ready presentations.\footnote{\url{https://bookdown.org/yihui/rmarkdown/slidy-presentation.html}} Slidy presentations are viewable in a browser like typical PowerPoint or Keynote presentations, but have additional features like scrollable slides, interactive figures, and animations.
 
@@
 
-Other research assets may include supplementary materials, stimulus sets, code for a computational model, additional analyses, or other documents, like a public-facing blog.
-\emph{vertical} is flexible with respect to the inclusion of additional assets not specified by the template.
+Other research assets may include supplementary materials, stimulus sets, code for a computational model, additional analyses, or other documents, such as blogs.
+Additional assets can be included by creating files in appropriate directories, and linking them to the resulting website (see below).
 For example, arbitrary R Markdown documents can be included in the vignettes directory.
-When the project is compiled (see below), any R Markdown files in that directory are compiled and served on the website.
+When the project is compiled (see below), R Markdown files in that directory are automatically compiled and served on the website (see Wickham and Hesselberth (2019)).
 
@@
 
-Finally, \emph{vertical} projects can include experiments in the experiments directory.
-These materials could be packaged as a compressed .zip file, which would then be accessible from an online repository.
+Finally, experiments should be placed in the experiments directory.
+They could be packaged as a compressed .zip file, which would then be accessible from an online repository.
 \emph{vertical} suggests using \emph{jsPsych} (De Leeuw, 2015), a JavaScript library for building browser-based experiments that can be served to the web or run locally.
 Selecting the \emph{jsPsych} module during project initialization downloads the most recent version of \emph{jsPsych} to the experiments folder.
-A basic html template is also provided, which loads the jsPsych library and presents some basic word stimuli.
-To further edit the experiment, users can simply edit the html file as they would when creating any other jsPsych experiment.
-When the \emph{vertical} project is compiled, this experiment is immediately available from the website navigation bar.
+A basic HTML template is also provided, which loads the jsPsych library and presents some basic word stimuli.
+To further edit the experiment, users can simply edit the HTML file as they would when creating any other jsPsych experiment.
+When the \emph{vertical} project is compiled, the experiment is available on the project website.
 
-For users who wish to work entirely within R Markdown, we suggest \emph{jspsychr} (Crump, 2019),\footnote{\url{https://crumplab.github.io/jspsychr/}} an R package that helps creating jsPsych experiments with R Markdown in RStudio.
-This option allows using RStudio's ability to pass objects between R and JavaScript when constructing experiment materials.
+For working entirely with R Markdown, we suggest \emph{jspsychr}, an R package that helps creating jsPsych experiments with R Markdown in RStudio (Crump, 2019).\footnote{\url{https://crumplab.github.io/jspsychr/}} RStudio enables passing objects between R and JavaScript, making this an attractive option for experienced R users.
 The template creates a sample \emph{jsPsych} experiment produced by \emph{jspsychr}, which contains the R Markdown script to generate the experiment file, the html file to run the experiment, and a folder to save the data and document the experiment.
 
-Browser-friendly experiments have several benefits over local-only experiments.
-First, scripts are open-source and easily reusable (no installation is required to run an experiment).
-Second, scripts can be shared in a self-contained manner such that the experiment is reproducible and the methods are transparent and verifiable.
-Third, because the experiment is an html file it can be shared and run as part of the project website, facilitating understanding of the experimental procedures.
-Sharing experiment source code with a demonstration is not yet common, but could benefit the review process and readers' understanding of the procedures by making methods more transparent.
+Browser-friendly experiments have several benefits.
+First, scripts are open-source and easy to use and share (no installation is required to run an experiment).
+Scripts can be shared in a self-contained manner such that the experiment is reproducible and the methods are transparent and verifiable.
+Third, because the experiment is an HTML file it can be shared and run as part of the project website, facilitating understanding of the experimental procedures.
+Sharing experiment source code with a demonstration could benefit the review process and readers' understanding of the procedures.
 
@@
 Above, we discussed using R Markdown, R package standards, and other R resources for creating research products.
-Then next step is sharing and communicating these products.
-When the products are created as suggested above, \emph{vertical} facilitates sharing the entire project as an R package for easy access to the data and functions, as a version controlled repository for access to the complete source code for reproducing every component of the project, and as a website, where the components are easily viewed by others.
+The next step is sharing and communicating these products.
+When the products are created as suggested above, \emph{vertical} facilitates sharing the entire project as an R package for easy access to the data and functions, as a version controlled repository for the project's source code, and as a website, where the components can be viewed in their final format.
 
@@
 
-An organizing principle of our approach is that all project materials are stored in a single directory on the user's local filesystem.
-When version control, such as Git, is enabled in this directory, the entire project can be seamlessly shared with others through online platforms such as GitHub and GitLab.
-Version control systems provide filesystems with additional functioning for keeping track of files' versions, improved options for collaborating, and seamless integration with online collaboration platforms (GitHub and GitLab); however, the details are outside the scope of this manuscript and we refer readers to Vuorre \& Curley (2018).
+The core of our proposal is that all project materials are stored in one directory on the user's local filesystem.
+When version control, such as Git, is enabled in this directory, the entire project can be seamlessly shared with others through online platforms such as GitHub, GitLab, and OSF.
+Version control systems provide filesystems with additional functioning for keeping track of files' versions, improved options for collaborating, and seamless integration with online collaboration platforms; however, the details are outside our scope, and we refer readers to Vuorre \& Curley (2018).
 By default, \emph{vertical} projects enable Git for new projects.
-Further, projects developed with Git (on a sharing platform such as GitHub or GitLab) can be seamlessly integrated with the Open Science Framework for long-term curation and, for example, to enable digital object identifiers (DOI).
 
@@
 
-An additional benefit of the R package organizing principles, and using the \emph{vertical} helper tools, is that the entire project can be showcased as a website (see Figure \ref{fig:vertical-project-website}.
-The website is compiled by \emph{pkgdown} (Wickham \& Hesselberth, 2019) in one command (\texttt{vertical::build\_vertical()}), or by clicking Addins -\textgreater{} Build Vertical in RStudio, and expertise with web standards is not required.
-For instructions on customizing the resulting website, see the \emph{pkgdown} instructions (\url{https://usethis.r-lib.org/}).
+An additional benefit of the R package organizing principles is that the entire project can be easily showcased as a website (see Figure \ref{fig:vertical-project-website}.
+The website is compiled by \emph{pkgdown} (Wickham \& Hesselberth, 2019) (click ``Addins'' -\textgreater{} ``Build Vertical'' in RStudio), which is automatically included with \emph{vertical} projects.
+For instructions on customizing the website, see \url{https://pkgdown.r-lib.org/}.
 
 Building a \emph{vertical} project as shown above creates the website in the docs folder, from where it can be viewed.
-During this process, all the R Markdown files within the project are evaluated while the documents are compiled, thus ensuring that all analyses, and thus products, are up to date.
-Importantly, the website can be easily made available on the internet.
-To do so, users should integrate their project with a GitHub repository (Vuorre \& Curley, 2018), and then enable GitHub pages under the repository's settings.
-An example of the resulting website can be viewed at \url{https://mvuorre.github.io/rparp/}.
+During this process, all the R Markdown files within the project are evaluated, ensuring that all analyses, and thus products, are up to date and use the same data in the same computational environment.
+The website is easily made available on the internet by integrating the project with a GitHub repository (Vuorre \& Curley, 2018), and then enabling GitHub pages under the repository's settings.
+See \url{https://crumplab.github.io/vertical/} for an example.
 
@@
 
-If the project contains R functions, data, and documentation following the R package standard (as discussed above), then others can install the package on their computers for fast access to the data and functions from within R.
-Again, integrating the project with GitHub is beneficial, because then others can install the package from GitHub using the \emph{devtools} package in R (\texttt{devtools::install\_github("account/repository")}).
+If the project contains R functions, data, and documentation following the R package standard, then others can install the package on their computers for fast access to the data and functions from R.
+Again, integrating the project with GitHub is beneficial, because then others can install the package from GitHub using the \emph{devtools} package, as was shown above in how to install \emph{vertical}.
 Sharing as an R package is very convenient for other R users, as any packaged data and functions are immediately available for use, natively in R, after package installation.
@@
 
-Our suggested approach has wider implications for making research practices in psychology more transparent and reproducible.
+Our suggestion has wider implications for research transparency and reproducibility.
 Consider the implied truth claims associated with any peer-reviewed manuscript.
 The act of publishing can be seen as a validity assertion: that the truth claims made by the authors are verifiably true, or more weakly that the research assets are verifiable.
-Unlike a set of conclusions from premises that may be tested for logical validity, conclusions and claims from a research project are often not as easily scrutinized and tested.
-Reviewers extend trust that methods and results were as written when assets can not be verified directly.
-Sharing research assets in the manner discussed here would greatly facilitate scrutinizing and testing all components of the research project, and thus potentially increase the reliability of the resulting scientific products.
+Unlike a set of conclusions from premises that may be tested for logical validity, research conclusions and claims are not as easily scrutinized and tested.
+Reviewers trust that methods and results were as written when assets can not be verified directly.
+Sharing in the manner discussed here would facilitate scrutinizing and testing research projects, and thus potentially increase the reliability of the resulting scientific products.
 
-Furthermore, organizing research assets in the manner discussed here allow its components to be formally tested, a process referred to as unit testing in software engineering.
-It refers to the practice of verifying that functions perform as intended in a variety of cases.
+Furthermore, organizing research assets in the manner discussed here allows computations to be formally tested, a process referred to as unit testing in software engineering.
+It is the practice of verifying that functions perform as intended in a variety of cases.
 We make a loose analogy between building a \emph{vertical} project and unit testing.
@@
 Successfully building a \emph{vertical} project is loosely analogous to performing unit tests that confirm and validate the process of asset creation.
-So, compiling a vertical project allows a researcher to unit test their own research project and verify its assets.
-By sharing a \emph{vertical} project, other researchers can perform the same verification.
+So, compiling a project is a test of the research project and verification its assets.
+Other researchers can perform the same verification, if the project's source is shared.
 We note that \emph{vertical} does not conduct proper unit tests on functions included in an R package.
@@
 Software developers routinely use version control platforms, such as GitHub and GitLab to coordinate software development across a team of collaborators (see, Vuorre \& Curley, 2018).
-We recognize that scientific research is \emph{not} software development.
-However, some components of the research workflow readily conform to the software development model---e.g., developing experimental programs, data analysis---and as such stand to benefit from borrowing best practices and tools from that domain.
+We recognize that scientific research is not software development.
+However, some components of the research workflow readily conform to the software development model---e.g., developing experiments and data analyses---and as such stand to benefit from borrowing best practices and tools from that domain.
 
-Research projects can be hosted on GitHub throughout the lifespan of the project to make use of tools for collaboration.
-With this in mind, \emph{vertical} strongly suggests using Git (along with its online platforms).
+Research projects can be hosted on GitHub throughout the lifespan of the project to make use its collaboration tools.
+With this in mind, \emph{vertical} suggests using Git (along with its online platforms).
 For example, we used GitHub as a collaboration tool to create \emph{vertical} and to write this manuscript (using \emph{papaja}).
@@
 
-\emph{vertical} may be most beneficial to users who are already well-versed in R; and, although the website tutorial is intended for R beginners, it is not a tutorial on the R language.
+\emph{vertical} may be most beneficial to users who already use R.
+Although our website tutorial is intended for R beginners, it is not a tutorial on the R language.
 Nevertheless, \emph{vertical} could be used for pedagogical purposes, like a museum guide offering an orientation to collections in a large museum.
@@
 \emph{vertical} is a modular open-source tool that can be extended by anyone.
-We added content modules that we think are common in many psychology research projects, and can incorporate new modules as they are identified and suggested by a user base.
+We included modules that we think are common, and can incorporate new modules as they are identified and suggested.
 Users can fork the \emph{vertical} source code on GitHub, make suggested changes, and submit pull requests for review and inclusion.
@@
 
-Researchers may improve the reproducibility of their work by borrowing standards from software development, such as the R package standard, discussed here for sharing research products.
-By organizing research assets in a standardized format, their reuse will be less time consuming and error-prone.
-Standardized formats also facilitate large-scale meta-scientific investigations (Gorgolewski et al., 2016), and inclusion of data sets for meta-analyses.
-
-More generally, in addition to the convenience of a single-platform \emph{vertical} workflow, adopting R standards during asset creation has the side effect of demanding some amount of attention on how research products are curated, stored, documented, and shared.
-Drawing attention to these details may serve to increase the reproducibility of our work.
+Borrowing standards and best practices from software development, such as the R package standard discussed here, may serve to dramatically improve reproducibility.
+By organizing research assets in a standard format, their reuse will be less time consuming and error-prone.
+Standard formats also facilitate large-scale meta-scientific investigations (Gorgolewski et al., 2016), and inclusion of data sets for meta-analyses.
+More generally, in addition to the convenience of a single-platform \emph{vertical} workflow, adopting R standards during asset creation demands some amount of attention on how research products are curated, stored, documented, and shared.
+Drawing attention to these details may also serve to increase the reproducibility of our work.
 
@@
 Retrieved from \url{https://www.psychologicalscience.org/observer/why-you-should-become-a-user-a-brief-introduction-to-r}
-
-\leavevmode\hypertarget{ref-R-citr}{}%
-Aust, F. (2019). \emph{Citr: 'RStudio' add-in to insert markdown citations}. Retrieved from \url{https://CRAN.R-project.org/package=citr}
-
-\leavevmode\hypertarget{ref-R-papaja}{}%
-Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}
-
-\leavevmode\hypertarget{ref-R-jspsychr}{}%
-Crump, M. J. C. (2019). \emph{Jspsychr: Templates and functions for writing and running jspsych experiments from r-studio}. Retrieved from \url{https://github.com/CrumpLab/jspsychr}
-
-\leavevmode\hypertarget{ref-R-vertical}{}%
-Crump, M.
-J.
-C., \& Vuorre, M.
-(n.d.).
-\emph{Vertical: Reproducible worfklow for psychological science research asset creation and communication}.
-Retrieved from \url{https://github.com/CrumpLab/vertical}
-
-\leavevmode\hypertarget{ref-de2015jspsych}{}%
-De Leeuw, J. R. (2015). JsPsych: A javascript library for creating behavioral experiments in a web browser. \emph{Behavior Research Methods}, \emph{47}(1), 1--12.
-
-\leavevmode\hypertarget{ref-gorgolewski_brain_2016}{}%
-Gorgolewski, K.
-J., Auer, T., Calhoun, V.
-D., Craddock, R.
-C., Das, S., Duff, E.
-P., \ldots{} Poldrack, R.
-A.
-(2016).
-The brain imaging data structure, a format for organizing and describing outputs of neuroimaging experiments.
-\emph{Scientific Data}, \emph{3}, 160044.
-\url{https://doi.org/10.1038/sdata.2016.44}
-
-\leavevmode\hypertarget{ref-HardwickeDataavailabilityreusability2018}{}%
-Hardwicke, T.
-E., Mathur, M.
-B., MacDonald, K., Nilsonne, G., Banks, G.
-C., Kidwell, M.
-C., \ldots{} Frank, M.
-C.
-(2018).
-Data availability, reusability, and analytic reproducibility: Evaluating the impact of a mandatory open data policy at the journal Cognition.
-\emph{Royal Society Open Science}, \emph{5}(8), 180448.
-\url{https://doi.org/10.1098/rsos.180448}
-
-\leavevmode\hypertarget{ref-HoutkoopDataSharingPsychology2018}{}%
-Houtkoop, B.
-L., Chambers, C., Macleod, M., Bishop, D.
-V.
-M., Nichols, T.
-E., \& Wagenmakers, E.-J.
-(2018).
-Data Sharing in Psychology: A Survey on Barriers and Preconditions.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(1), 70--85.
-\url{https://doi.org/10.1177/2515245917751886}
-
-\leavevmode\hypertarget{ref-KleinPracticalGuideTransparency2018}{}%
-Klein, O., Hardwicke, T.
-E., Aust, F., Breuer, J., Danielsson, H., Mohr, A.
-H., \ldots{} Frank, M.
-C.
-(2018).
-A Practical Guide for Transparency in Psychological Science.
-\emph{Collabra: Psychology}, \emph{4}(1), 20.
-\url{https://doi.org/10.1525/collabra.158}
-
-\leavevmode\hypertarget{ref-LindsaySharingDataMaterials2017}{}%
-Lindsay, D. S. (2017). Sharing Data and Materials in Psychological Science. \emph{Psychological Science}, \emph{28}(6), 699--702. \url{https://doi.org/10.1177/0956797617704015}
-
-\leavevmode\hypertarget{ref-MartoneDatasharingpsychology2018}{}%
-Martone, M. E., Garcia-Castro, A., \& VandenBos, G. R. (2018). Data sharing in psychology. \emph{American Psychologist}, \emph{73}(2), 111--125. \url{https://doi.org/10.1037/amp0000242}
-
-\leavevmode\hypertarget{ref-morey_peer_2016}{}%
-Morey, R.
-D., Chambers, C.
-D., Etchells, P.
-J., Harris, C.
-R., Hoekstra, R., Lakens, D., \ldots{} Zwaan, R.
-A.
-(2016).
-The Peer Reviewers Openness Initiative: Incentivizing open research practices through peer review.
-\emph{Royal Society Open Science}, \emph{3}(1), 150547.
-\url{https://doi.org/10.1098/rsos.150547}
-
-\leavevmode\hypertarget{ref-nosek_promoting_2015}{}%
-Nosek, B.
-A., Alter, G., Banks, G.
-C., Borsboom, D., Bowman, S.
-D., Breckler, S.
-J., \ldots{} Yarkoni, T.
-(2015).
-Promoting an open research culture.
-\emph{Science}, \emph{348}(6242), 1422--1425.
-\url{https://doi.org/10.1126/science.aab2374}
-
-\leavevmode\hypertarget{ref-R-base}{}%
-R Core Team. (2019). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
-
-\leavevmode\hypertarget{ref-Rouderwhatwhyhow2016}{}%
-Rouder, J. N. (2016). The what, why, and how of born-open data. \emph{Behavior Research Methods}, \emph{48}(3), 1062--1069. \url{https://doi.org/10.3758/s13428-015-0630-z}
-
-\leavevmode\hypertarget{ref-RouderMinimizingMistakesPsychological2019}{}%
-Rouder, J.
-N., Haaf, J.
-M., \& Snyder, H.
-K.
-(2019).
-Minimizing Mistakes in Psychological Science.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(1), 3--11.
-\url{https://doi.org/10.1177/2515245918801915}
-
-\leavevmode\hypertarget{ref-RStudioTeamRStudioIntegratedDevelopment2016}{}%
-RStudio Team. (2016). \emph{RStudio: Integrated Development Environment for R}. Boston, MA: RStudio, Inc. Retrieved from \url{http://www.rstudio.com/}
-
-\leavevmode\hypertarget{ref-SilberzahnManyAnalystsOne2018}{}%
-Silberzahn, R., Uhlmann, E.
-L., Martin, D.
-P., Anselmi, P., Aust, F., Awtrey, E., \ldots{} Nosek, B.
-A.
-(2018).
-Many Analysts, One Data Set: Making Transparent How Variations in Analytic Choices Affect Results.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(3), 337--356.
-\url{https://doi.org/10.1177/2515245917747646}
-
-\leavevmode\hypertarget{ref-R-posterdown}{}%
-Thorne, W.
-B.
-(2019).
-\emph{Posterdown: An r package built to generate reproducible conference posters for the academic and professional world where powerpoint and pages just won't cut it}.
-Retrieved from \url{https://github.com/brentthorne/posterdown}
-
-\leavevmode\hypertarget{ref-VanpaemelAreWeWasting2015}{}%
-Vanpaemel, W., Vermorgen, M., Deriemaecker, L., \& Storms, G.
-(2015).
-Are We Wasting a Good Crisis?
-The Availability of Psychological Research Data after the Storm.
-\emph{Collabra: Psychology}, \emph{1}(1), Art.
-3.
-\url{https://doi.org/10.1525/collabra.13}
-
-\leavevmode\hypertarget{ref-VuorreCuratingResearchAssets2018}{}%
-Vuorre, M., \& Curley, J.
-P.
-(2018).
-Curating Research Assets: A Tutorial on the Git Version Control System.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 219--236.
-\url{https://doi.org/10.1177/2515245918754826}
-
-\leavevmode\hypertarget{ref-Wichertspooravailabilitypsychological2006}{}%
-Wicherts, J.
-M., Borsboom, D., Kats, J., \& Molenaar, D.
-(2006).
-The poor availability of psychological research data for reanalysis.
-\emph{American Psychologist}, \emph{61}(7), 726--728.
-\url{https://doi.org/10.1037/0003-066X.61.7.726}
-
-\leavevmode\hypertarget{ref-R-testthat}{}%
-Wickham, H. (2011). Testthat: Get started with testing. \emph{The R Journal}, \emph{3}, 5--10. Retrieved from \url{https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf}
-
-\leavevmode\hypertarget{ref-WickhamPackagesOrganizeTest2015}{}%
-Wickham, H. (2015). \emph{R Packages: Organize, Test, Document, and Share Your Code}. "O'Reilly Media, Inc.". Retrieved from \url{http://r-pkgs.had.co.nz/}
-
-\leavevmode\hypertarget{ref-R-pkgdown}{}%
-Wickham, H., \& Hesselberth, J. (2019). \emph{Pkgdown: Make static html documentation for a package}. Retrieved from \url{https://CRAN.R-project.org/package=pkgdown}
-
-\leavevmode\hypertarget{ref-R-devtools}{}%
-Wickham, H., Hester, J., \& Chang, W. (2019). \emph{Devtools: Tools to make developing r packages easier}. Retrieved from \url{https://CRAN.R-project.org/package=devtools}
-
-\leavevmode\hypertarget{ref-WilkinsonFAIRGuidingPrinciples2016}{}%
-Wilkinson, M.
-D., Dumontier, M., Aalbersberg, I.
-J., Appleton, G., Axton, M., Baak, A., \ldots{} Mons, B.
-(2016).
-The FAIR Guiding Principles for scientific data management and stewardship.
-\emph{Scientific Data}, \emph{3}, 160018.
-\url{https://doi.org/10.1038/sdata.2016.18}
-
-\leavevmode\hypertarget{ref-R-rmarkdown}{}%
-Xie, Y., Allaire, J. J., \& Grolemund, G. (2018). \emph{R markdown: The definitive guide}. Boca Raton, Florida: Chapman; Hall/CRC. Retrieved from \url{https://bookdown.org/yihui/rmarkdown}
-
-\leavevmode\hypertarget{ref-yee_why_2017}{}%
-Yee, S.
-J.
-W., \& Debbie.
-(2017).
-Why You Should Become a UseR: A Brief Introduction to R.
-\emph{APS Observer}, \emph{30}(3).
-Retrieved from \url{https://www.psychologicalscience.org/observer/why-you-should-become-a-user-a-brief-introduction-to-r}
 \end{cslreferences}

### 2020-01-15 R/build.R

@@ -17,6 +17,7 @@ build_vertical <- function(clean=TRUE,...) {
 
   for (i in list.files(pattern = "\\.Rmd", recursive = TRUE)) {
     if(unlist(strsplit(i,split=.Platform$file.sep))[1] %in% c("experiments","vignettes","inst") == FALSE){
+      usethis::ui_info(paste("Knitting ", i))
       out.file <- rmarkdown::render(i, output_dir = paste0("docs/", dirname(i)), quiet=TRUE)
       if (!any(grepl(tools::file_path_sans_ext(i), readLines("_pkgdown.yml")))) {
         usethis::ui_info(paste(file.path(dirname(i),basename(out.file)),

### 2020-01-15 manuscript/manuscript.Rmd
Cut 5 words and fix typo.

@@
 We first outline how data and functions are included in R packages, and then turn to including other assets, such as manuscripts and posters.
-A complete guide to R packages is outside our scope; for a complete treatment, see Wickham [-@WickhamPackagesOrganizeTest2015].
+A complete guide to R packages is outside our scope (see @WickhamPackagesOrganizeTest2015).
 
@@
 
-An additional benefit of the R package organizing principles is that the entire project can be easily showcased as a website (see Figure \@ref(fig:vertical-project-website).
+An additional benefit of the R package organizing principles is that the entire project can be easily showcased as a website (see Figure \@ref(fig:vertical-project-website)).
 The website is compiled by *pkgdown* [@R-pkgdown] (click "Addins" -> "Build Vertical" in RStudio), which is automatically included with *vertical* projects.

### 2020-05-18 manuscript/manuscript.Rmd
Add Author Contributions section to MS
Revision draft 1

@@
 Although more transparent communication of these research products has recently been encouraged [@HoutkoopDataSharingPsychology2018; @LindsaySharingDataMaterials2017; @VanpaemelAreWeWasting2015; @Wichertspooravailabilitypsychological2006; @KleinPracticalGuideTransparency2018; @MartoneDatasharingpsychology2018; @RouderMinimizingMistakesPsychological2019; @Rouderwhatwhyhow2016], these efforts often focus narrowly on sharing data (and sometimes analysis code).
-Further, the practical value of sharing is often limited by poor documentation, incompatible file formats, and lack of organization, resulting in low rates of reproducibility [@HardwickeDataavailabilityreusability2018].
+Further, the practical value of sharing is often limited by poor documentation, incompatible file formats, and lack of organization, resulting in low rates of reproducibility[^reproducibility] [@HardwickeDataavailabilityreusability2018].
 Standardization of protocols for sharing would be beneficial; but, such standards have not emerged, possibly due to the variance of research in psychology.
@@
 
-Some advances in data-sharing standards have emerged: In much of psychological science it is now customary to share data on Open Science Framework (OSF).
+[^reproducibility]: By "reproducibility", we mean computational or analytic reproducibility by which previous results can be reproduced using existing data.
+Contrast this to "replicability", in which entire procedures are repeated to generate new data to test a scientific idea.
+
+Some advances in data-sharing standards have emerged: It is becoming more popular to share data on the Open Science Framework (OSF).
 However, those materials often contain idiosyncratic file organization and minimal or missing documentation for raw data.
@@
 
-Learning and following R conventions for declaring functions has a pedagogical benefit to the researcher and may improve their practices.
-There is also a reuse benefit: Functions can be difficult to find in old scripts, but easy to find and load if they are called from an existing package.
+Wrapping packages also provide a reuse benefit: Functions can be difficult to find in old scripts, but easy to find and load if they are called from an existing package.
 Thus, formally including one's functions in R packages facilitates reproducibility and sharing.
@@
 First, raw data in any format is stored in the "data-raw" directory to ensure that the method of sharing remains software-agnostic.
+If possible, that raw data should be in a text based format, rather than proprietary formats such as SPSS files.
 Second, instructions for converting and cleaning the raw data into analyzable format should also be included in this directory (the raw data files should not themselves be modified).
@@
 
-Finally, the resulting data object in "data" should be documented in a standardized format by placing a `data.R` file in the "R" directory, a process discussed in more detail in our complete online supplementary tutorial, and in @WickhamPackagesOrganizeTest2015.
+Finally, the resulting data object in "data" should be documented in a standardized format by placing a `data.R` file in the "R" directory, a process discussed in more detail in our complete online supplementary tutorial (https://crumplab.github.io/vertical/), and in @WickhamPackagesOrganizeTest2015.
 That standard format of documentation is especially useful because it allows datasets' documentation to be viewed in R (e.g., type `?attitude` in the R console) and online (see below).
@@
 When version control, such as Git, is enabled in this directory, the entire project can be seamlessly shared with others through online platforms such as GitHub, GitLab, and OSF.
-Version control systems provide filesystems with additional functioning for keeping track of files' versions, improved options for collaborating, and seamless integration with online collaboration platforms; however, the details are outside our scope, and we refer readers to Vuorre & Curley [-@VuorreCuratingResearchAssets2018].
+Version control systems provide filesystems with additional functioning for keeping track of files' versions, improved options for collaborating, and seamless integration with online collaboration platforms.
+For example, old versions of manuscripts are retained in the project's history.
+However, the details are outside our scope, and we refer readers to Vuorre & Curley [-@VuorreCuratingResearchAssets2018].
 By default, *vertical* projects enable Git for new projects.
+
+Because of the integration between Git(Hub) and one's local filesystem, users should primarily interact with Git(Hub).
+However, for long-term preservation, users should also link the project's GitHub repository to OSF.
+This can be done through the "Settings" page on an OSF project's page.
+Then, one can create an OSF registration of that project, which preserves the GitHub repository's current status on OSF permanently.
+This way, users gain both GitHub's convenience and OSF's long-term preservation.
 
@@
 Again, integrating the project with GitHub is beneficial, because then others can install the package from GitHub using the *devtools* package, as was shown above in how to install *vertical*.
+We recommend sharing most data packages through GitHub (as discussed here), rather than the official CRAN channel.
 Sharing as an R package is very convenient for other R users, as any packaged data and functions are immediately available for use, natively in R, after package installation.
 We expand on the extended benefits of using R packages for transparency and reproducibility in the general discussion.
+
+## Reproducibility
+
+An important component of computational reproducibility is specifying the (software) environment in which the computations were done, or should be done.
+R packages are a robust solution to this problem, because the package's metadata in the DESCRIPTION file can and should include the R packages and their versions that are required for the project.
+For example, if the project required using a specific version of R, and a specific version of an R package, this information could be written in the DESCRIPTION file.
 
@@
 And, the software we have discussed, RStudio, works well with other languages, including JavaScript, Python, and C++.
+
+## Comparison to other approaches
+
+We are not the first to recommend some form of standardization of workflows and procedures within psychology.
+First, Psych-DS is a collaborative project that aims to "promote the adoption of good practices in the management of scientific data" and "create a machine-readable format for these datasets that can support tools for analysis, discovery, and preparation of datasets in psychology" (https://github.com/psych-ds/psych-DS).
+Our suggestion is directly compatible with that effort, insofar as the R package standard doesn't require any special format from the raw data.
+In fact, future work might bring these two approaches together to inform both how the raw data should be formatted (Psych-DS), and how research products based on that raw data could best be curated and shared (present work).
+
+There are also at least two other R packages designed to help researchers organize their workflows and facilitate reproducibility and data/code sharing.
+We highlight these packages here, while also discussing their similarities and differences to our approach.
+*workflowr* is an R packages that makes it easy to create version controlled and well-organized repositories of research materials, and share them as websites on GitHub [@blischakCreatingSharingReproducible2019; https://github.com/jdblischak/workflowr].
+The *workflowr* R package itself is quite mature, and provides helpful tools for compiling many R Markdown documents sequentially, but has a slightly different approach to our proposal, in that workflowr does not itself provide much of a standardized organization for the materials (although it does provide a template).
+Further, the materials are not organized as an R package, as they would be in the (extended) *vertical* approach.
+Additionally, *workflowr* does not easily incorporate other types of materials (manuscripts, slides, posters, experiments) in the resulting website, unlike *vertical*.
+However, it may be slightly easier to use otherwise, for the many useful functions that it provides.
+
+Another approach is contained in the *rrtools* R package [@marwickPackagingDataAnalytical2018; https://github.com/benmarwick/rrtools], which, unlike *workflowr* but like our approach, centers on the R package architecture.
+Additionally, *rrtools* helps set up a Docker container and continuous integration services for the project, both tools that we hope become more widely adopted in the future.
+However, it does not help users set up other important research materials as part of the project's materials, such as slides and experiments, nor does it help users share their projects as websites.
+
+More generally, however, tools and descriptions, such as Psych-DS, *rrtools*, *vertical*, and *workflowr* are pointing to the same important and general idea: Psychological sciences would benefit from standardizing, as much as possible, the ways in which computational work is curated.
+We think that these tools and principles are in their infancy, and their importance is only now being recognized.
+Therefore, we expect these tools to evolve and mutate to best suit the practical researchers' needs, and hope that our suggestion provides some insight to how these tools might work in the future.
 
@@
 
+Future work should also identify commonalities between our suggestion, and other related work such as rrtools [@marwickPackagingDataAnalytical2018] and workflowr [@blischakCreatingSharingReproducible2019], and potentially work to unify these approaches to a more comprehensible and user-friendly tool for research project organization.
+
 ## FAIR
@@
 
+# Author Contributions
+
+MV and MC jointly generated the idea for the manuscript; MV wrote the first draft, which both authors then critically edited.
+*vertical* was programmed by MC and MV.
+Both authors approved the final submitted version of the manuscript.
+
 # References

### 2020-05-22 manuscript/manuscript.Rmd
More revisions to MS

@@
 
-An important component of computational reproducibility is specifying the (software) environment in which the computations were done, or should be done.
-R packages are a robust solution to this problem, because the package's metadata in the DESCRIPTION file can and should include the R packages and their versions that are required for the project.
-For example, if the project required using a specific version of R, and a specific version of an R package, this information could be written in the DESCRIPTION file.
+It is not uncommon to discover breaking changes in previously working code as a result of updates to a computational environment (e.g., a new language version or library dependency).
+As a result, an important component of computational reproducibility is specifying the (software) environment in which the computations were done, or should be done.
+In general, there is a continuum of static to dynamic approaches to this problem.
+A static approach could be a list of the requirements for the computational environment; whereas, a more dynamic approach could be a website or other container that actively runs the required environment (e.g., cloud computing tools such as Binder (https://mybinder.org/), CodeOcean (https://codeocean.com/), or RStudio Cloud (https://rstudio.cloud/)).
+
+For simplicity, we adopt R packages as a robust static solution.
+For example, an R package's metadata in the DESCRIPTION file can and should include the R packages and their versions that are required for the project.
+This provides a clear and standardized location for declaring the computational environment  necessary to run the code in the project.
+Conveniently, end users can opt to install the package with dependencies, which automatically installs libraries listed in the DESCRIPTION file.
+However, even though the DESCRIPTION may describe the necessary computational environment, an end user may have to take additional steps to create it themselves (e.g., installing R on their computers).
+Nevertheless, this static approach is not incompatible with dynamic approaches, and advanced users could conceivably adapt vertical websites to link to cloud computing environments that preserve and execute project scripts.
 
@@
 We highlight these packages here, while also discussing their similarities and differences to our approach.
-*workflowr* is an R packages that makes it easy to create version controlled and well-organized repositories of research materials, and share them as websites on GitHub [@blischakCreatingSharingReproducible2019; https://github.com/jdblischak/workflowr].
+*workflowr* is an R package that makes it easy to create version controlled and well-organized repositories of research materials, and share them as websites on GitHub (@blischakCreatingSharingReproducible2019; https://github.com/jdblischak/workflowr).
 The *workflowr* R package itself is quite mature, and provides helpful tools for compiling many R Markdown documents sequentially, but has a slightly different approach to our proposal, in that workflowr does not itself provide much of a standardized organization for the materials (although it does provide a template).
@@
 
-Another approach is contained in the *rrtools* R package [@marwickPackagingDataAnalytical2018; https://github.com/benmarwick/rrtools], which, unlike *workflowr* but like our approach, centers on the R package architecture.
+Another approach is contained in the *rrtools* R package (@marwickPackagingDataAnalytical2018; https://github.com/benmarwick/rrtools), which, unlike *workflowr* but like our approach, centers on the R package architecture.
 Additionally, *rrtools* helps set up a Docker container and continuous integration services for the project, both tools that we hope become more widely adopted in the future.

### 2020-05-22 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 
-We sought for a standard for organizing and sharing that would adhere to the FAIR (Findable, Accessible, Interoperable, Reusable) guidelines to maximize the reuse potential of data and support "discovery through good data management" [@WilkinsonFAIRGuidingPrinciples2016, see also general discussion].
+We sought a standard for organizing and sharing that would adhere to the FAIR (Findable, Accessible, Interoperable, Reusable) guidelines to maximize the reuse potential of data and support "discovery through good data management" [@WilkinsonFAIRGuidingPrinciples2016, see also general discussion].
 Additionally, we recognized the added value of including other research outputs ("products"; e.g., manuscripts) beyond datasets in a reproducible collection of materials.
@@
 Broadly, there are 3 steps to including data in an R package: 1.
-Placing raw data in the "data-raw" directory, 2.
+placing raw data in the "data-raw" directory, 2.
 creating an R script that processes the raw data and creates an R data object into the "data" directory, and 3.

### 2020-05-22 manuscript/manuscript.Rmd
Update affiliation

@@
     corresponding: yes
-    address: 406 Schermerhorn Hall, 1190 Amsterdam Avenue MC 5501, New York, NY 10027
-    email: mv2521@columbia.edu
+    address: 1 St Giles, Oxford OX1 3JS, United Kingdom
+    email: matti.vuorre@oii.ox.ac.uk
   - name: Matthew J. C. Crump
@@
   - id: "1"
-    institution: Department of Psychology, Columbia University, New York, USA
+    institution: Oxford Internet Institute, University of Oxford, United Kingdom
   - id: "2"
